### 2016-06-13 main.tex
Add back in the paragraph about non-matching data, now that that works (test running now...).

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

% Remove comment for double spacing
\usepackage{setspace} 
\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

%%% Begin BWP
%\usepackage{amsmath, amsthm, amssymb, wasysym, graphicx}
%\usepackage[small, hang, bf]{caption}
%\usepackage{natbib}
%\renewcommand\cite{\citep}
%\newcommand\citepossessive[1]{\citeauthor{#1}'s \citeyearpar{#1}}
\newcommand\eq[1]{Eq.~\ref{#1}}
\newcommand\fig[1]{Fig.~\ref{#1}}
\newcommand\sref[1]{Section~\ref{#1}}
\let\oldmarginpar\marginpar
\renewcommand{\marginpar}[1]{\oldmarginpar{\linespread{1}\scriptsize{#1}}}

% PLOS wants \paragraph for some reason...
\renewcommand{\subsubsection}[1]{\paragraph{#1}}

\setlength{\marginparwidth}{55mm}

\newcommand\argmin{\mathop{\textrm{{\rm argmin}}}\limits}
\newcommand{\noprint}[1]{}

\hyphenation{cross-valid-ation}
%%% End BWP

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
\begin{flushleft}
{\Large
\textbf\newline{A Fast and Accurate Zebra Finch Syllable Detector}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Ben Pearre\textsuperscript{1,a,\textcurrency},
L.~Nathan Perkins\textsuperscript{1,b},
Jeffrey E.~Markowitz\textsuperscript{2,c},
Timothy J.~Gardner\textsuperscript{1,d}
\\
\bigskip
\textsuperscript{1} Department of Biology, Boston University, Boston, Massachusetts, United States of America\\
\textsuperscript{2} Department of Neurobiology, Harvard Medical School, Boston, Massachusetts, United States of America
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.
Author \textsuperscript{a}~contributed the original MATLAB and LabVIEW implementations and most of the manuscript.
\textsuperscript{b}~contributed the Swift implementation and its description herein, helped debug the training code, and measured the timing for most of the detectors.
\textsuperscript{c}~performed the fibre photometry experiment, wrote several types of filter-bank software against which we compared, made suggestions for the neural network detector, and reformatted the MATLAB code for distribution.
\textsuperscript{d}~provided a stream of suggested techniques, metrics, and goals.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency a Insert current address of first author with an address update
% \textcurrency b Insert current address of second author with an address update
% \textcurrency c Insert current address of third author with an address update

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
\textsuperscript{\textcurrency} Corresponding author: bwpearre@gmail.com (BP)

\end{flushleft}

\reversemarginpar
%%% End PLOS header template

\begin{abstract}
  The song of the adult male zebra finch is strikingly stereotyped.
  Efforts to understand motor output, pattern generation, and learning
  have taken advantage of this consistency by investigating the bird's
  ability to modify specific parts of song under external cues, and by
  examining timing relationships between neural activity and vocal
  output.  Such experiments require that precise moments during song
  be identified in real time as the bird sings.  Various
  syllable-detection methods exist, but many require special hardware,
  software, and know-how, and details on their implementation and
  performance are scarce.  We present an accurate, versatile, and fast
  syllable detector that can control hardware at precisely timed
  moments during zebra finch song.  Many moments during song can be
  isolated and detected with false negative and false positive rates
  well under 0.5\% and 0.01\% respectively. The detector can run on a
  stock Mac Mini with triggering delay of less than a millisecond and
  a jitter of $\sigma\approx 2$ milliseconds.
\end{abstract}

\linenumbers

\section{Introduction}

The adult zebra finch ({\em Taeniopygia guttata}) sings a song made up
of 2--6 syllables, with longer songs taking on the order of a
second. The song may be repeated hundreds of times per day, is almost
identical each time, and several brain areas reflect this consistency
in highly stereotyped neural firing patterns. This consistency makes
the zebra finch one of the most popular models for the study of the
neural basis of learning, audition, and control.

This consistency allows a variety of experiments if precise moments in
song can reliably be detected quickly enough to trigger other
apparatus during singing.  A common area of study with this
song-triggering technique is the anterior forebrain pathway (AFP), a
homologue of mammalian basal ganglia consisting of a few distinct
brain areas concerned with the acquisition and learning of song.  For
example, \cite{Kao2005} stimulated the lateral magnocellular nucleus
of the anterior nidopallium (LMAN)---the output nucleus of the
AFP---at precisely timed moments during song and showed that this area
repeatably controls specific variables in song output.
\cite{Wang2008interhemispheric} stimulated LMAN and the high vocal
centre (HVC) in one hemisphere or the other, and found that control of
song rapidly switches between hemispheres.  \cite{Keller2009}
investigated the response of cells in field L and the caudolateral
mesopallium during song with and without feedback perturbations as
well as external sounds, and concluded that these regions may hold a
representation of song against which auditory signals are compared.
\cite{Tumer2007pitchshift} played a disruptive sound during renditions of a syllable that were
slightly above (or below) its average pitch, and showed that the
presence of this apparently random natural variability in songbird
motor output is used to drive change in the song.  \cite{Andalman2009}
extended this experiment and showed that AFP produces a corrective
signal to bias song away from those disruptions.  \cite{Warren2011}
showed that that AFP transfers this signal to the robust nucleus of
the arcopallium (RA) using NMDA-receptor--mediated glutamatergic
transmission. By looking at song recovery after applying the same
pitch-shift paradigm, \cite{Canopoli2014} showed that the caudal
medial nidopallium is implicated in memorising or recalling a recent
song target, but for neither auditory processing nor directed motor
learning.

% THEIR WORDS: (NCM, a high auditory area) impaired recovery of the original pitch even several weeks after withdrawal of the reinforcing stimuli. Because NCM lesions spared both successful noise-avoidance behavior and birds' auditory discrimination ability, our results show that NCM is not needed for directed motor changes or for auditory discriminative processing, but is implied in memorizing or recalling the memory of the recent song target.

Despite the power and versatility of such experiments, there is no standard syllable detector.  Considerations include:
\begin{description}
\item[Accuracy:] How often does the system produce false positives or false negatives?
\item[Latency:] The average delay between the target syllable being sung and the detection.
\item[Jitter:] The amount that latency changes from instance to instance of song.  Our measure of jitter is the standard deviation $\sigma$ of latency.
\item[Versatility:] Is detection possible at ``difficult'' syllables?
\item[Ease of use:] How automated is the process of programming a detector?
\item[Cost:] What are the hardware and software requirements?
\end{description}

A variety of syllable-triggering systems have been used, but few have
been documented or characterised in detail.  In 1999,
\cite{Leonardo1999} used groups of IIR filters with hand-tuned logical
operators.  Their system had a latency of 50 or 100 milliseconds (ms),
and they do not report on jitter or accuracy.  As access to
computational resources has improved, approaches have changed: in
2009, \cite{Andalman2009} still used hand-tuned filters, but ran them
on a Tucker-Davis Technologies digital signal processor.  They report
a latency of around 4 ms, but as with other filter-bank techniques, it
is not strictly a syllable detector but rather a pitch and timbre
detector---it cannot identify a frequency sweep, or distinguish a
short chirp from a long one---and thus requires careful selection of
target syllables.  Furthermore, the method is neither inexpensive nor,
based on our experience with a similar technique, accurate.
\cite{Keller2009} applied a neural network to a spectral image of
song.  They report a jitter of 4.3 ms, but further implementation and
performance details are not available.  In 2011, \cite{Warren2011}
matched spectral templates to stable portions of syllables in 8-ms
segments.  They report a jitter of 4.5 ms, and false-negative and
false-positive rates of up to 2\% and 4\%, respectively.  Hardware
requirements and ease of use were not reported.  In 2013,
\cite{Skocik2013} described in detail a system that matches spectral
images of template syllables using a correlation coefficient.  With a
fast desktop (Intel i7 six-core) running Linux and equipped with a
National Instruments data acquisition card, it boasts a hardware-only
(not accounting for the time taken to compute a match with a syllable)
latency and jitter of just a few microseconds, and the detection
computation they use should not much increase that.  Drawbacks are
that some hand-tuning may still be required, and that they report
false-negative rates around 4\% and 7\% for zebra finches and
Bengalese finches respectively, measured on a small dataset.  In much
other work, an allusion is made to a syllable detector, but no further
information is provided.

We developed a standalone detector that learns to match moments in the
song using a neural network applied to the song spectrogram, and
outputs a TTL pulse (a brief 5-volt pulse) at the chosen moment. The approach consists of
three steps:

\begin{enumerate}
\item Record and align a corpus of training songs.  The technique has been published in \cite{Poole2012}.
\item Choose one or more instants in the song that should create a trigger event, and train a
  neural network to recognise them. This step is carried out offline.  While any neural network software would produce equivalent results, we used MATLAB 2015b's neural network toolbox.
\item Once trained and saved, the neural network is used by a realtime
  detection program that listens to an audio signal and indicates detection of the target
  syllables via a TTL pulse.  We present three implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
\end{enumerate}
This method makes the following contributions:
\begin{itemize}
\item Fast: sub-millisecond latencies, with jitter around 2 ms.
\item Accurate: false positive rates under 0.02\% and false negative rates under 0.5\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
\item State-of-the-art performance with default parameters.
\item Runs on inexpensive hardware.
\item Described in detail here, with reference implementations provided and benchmarked.
\end{itemize}

%We present the method in Section~\ref{sec:method}.  Section~\ref{sec:results} presents our measurements.
%We
%conclude in Section~\ref{sec:conclusion}, and point to software
%resources in Appendix~\ref{sec:resources}.

\section{Materials and Methods}
\label{sec:method}

\subsection{Learning a detector}

We begin with a few hundred recordings of a given bird's song, in this
case made inside a sound-isolating chamber in which was mounted a
recording microphone (Audio-Technica AT803).  The songs are
time-aligned as described in \cite{Poole2012}.

We rely on two circular buffers:
\begin{description}
\item[Audio buffer:] This contains the most recent audio samples, and
  is of the length required to compute the Fast Fourier Transform
  (FFT)---usually 256 samples.
\item[FFT buffer:] The results of each new FFT are placed here.  It
  contains the $n_\textrm{fft}$ most recent FFTs, which will serve as
  inputs to the neural network (described below).
\end{description}

Audio is recorded at some sample rate $1/t_\mathrm{sample}$ (for example, 44.1 kHz), and new data are appended to the circular audio buffer.

The spectrogram is computed at regular intervals---the useful range
seems to be roughly 1--5 milliseconds, which we refer to as the frame
interval $t_\textrm{fft}$.  At each frame, a spectrum is computed from
the most recent 256 audio samples in the audio buffer, and the result
is appended to the FFT buffer.  For example, if $t_\textrm{fft}=1$ ms
and the recording sample rate $1/t_\textrm{sample}=40$ kHz, then in
order to compute a new FFT, $40000\cdot 0.001=40$ new audio samples
must be appended to the audio buffer, the FFT is computed using the
most recent 256 samples in that buffer, and the result is appended to
the FFT buffer.  Over time, the successive spectra will look something
like \fig{fig:song}.

\begin{figure}
  \includegraphics[width=\textwidth]{Fig1}
  \caption{This image was made by superposing the spectra of our 2818
    aligned songs.  Our example detection points, $t^*_1\ldots t^*_6$,
    are shown as red lines, with the recognition regions (30 ms
    $\times$ 1--8 kHz) marked as rectangles.}
  \label{fig:song}
\end{figure}

Note that this results in time being discretised into chunks of length
$t_\mathrm{fft}$.  Because each Fourier transform computation contains
a small number $n_\mathrm{fft}$ of new audio samples (in the example
above, 40 new samples and $256-40$ that have already been examined),
we tried implementing the Sliding Discrete Fourier transform (SDFT)
\cite{Jacobsen2003SlidingDFT}.  This allows
$t_\textrm{fft}=t_\textrm{sample}$.  Practically, the operating system
retrieves new audio samples from the hardware several at a time, so
the full benefit of the SDFT is difficult to see in practice.
Furthermore, we found that FFT implementations are sufficiently highly
optimised that discretising time into chunks of $t_\mathrm{fft}$ as we
have done produced similar results with simpler software.

When using consumer audio hardware that operates best at 44.1 kHz, the
requested frame interval may not line up with the sample rate, so the
actual frame interval may be different from the intended.  For
example, at 44.1 kHz a 1-ms frame interval requires a new FFT every
44.1 samples.  This must be rounded to 44 samples, resulting in
$t_\textrm{fft}=\lfloor 44.1 \rceil / 44.1 \approx 0.9977$ ms.

One or more target moments during the song must be chosen.  Our
interface presents the time-aligned spectrogram averaged over training
songs, and requires manual input of the target times, $t^*$.  Then we
assemble the training set from the song data, train the network,
compute optimal output unit thresholds, and save the network object
and an audio test file.

\subsubsection{Recognition region}
\label{sec:recognitionregion}

The neural network's inputs are the FFT values from a rectangular
region of the spectrogram covering a predefined range of frequency
values $F$ (for example, 1--8 kHz) at some number of the most recent
frames $n_\textrm{fft}$.  Any time $t$ falls within frame
$\tau(t)$, and $t-t_\textrm{fft}$ falls within the previous frame, so the recognition region that the neural network receives as input
consists of the spectrogram values over $F$ at $\tau(t)$ and those
from the contiguous set of recent frames: $T = \{$ $\tau(t)$,
$\tau(t-t_\textrm{fft})$, $\tau(t-{2t_\textrm{fft}})\ldots$
$\tau(t-n_\textrm{fft}t_\textrm{fft})\}$.  We have found that a 30-ms
time window---which will yield $n_\textrm{fft} = |T|=\lfloor 30\textrm{
  ms}/t_\textrm{fft}\rceil$ frames---of frequencies spanning 1--8 kHz
works well for most syllables.

Six examples of chosen target moments in the song, and their
corresponding recognition regions $F$ and $T$, are shown in
\fig{fig:song}.

%% \subsubsection{Song micro-realignment}

%% As conditions change, and especially during undirected song, syllable
%% length and relative timing may vary slightly, which introduces
%% variations in the precise timing of each syllable. The alignment
%% software we use ensures that songs are aligned at the point midway
%% through the song, but if the target syllable is not at that point, it
%% is helpful to re-align the songs at the point of interest.  This may
%% be accomplished by looking for peaks in the correlation of the
%% time-domain signal with the song whose spectrogram is closest to
%% average over the training set.\marginpar{I hope that's the right thing
%%   to do. Seems to work, anyway\dots}

\subsubsection{Building the training set}

The training set is created in a fashion typical for neural networks:
at each time frame $t$ the rectangular $|F|\times |T|$ recognition
region in the spectrogram as of time $t$ is reshaped into a vector
$\xi_t$, which will have length $|F||T|$ and contain the spectra in
$F$ taken at all of the times in the set $T$: from
$\tau(t-n_\textrm{fft}t_\textrm{fft})$ to $\tau(t)$.  These vectors
are placed into a training matrix, $\Xi$, such that each column
$\xi_t$ holds the contents of the recognition region---containing
multiple frames from the spectrogram---as of one value of $t$.

Training targets $y_t$ are vectors with one element for each desired
detection syllable.  That element is, roughly, 1 if the input vector
comes from the target time ($t=t^*$), 0 otherwise, for each target
syllable (of which there may be any number, although they increase
training time, and in our implementations the number of distinct
output pulses is constrained by hardware).  Since the song alignment
may not be perfect, and due to sample aliasing, a
strict binary target may ask the network to learn that, of two
practically identical frames, one should be a match and the other
not. Thus it is preferable to spread the target in time, such that at
the target moment it is 1, and at neighbouring moments it is
nonzero. We found that a Gaussian smoothing kernel around the target
time with a standard deviation of 2 ms serves well.

\noprint{
  % I screwed up: the results are without cage noise.  Rather than
  repeat all measurements, let's just not mention it.  }

With inputs well outside the space on which a neural network has been
trained, its outputs will be essentially random. In order to reduce
the false positive rate it is useful to provide negative training
examples that include silence, cage noise, non-song vocalisations, and
perhaps songs from other birds.  Although it will depend on the makeup
of the non-song data, we have found that training with as low as a 1:1
ratio of non-song to song yields excellent results.

\subsubsection{Normalisation}

In order to present consistent and meaningful inputs to the neural
network, we normalise the incoming data stream so that changes in the
song structure of the sound are emphasised over changes in volume, and
to maximise the effectiveness of the neural network's training
algorithm.

The first normalisation step is designed to eliminate differences in
amplitude due to changes in the bird's location and other variations
in recording.  Each recognition region vector $\xi_t$---during
training, each {\em column} of the training matrix $\Xi$---is
normalised using MATLAB's {\tt zscore} function $\hat{\xi}_t = (\xi_t
- \mu_{\xi_t}) / \sigma_{\xi_t}$, so that the content of each window
has mean $\mu_{\hat{\xi}_t}=0$ and standard deviation
$\sigma_{\hat{\xi}_t}=1$.

The second step is designed to ensure that the neural network's inputs
have a range and distribution for which the training function can
easily converge.  Each element $i$ of $\hat{\xi}$ is normalised across
the entire training set---each {\em row} of $\Xi$---in the same way:
$\check{\xi}^i = (\hat{\xi}^i -
\mu_{\hat{\xi}^i})/\sigma_{\hat{\xi}^i}$, so that the values of that
point across the whole training set have mean $\mu_{\check{\xi}^i}=0$
and standard deviation $\sigma_{\check{\xi}^i}=1$.  This is
accomplished during training by setting MATLAB's neural network
toolbox normalisation scheme to {\tt mapstd}, and the scaling
transform is saved as part of the network object used by the realtime
detector so that it may be applied to unseen data at runtime.

These two normalisation steps provide a set of inputs that are more
robust to outliers and less likely to produce false positives during
silence than other normalisation schemes, such as linear or
L\textsubscript{2} normalisation.

\subsubsection{Neural Networks}

While any learned classifier might suffice, we chose a two-layer
feedforward neural network.  In brief, our network takes an input
vector $\xi_t$---as described above---and produces an output vector
$y_t$, and when any element of $y_t$ is above a threshold (described
below), the detector reports a detection event.  The network uses two
matrices of weights, $W_0$ and $W_1$, and two vectors of biases, $b_0$
and $b_1$.  The first takes the input $\xi_t$ to an intermediate
stage---the ``hidden layer'' vector.  To each element of this vector
is applied a nonlinear squashing function such as {\tt tanh}, and then
the second weight matrix $W_1$ is applied.  This produces output
$y_t$:
\begin{equation*}
  y_t = W_1 \tanh (W_0 \xi_t + b_0) + b_1
\end{equation*}
During the network's training phase, the elements of the matrices and bias vectors are learned by back-propagation of errors
over a training set.  A more detailed explanation of neural networks
may be found in \cite{hkp}.

Essentially, after training, the network is available in the form of
the two matrices $W_0$ and $W_1$ and the two vectors $b_0$ and $b_1$,
and running the network consists of two matrix multiplications, two
vector additions and the application of the squashing function.

\subsubsection{Training the network}

The network is trained using MATLAB's neural network toolbox. We tried
a variety of feedforward neural network geometries, from simple
1-layer perceptrons to geometries with many hidden nodes, as well as
autoencoders. Perhaps surprisingly, even the former yields excellent
results on many syllables, but a 2-layer perceptron with a very small
hidden layer---with a unit count 2-4 times the number of target
syllables---was a good compromise between accuracy and training
speed. For more variable songs, deep structure-preserving networks may
be more appropriate, but they are slow to train and unnecessary for
zebra finch song.

We used the Levenburg-Marquardt algorithm, which is the MATLAB
toolbox's default.  It is a fast algorithm, but is memory-intensive,
so many target syllables or high FFT frame rates require a large
amount of RAM and increase training time to hours. Other training
algorithms that use less RAM are much slower, and by default they will
often terminate prematurely due to their performance gradient going to
0.

\subsubsection{Computing optimal output thresholds}
\label{sec:optimalthresholds}
After the network is trained, outputs of the network for any input are
now available, and will be in (or, due to noisy inputs and imperfect
training, close to) the interval (0, 1). We must choose a threshold
above which the output is considered a positive detection. Finding the
optimal threshold requires two choices. The first is the relative cost
of false negatives to false positives, $C_n$. The second is the
acceptable time interval: if the true event occurs at time $t$, and
the detector triggers at any time $t\pm\Delta t$, then it is
considered a correct detection. Then the optimal detection threshold
is the number that minimises $\textrm{[false positives]}
+C_n\cdot\textrm{[false negatives]}$ over the training set, using the
definitions of false positives and negatives given in
Section~\ref{sec:accuracy}. Since large portions of the cost function
are flat, hillclimbing is ineffective.  Random-restart hillclimbing
would work, but a brute-force search requires fractions of a
second. For the results presented here, we have used $C_n=1$ and
$\Delta t=10$ ms.

\subsubsection{De-bouncing}

During runtime, the network may produce above-threshold responses to nearby
frames.  Thus, after the first response, subsequent responses are suppressed
for 100 ms.  However, for the accuracy figures presented here, we used
the un-de-bounced network response.

\subsubsection{Our parameter choices}

We used an FFT of size 256; a Hamming window; and chose a target
spectrogram frame interval of $t_\textrm{fft}=1.5$ milliseconds,
resulting in a true frame interval of $t_\textrm{fft}=\lfloor 1.5\cdot
44.1\rceil /44.1\approx 1.4966$ ms.  We set the network's input space
to 30 ms long, and to span frequencies from 1--8 kHz, which contains
the fundamentals and several overtones of most zebra finch
vocalisations.

We found these parameters to work well across a variety of target
syllables, but various other parameter sets yield results similar to
those presented here.  Some of the parameters trade off detection
accuracy or temporal precision vs.~training time. For example,
decreasing the frame interval generally decreases both latency and
jitter, but also increases training time.

\subsection{Realtime detection}

The architecture of the realtime detector requires that the most
recent $n_\textrm{fft}$ spectrograms be fed to the neural network
every frame interval.  Audio samples from the microphone are appended
to the circular audio buffer.  Every $t_\textrm{fft}$ seconds a new
spectrogram is calculated by applying the Hamming window to the
contents of the buffer, performing an FFT, and extracting the
power. Outputs of the spectrogram from the target frequency band are
appended to the circular FFT buffer.  The spectrograms are sent to a
static implementation of the previously trained neural network.

We tested three implementations of the realtime detector.  For all of
these tests, we ran the detector processes under the operating
systems' default schedulers and process priorities, running typical
operating system daemon processes but no user loads.  The computers
had ample memory resources.

\subsubsection{Swift}

This detector uses the Swift programming language and Core Audio
interface included in Apple's operating systems.

The Core Audio frameworks provide an adjustable hardware buffer size
for reading from and writing to audio hardware (different from our two
circular buffers). Tuning this buffer size provides a tradeoff between
the jitter in the detection and the processor usage needed to run the
detector. We used buffer sizes ranging from 8 samples (0.18 ms at 44.1
kHz) to 32 samples (0.7 ms at 44.1 kHz) depending on the frame size 
used by the detector.

Vector operations---applying the Hamming window, the FFT, input
normalisation, matrix multiplication, and the neural network's
transfer functions---are performed using the Accelerate framework
(vDSP and vecLib), which use modern vector-oriented processor
instructions to perform calculations.

When the neural network detects a match, it instructs the computer to
generate a TTL pulse that can be used to trigger downstream hardware.
This pulse can be either written to the computer's audio output buffer
(again, in 8- to 32-sample chunks) or sent to a microcontroller
(Arduino) via a USB serial interface. Sending the trigger pulse via
the serial interface and microcontroller is noticeably faster (2.2 ms
lower latency), likely due to the fact that the audio buffer goes
through hardware mixing and filtering prior to output.

The above code can be run on multiple channels of audio on consumer
hardware (such as a 2014 Mac Mini) with little impact on CPU usage
($<15\%$). Depending on the experimental needs, latency can potentially
be further decreased (at the expense of processor usage) by adjusting the audio
buffer sizes.

We ran the Swift detector on a Late 2014 Mac Mini with a Intel Core i5
processor at 2.6GHz  with 16 gigabytes of RAM, running Mac OS X 10.11.

% technically, it is LabVIEW
\subsubsection{LabVIEW}

This implementation requires special software and hardware: LabVIEW
from National Instruments---we used 2014 service pack 1---and a data acquisition card---we use the
National Instruments PCI-6251 card on a PC with an Intel Core i5-4590
processor at 3.7GHz (a relatively low-end machine) with 32 gigabytes
of RAM, running Microsoft Windows 8.1 Pro.

This implementation has several drawbacks: it requires Windows and
expensive hardware and software and due to the programming language it
is difficult to modify and debug---indeed, a persistent bug in our
implementation currently renders it substantially less accurate than
the other detector implementations on some syllables. However, our
test configuration usually achieved excellent performance, and further
gains should be possible if the implementation were retargeted onto
field-programmable gate array (FPGA) hardware---which would have the additional benefit of providing
deterministic ``hard realtime'' guarantees---or just run on a faster
desktop system.

\subsubsection{MATLAB}

This detector uses the built-in audio input and output hardware on a
compatible computer.  We tested on a 2014 Mac Mini (the same machine used
for the Swift detector described above) and 2015 Mac Pro.
The Mac Pro does not have an audio input jack, so input was through an
M-Audio MobilePre external USB audio interface.  Despite the faster
processor, the latter system did not achieve higher performance than
the former, due to USB data transfer overhead.

Because of how MATLAB's DSP toolbox interfaces with audio hardware,
there is a double buffer both reading from and writing to audio
hardware. As a result, much of the code focuses on a lightweight audio
hardware interface, in order to have the smallest achievable audio
buffer. To help achieve this, the MATLAB implementation spreads data
acquisition and processing across two threads, due to the higher
computational overhead of the interpreted programming language.

The most versatile implementation, MATLAB runs on a variety of
hardware and operating systems, and is perhaps the easiest to modify.
While it did not perform as well on our test system as the other
implementations, the convenience may outweigh the slight timing
performance penalty for some experiments.  Key to minimising jitter is
the size of the audio buffer: on a 2014 Mac Mini running MATLAB 2015b
the smallest buffer size that did not result in read overruns was
about 4 ms.

Similar to the Swift implementation, it is also possible to modify the
MATLAB implementation to generate the TTL pulse from a microcontroller
(Arduino) controlled via a serial over USB interface. This eliminates
the double buffer required when sending triggers via the audio
interface, reducing latency by half.

\subsection{Quantification}
\label{sec:quantify}

We divided the data into training and test sets.  For the results
reported in the following section, the training set consisted of 1000
instances of the song of one bird.
Our test set consisted of an additional 1818 iterations of the bird's
song.

The MATLAB neural network toolbox further divides our ``training'' set
into internal training, validation, and test sets.  We did not stray
from the toolbox's default values, and do not discuss MATLAB's
internal subdivision of our training set.  Because the dataset
consists of temporally aligned songs, ground truth is available for
each song (albeit with minor variations due to the alignment process
used \cite{Poole2012}), and thus the detector can be checked by
presenting the recorded training songs as well as the canonical
detection events. To this end, besides the trained network object, our
learning code produces an audio file consisting of all of the training
data on the left audio channel and a delta function at each aligned
moment of target syllable presentation on the right channel. Thus,
when played on any audio player, the left channel may be provided as
input to the detector, and the the detector's output pulses may be
compared against the right channel.

\subsubsection{Accuracy}
\label{sec:accuracy}

We define the accuracy of the network based on its classification
performance per frame. In order to avoid the apparent problem of every
non-detected non-syllable counting as a true negative, we also tried
defining accuracy on a per-song basis, such that a song without the
target syllable counted as a single true negative.  Computing the
optimal output thresholds on a per-frame basis resulted in higher
thresholds and thus a lower false-positive rate, with minimal
consequences to the false-negative rate, while also providing a valid
definition of false-positive rate for data streams that had not been
segmented into songs.

%% We define accuracy on a per-song basis, as follows:

%% \begin{description}
%%   \item[True positive:] A song for which the detector fires within
%%     20ms of the intended moment.
%%   \item[True negative:] So that every non-detected non-syllable does
%%     not count, a true negative is a complete song during which the
%%     detector did not fire at any point outside the true positive
%%     region.
%%   \item[False positive:] A detection event more than 20ms from the
%%     target syllable.\marginpar{Or a song with $>0$ false pos?}
%%   \item[False negative:] A song's target interval during which a
%%     detection event should have occurred, but didn't.
%% \end{description}

The accuracy as defined above is used for computing the optimal
thresholds above which the network's output should be interpreted as a
match on the training data as described in
\sref{sec:optimalthresholds}, for evaluation of the detectors on the
training songs, and while live.

\noprint{ Since the network will output values $o_t$ between 0 and 1
  at each moment $t$ in an attempt to match the training output, the
  optimal threshold $\tau\in[0,1]$ for the output neuron should be
  computed.  Given the relative cost of false positives vs.~false
  negatives $C$, and the acceptable time difference between target
  syllable and correct output $\Delta t_d$, we compute the optimal
  threshold for an output element according to the definitions above:
\begin{eqnarray*}
  \textrm{true positives}_\tau &=& \textrm{size of set}_{s\in \textrm{target songs}} o_t > \tau, \left| t \leq \Delta t_d \right| \\
  \textrm{false negatives}_\tau &=& \textrm{size of set} {s\in\textrm{target songs}} - \textrm{size of set} \textrm{true positives} \\
  \textrm{false positives}_\tau &=& \textrm{size of set}_{s\in \textrm{target songs}} o_t > \tau, \left| t > \Delta t_d \right| \\
  \widehat{\tau} &=& \argmin_\tau C\textrm{false positive} + \textrm{false negatives}
\end{eqnarray*}
}

\subsubsection{Timing}

We evaluate the time taken from the presentation of the target
syllable to the firing of the detector's TTL pulse. While playing the
audio test file from any audio playback device, the TTL output from
the ground-truth channel of the audio output may be used as the
trigger pulse for an oscilloscope, and compared to the TTL pulse
produced by the detector implementation, which sees only the birdsong
channel of the audio file. For this purpose we used a pulse generator
(Philips PM 5715, with a listed latency of 50 ns, jitter of $\leq
0.1\%$ or 50 ps, whichever is greater) to widen the detector's output
spike to a number much larger than the jitter ($\sim 100$ ms).  This
obviates pulse length variability in the output device by essentially
discarding the falling edge of the output pulse.  The oscilloscope is
then set to averaging mode (128-trigger average) in order to collect
timing data. The canonical signal is the trigger at $t=0$, and the
average of the detector's detection events will be seen as a
low-to-high transition with form approximating the cumulative
probability distribution function (CDF) of the detector's output in
response to the chosen song event, with individual detection events visible as steps in this curve.

% Use to say: "Mean latency is then given as the halfway point of that detection 
% sigmoid."
%
% I don't think this description matches the calculation, this description
% seems to better match a "median" latency. The mean latency is calculated
% as an average latency of all detection events, which is the same as 
% fitting a normal distribution to the detection events and taking the mean
% of the normal distribution (the MLE estimator is the same as the sample mean).

Mean latency is then given as the average time between the 
canonical signal and the corresponding detection event. 
It is a helpful number, but not a critical one, since a
detector with high but constant latency can often be trained to
trigger at a point somewhat before the true moment of interest,
especially given the stereotypy of zebra finch song.  Often more
important is latency jitter: how much variability is there in the
latency?  We define jitter as the standard deviation of the latency.

When measuring timing, it is useful to compare against a theoretical
optimal, in order to control for any imprecision in the song
extraction and alignment of our real data inherent in the method we
use (described in \cite{Poole2012}).  We propose two measures thereof:

First we test the ``ideal'' latency and jitter in the time shift used
in calculating new columns of the spectrogram. By passing a recorded
audio sequence into the detector offline, and assuming no input or
output latency, we compare how many additional audio samples beyond
the syllable are needed before the spectrogram can match the inputs
needed to trigger the neural network.  This latency reflects the FFT
size used for calculating the spectrogram, the FFT time shift between
columns in the spectrogram, and the width of the Gaussian smoothing
kernel applied to the ground truth data when training the neural
network, but ignores computation times, audio buffers, and other
operating system overhead.

Next we use a ``$\delta$-syllable'' consisting of a $\delta$ function
in the time domain, and train the network to trigger 5 ms after this
pulse.  This song is fed into the live detector.  The results for this
measurement show the latency and jitter inherent in the complete
detector including audio read and write buffers, classifier
computation, and FFT time shift aliasing, but excluding imprecision in
the song alignment as well as detection timing effects due to
differences between instances of the bird's song.

Finally, we measure latency on real bird song aligned as in
\cite{Poole2012}.  This extracts and aligns the centres of the sampled
songs, and the canonical signal is given with respect to that
alignment.  These timing measurements reflect not only detector
performance, but also the variability of the bird's song.

\subsection{Fibre Photometry}
\label{sec:fp-methods}

In addition to synthetic tests of detector accuracy and 
latency, we piloted the syllable detector in an experiment
providing altered auditory feedback (AAF) to zebra finches. AAF, or 
playback of a perturbing white noise stimulus, is known to disrupt 
the highly-stereotyped zebra finch song. By probabilistically 
targeting the AAF to a specific syllable, it is possible 
to selectively destabilize parts of the song. During this 
feedback, signal from genetically-encoded calcium indicators was 
recorded via fibre photometry to identify a neural representation of 
an error signals associated with the feedback.

The initial experiment design used a DSP with hand-designed filter-banks to
detect syllables, but these required tuning and were only conducive
to detecting harmonic stacks. This experiment provided an ideal design
to pilot our neural-network-based syllable detector.

\subsubsection{Surgery}

GCaMP6s and GCaMP6f were expressed using a lentivirus under the RSV 
promoter \cite{Markowitz:2015ko}. Injections were targeted under HVC 
using stimulation of Area X with a bipolar stimulation electrode. 
After injecting the virus, we implanted a fibre-optic cannula above 
HVC ($400 \mu m$ diameter, Doric Lenses). The location of the injection 
was verified histologically after each experiment.

\subsubsection{Photometry}

Fibre photometry \cite{Adelsberger:2005dy,Schulz:2012fv,Cui:2013dq,Adelsberger:2014jd,Gunaydin:2014dh} 
was performed in N=3 birds at least 2 weeks post-injection using a 
completely fibre-coupled light path. A 470 nm fibre-coupled LED (Mightex) 
was sinusoidally driven at 5 kHz by a TDT RX8. The output of the LED 
was filtered with a GFP excitation filter (Doric Lenses Fluorescence Cube) 
and coupled to an optical rotary joint (Doric Lenses). The rotary joint 
was connected via a multimode fibre optic patch cord (400 $\mu$m, .48 NA, 
Doric Lenses) to the implanted cannula using a brass sleeve (Doric Lenses). 
LED power was calibrated such that optical power at the tip was 
20--300 $\mu$W (LED was turned on only during vocalizations, detected using 
a TDT RX8). Fluorescence was collected through the same patch cord, 
filtered with a GFP emission filter, and sent to a silicon photomultiplier 
(SensL MiniSM 30000). The output of the photomultiplier was demodulated 
using a dual-phase lock-in amplifier (10-ms time-scale, SR830, Stanford 
Research Systems) and digitized along with the bird's song at 24.414 kHz 
using a National Instruments data acquisition board (PCIE-6323).

\subsubsection{Syllable detection}

User-selected syllables were detected using the Swift implementation of the
syllable detector, and the TTL output triggered altered auditory feedback 
via the TDT RX8.

\subsubsection{Data analysis}

$\Delta F / F0$ was estimated at each point by subtracting and dividing by 
F0, defined as the 11th percentile computed in a 300-ms sliding window. 
Data were then smoothed with a 50-ms exponential window. Fluorescence traces
during singing were normalized by subtracting the trial-average signal, in 
order to remove song-based modulation.

\subsection{Ethics}

All procedures were approved by the Institutional Animal Care and Use Committee of Boston University (protocol number 14-029).

\section{Results}
\label{sec:results}

\fig{fig:song} shows the song for which we present accuracy and timing results.  Six
example target trigger points are shown, at 150, 200, 250, 300, 350,
and 400 milliseconds after the beginning of the aligned samples, as
indicated by the red lines, and referred to henceforth as $t^*_1\ldots
t^*_6$ respectively.  The rectangles show the recognition window (30
ms, 1--8 kHz) for each instant in the song.

The first two triggers, $t^*_1$ at 150 ms and $t^*_2$ at 200 ms, are
far from pure tones, but are closer to coloured noise.  They would be
difficult for a harmonic-stack filter bank to detect, and especially to
pinpoint in time.  $t^*_3$ (250 ms) and $t^*_6$ (400 ms) are rich in
overtones and contain downward slides with changing timbres.  $t^*_4$
(300 ms) occurs near the beginning of a more typical harmonic stack amenable to a variety of
recognition techniques, although consistently detecting a point
partway through the syllable demands a detector that can use multiple
time steps.  $t^*_5$ (350 ms) shows a steady pitch followed by a
complex changing tone structure.

An example of the detector's output for three of these syllables is
shown in \fig{fig:detection_raster}.  The total audio energy of each song is shown as a single grayscale row.  By sorting according to time of
each syllable's detection, the rest of the song is shown in the
context of the timing for that moment.  This gives an intuition of the
timing variability within each song and across songs (which may be responsible
for some of our measured jitter).

\begin{figure}
  \includegraphics[width=\textwidth]{Fig2}
  % 9x10 inches
  % Fonts: Helvetica, scaled 130%
  \caption{Each plot shows one network output unit's responses to all
    2818 presentations of the song shown in \fig{fig:song}.  For
    simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
    $t^*_6$. The horizontal axis is time relative to the beginning of
    the aligned song, and the vertical axis is an index for the 2818
    individual song presentations. The grey shading shows the audio
    amplitude of song Y at time T. Detection events on training songs
    are shown in cyan, with detections of unseen test songs in red.  To provide an intuition of intra-song variability, songs have been
    stably sorted by the time of detection events; thus, each of the
    three detection graphs shows the songs in a different order.}
  \label{fig:detection_raster}
\end{figure}

\subsection{Accuracy}

We allowed our training software to allocate our default of 4 hidden
units per syllable, and computed a new FFT every 1.5 ms.  Because our
timing test files are designed for our stereo playback software,
allowing only one channel for the ground-truth pulse, we trained one
detector for each syllable for the following results.  In order to
eliminate the large number of variables involved in microphone, cage,
playback and re-digitising, we evaluated the neural network's accuracy
directly on the digitised recording.  When training and runtime data
are gathered on the same hardware setup, this is the digital signal
that the detector will see---only the non-song data may be different.
%Accuracy \marginpar{Should be! Verify! And/or explain that there is a bug in the LabVIEW implementation, if there in fact is. ALSO need to present a picture of the test syllable or etc\dots} is identical across the three detectors.
\noprint{
  Typical confusion matrices for these syllables are:

  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_1$
    & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.96451\% & 0.000298\%\\ 
    neg & 0.0355\% & 99.99970\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_2$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.92903\% & 0.000596\%\\ 
    neg & 0.0710\% & 99.99940\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_3$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.68062\% & 0.00194\%\\ 
    neg & 0.319\% & 99.99806\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_4$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.82257\% & 0.00239\%\\ 
    neg & 0.177\% & 99.99761\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_5$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.78708\% & 0\\ 
    neg & 0.213\% & 100\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
  \begin{tabular}{r|cc}
    $t^*_6$ & \multicolumn{2}{c}{True} \\ 
    & pos & neg \\ 
    \hline  Detected pos & 99.71611\% & 0.00552\%\\ 
    neg & 0.284\% & 99.99448\%\\ 
  \end{tabular}
  \vspace{8pt}\par\noindent
}

Accuracies for our six test syllables are shown in
\fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
songs were always 100\%).  For each test syllable we trained
100 different
networks, which differ from each other in the random initialisation of
network weights and on which random subset of our songs was used for
training.  Each network's accuracy is shown as a single point in each
chart in \fig{fig:accuracies}.  Some syllables are easier to identify
reliably than others, allowing detection point choice to impact the
detector's accuracy.  For difficult syllables, the training process
occasionally yields a network that performs unusually poorly, but it
is easy to spot these networks by their performance either as measured
as part of the training process or on the test audio file, and
re-train when necessary.

\begin{figure}
  % Width 8, height 3, font 120%
  \includegraphics[width=\textwidth]{Fig3}
  %\includegraphics[width=\textwidth]{accuracies-2hidden}

  \caption{Accuracy variability over 100 different training runs for each
    of the six test syllables.  Each dot shows the test-set accuracy
    for an independently trained detector.  Because the horizontal
    positions have been randomised slightly so as not to occlude same-valued
    measurements, test syllable is also indicated by colour.}
  \label{fig:accuracies}
\end{figure}

We repeated the above experiment using only 2 hidden units per syllable.  Training under this condition is much faster, but the mean true positive rate per syllable decreases by an average of 18\% across our test syllables, and the mean false positive rate increases by roughly 23\%.

\subsection{Timing}

We evaluate variability in timing performance across three variables:
FFT frame interval; syllable choice; and detector implementation.

\subsubsection{FFT frame interval}

Detector latency and jitter depend on the FFT frame rate.  Our
1.5-ms--frame default is a compromise: shorter frames increase the
precision of timing, but also increase computational requirements both
during training and at runtime.  \fig{fig:TimingVsFrame} shows how
these numbers vary over a useful range of frame rates on our ideal
detector, the Swift detector with serial output, and for the LabVIEW
detector, for both the $\delta$-syllable and for the bird song at
$t^*_4$.

\begin{figure}
  \includegraphics[width=\textwidth]{Fig4}
  \caption{Timing varies as the FFT frame interval changes.  Here we
    show results for the ideal detector and the LabVIEW and
    Swift+serial implementations, for the constructed
    $\delta$-syllable and for trigger $t^*_4$ of the bird song.  The
    lines show latency; error bars are standard deviation (jitter).  Points have
    been shifted horizontally slightly for clarity; original positions are [0.5 1
      1.5 2 4] ms.}
  \label{fig:TimingVsFrame}
\end{figure}

\subsubsection{Syllable choice}

Syllable choice impacts detector performance, but despite the variety
of syllables chosen here, performance was fairly stable across
syllables.  \fig{fig:TimingVsSyllable} shows measured timing data for
the Swift+serial detector compared to the ideal, with
$t_{\textrm{fft}}=1.5$ ms.  Given the variety of the test syllables,
the variability is surprisingly low, and is summarised here with 95\%
confidence intervals (with n=6):
\vspace{8pt}\par\noindent
\begin{tabular}{r|cc}
  Detector & Latency & Jitter \\ 
  \hline   Ideal & $-0.8\pm 0.3$ ms & $2.1\pm 0.1$ ms \\
  Swift+serial & $0.0\pm 0.6$ ms & $2.0\pm 0.1$ ms
\end{tabular}
\vspace{8pt}\par\noindent
The negative latency is due to the way in which the network responds
to the song through time: as the recognition region looks increasingly
similar to the trained match, the network's evaluation of similarity
rises, and will generally cross the triggering threshold before it
reaches its maximum value.  A heuristic as simple as triggering at an
apparent turning point after crossing the threshold might improve
timing consistency at the expense of latency, but we did not test this.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig5}
  \end{center}
  \caption{Timing data for our 6 test syllables, for the ideal and the
    Swift+serial detectors, with an FFT frame rate of 1.5 ms.  Point
    centres show latency; error bars show jitter.}
  \label{fig:TimingVsSyllable}
\end{figure}

\subsubsection{Detector Implementations}

We compared latency and jitter across our different detector
implementations for the $\delta$-syllable and $t^*_4$, again with
$t_\mathrm{fft}=1.5$ ms.  Results are as follows:
\vspace{8pt}
\begin{tabular}{l|c|c|c| c}
  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird: $t^*_4$} \\
  Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
  \hline
  Ideal & 0.66 & 0.38 & -1.0 & 2.0 \\
  LabVIEW & 1.5 & 0.69 & -0.18 & 2.4 \\
  Swift+serial & 0.88 & 1.6 & -0.075 & 2.0 \\
  Swift+audio & 5.9 & 2.6 & 1.2 & 4.9 \\
  MATLAB+serial & 8.1 & 2.3 & 3.8 & 2.0 \\
  MATLAB+audio & 13 & 5.9 & 8.4 & 5.9
\end{tabular}
\vspace{8pt}\par\noindent
These data are also plotted in \fig{fig:TimingVsDetector}, and
\fig{fig:timing} gives a more detailed view of what the timing curves
look like for the five implementations of our detector on $t^*_4$.

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig6}
  \end{center}
  \caption{The different detectors for the constructed
    $\delta$-syllable and for the bird song at $t^*_4$.  Point centres
    show latency; error bars show jitter.}
  \label{fig:TimingVsDetector}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=8cm]{Fig7}
  \end{center}
  \caption{Raw timing curves for all detectors measured during
    detection of $t^*_4$ using 1.5-ms frames.  We extract the trigger events from each curve, from which we obtain the mean---latency---and
    standard deviation---jitter.}
  \label{fig:timing}
\end{figure}

\subsection{Fibre Photometry}
\label{sec:fp-results}

\fig{fig:fp-audio} shows 1,123 song performances from a single bird. 
At trial 858, the altered auditory feedback (AAF) was introduced, 
triggered by the first syllable in the song.

Measurements of the fluorescence from genetically encoded calcium indicators
in the HVC shelf reveal an AAF-sensitive response as compared with catch trials
(\fig{fig:fp-results}). This suggests a population of cells in the HVC
shelf that respond to AAF and may be involved in error signaling, given projections
from the shelf to AIV, an area area known to be critical for processing of 
sensory errors during singing \cite{MandelblatCerf:2014eu}.

The neural-network-based syllable detector enabled this experiment
by allowing low-latency AAF triggered off user-selected syllables, 
without hand-tuning and without relying on harmonic stacks needed 
for filter-bank-based syllable detection. The detector will allow
continued research into the neural encoding of these error signals.

\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{Fig8}
  \end{center}
  \caption{Audio amplitude stack plot of the experiment as aligned by
    the bird's song.  The top panel shows a spectrogram of the song
    with the triggered white-noise stimulus.  In the bottom panel,
    each row is average amplitude for a single performance of the song
    (1123 songs total).  The white-noise stimulus is visible as a
    bright yellow region after the first syllable in the final quarter
    of the trials.}
  \label{fig:fp-audio}
\end{figure}

\begin{figure}
  \begin{center}
    \includegraphics[width=7cm]{Fig9}
  \end{center}
  \caption{Fibre photometry during AAF reveals an AAF-sensitive response
  (lines indicate mean, shading indicates 95\% confidence interval, bootstrap).
  Blue are catch trials, red are AAF trials.}
  \label{fig:fp-results}
\end{figure}

\section{Discussion}
\label{sec:conclusion}

This syllable detector is appropriate for zebra finch song, and
although our tests were carried out on songs from that species, it is
also likely to work well for Bengalese finches.  It offers the
following benefits:
\begin{itemize}
\item The detector is accurate. False negative and false positive
  rates can be well under 0.5\% and 0.01\% respectively, and trading
  these two numbers off against each other is through a single
  relative-cost parameter.
\item Latency is generally under a millisecond, with jitter around 2
  ms.
\item Works on a wide range of target syllables using the default
  values described here, generally eliminating the need for
  hand-tuning.
\item Runs fast enough for syllable-modification experiments on
  inexpensive consumer-grade hardware, although we recommend that the
  training phase be run on a fast desktop system with 32 GB of RAM.
\item A single detector can generate different target pulses for
  multiple syllables at almost no additional computational cost during
  runtime, although training time will increase.
\end{itemize}

Although there are differences, the Swift+serial detector and the
LabVIEW implementation are roughly comparable in performance.  We
prefer the Swift implementation due to its lower hardware and software
requirements and the difficulty of debugging LabVIEW programmes.  With
serial output, MATLAB's performance is good, although its buffer
handling is sensitive to system load.

The song presented here was recorded with a fixed microphone mounted
inside the cage.  We found that higher accuracy is achieved when a
microphone is instead mounted on the bird's head, which maintains
volume and reduces changes in timbre as the bird moves.
% We have head mounted microphone data?

A common experimental paradigm requires detecting the frequency of
syllables.  Many pitch detection techniques rely on the spectrum,
which incurs no additional computational cost here since it is already
available.  For example, \cite{Canopoli2014} achieved good results
with the Harmonic Product Spectrum algorithm
\cite{Noll1970pitchdetection}.

In order to monitor syllable duration, the beginning and end of a
syllable may be detected by looking at the ratio of total energy in
the singing frequency band to the total energy, over some small time
window.  Any syllable thus identified that also contains a trigger
event may be monitored for duration.  Alternatively, the network can
be trained to recognise both the beginning and the end of the syllable
of interest.

In feedback experiments such as frequency- or duration-shifting, vocal
output changes over the course of the experiment.  The neural network
excels at identifying syllables close to its training set, so as vocal
output changes the detector may not recognise a match.  If the
detector must be robust to this shift, it may be retrained as often as
necessary as the bird learns, or data consisting of synthetically
pitch-shifted or duration-shifted target syllables over the
recognition region may be added to the training set.  We will test
these approaches in future work.

\section{Acknowledgments}
We would like to thank Winthrop F.~Gillis, William A.~Liberti, and Sanne Moorman for
their improvements to the draft.  This work was funded by NIH grants
5R01NS089679-02 and 5U01NS090454-02.  The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

\appendix

\section{Resources}
\label{sec:resources}

{\em If accepted, before final submission, all code will be moved to the Gardner Lab github repository, final release versions generated, and DOIs created as described in https://guides.github.com/activities/citable-code/ so that our results may be reproduced and the software deployed.  For now, in lieu of final release DOIs, we provide these links to our development repositories.  If you would like final versions and test datasets for review, please contact the corresponding author (bwpearre@bu.edu).}
%\marginpar{PLOS doesn't really say where such things should go.}
%\marginpar{Cleanup: mostly done.  But it would be nice to move all this to the lab's GitHub group.}
\begin{description}
  \item[Song alignment:] Jeff Markowitz's song alignment software \cite{Poole2012} can be found at
    {\tt https://github.com/jmarkow/zftftb}
\end{description}

 \noindent Our implementation of the syllable
    detector is available under an open-source license {\em (we are still evaluating which open source license is the most suitable for this software):}
    \begin{description}
    \item[Training the neural network:]\hfill
      \begin{trivlist}
      \item {\tt https://github.com/bwpearre/birds/align}
      \end{trivlist}

    \item[Runtime:]\hfill
      \begin{description}
      \item[MATLAB:] {\tt https://github.com/gardner-lab/syllable-detector-matlab}
      \item[Swift:] {\tt https://github.com/gardner-lab/syllable-detector-swift}
      \item[LabVIEW:] {\tt https://github.com/bwpearre/birds/align}
      \end{description}

\end{description}

    %\bibliography{birds}
\begin{thebibliography}{10}

\bibitem{Kao2005}
Kao MH, Doupe AJ, Brainard MS.
\newblock Contributions of an avian basal ganglia--forebrain circuit to
  real-time modulation of song.
\newblock Letters to Nature. 2005 February;433:638--643.
\newblock Available from:
  \url{http://www.nature.com/nature/journal/v433/n7026/abs/nature03127.html}.

\bibitem{Wang2008interhemispheric}
Wang CZH, Herbst JA, Keller GB, Hahnloser RHR.
\newblock Rapid Interhemispheric Switching during Vocal Production in a
  Songbird.
\newblock PLOS Biology. 2008 October;6(10).
\newblock Available from:
  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.0060250}.

\bibitem{Keller2009}
Keller GB, Hahnloser RHR.
\newblock Neural processing of auditory feedback during vocal practice in a
  songbird.
\newblock Nature. 2009 January;457:187--190.
\newblock Available from:
  \url{http://www.nature.com/nature/journal/v457/n7226/abs/nature07467.html}.

\bibitem{Tumer2007pitchshift}
Tumer EC, Brainard MS.
\newblock Performance variability enables adaptive plasticity of 'crystallized'
  adult birdsong.
\newblock Nature. 2007 December;450:1240--1244.
\newblock Available from:
  \url{http://www.nature.com/nature/journal/v450/n7173/abs/nature06390.html}.

\bibitem{Andalman2009}
Andalman AS, Fee MS.
\newblock A basal ganglia-forebrain circuit in the songbird biases motor output
  to avoid vocal errors.
\newblock Proceedings of the National Academy of Sciences of the United States
  of America. 2009 July;106(30):12518--12523.
\newblock Available from:
  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2709669/}.

\bibitem{Warren2011}
Warren TL, Tumer EC, Charlesworth JD, Brainard MS.
\newblock Mechanisms and time course of vocal learning and consolidation in the
  adult songbird.
\newblock Journal of Neurophysiology. 2011 October;106(4):1806--1821.
\newblock Available from:
  \url{http://jn.physiology.org/content/106/4/1806.full}.

\bibitem{Canopoli2014}
Canopoli A, Herbst JA, Hahnloser RHR.
\newblock A Higher Sensory Brain Region Is Involved in Reversing
  Reinforcement-Induced Vocal Changes in a Songbird.
\newblock The Journal of Neuroscience. 2014;34(20):7018--7026.
\newblock Available from:
  \url{http://www.jneurosci.org/content/34/20/7018.full}.

\bibitem{Leonardo1999}
Leonardo A, Konishi M.
\newblock Decrystallization of adult birdsong by perturbation of auditory
  feedback.
\newblock Nature. 1999 June;399:466--470.
\newblock Available from:
  \url{http://www.nature.com/nature/journal/v399/n6735/abs/399466a0.html}.

\bibitem{Skocik2013}
Skocik M, Kozhevnikov A.
\newblock Real-time system for studies of the effects of acoustic feedback on
  animal vocalizations.
\newblock Frontiers in Neural Circuits. 2013 January;6(111).
\newblock Available from:
  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3539774/}.

\bibitem{Poole2012}
Poole B, Markowitz JE, Gardner TJ.
\newblock The Song Must Go On: Resilience of the Songbird Vocal Motor Pathway.
\newblock PLOS One. 2012 June;7(6).
\newblock Available from:
  \url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0038173}.

\bibitem{Jacobsen2003SlidingDFT}
Jacobsen E, Lyons R.
\newblock The sliding DFT.
\newblock IEEE Signal Processing Magazine. 2003 Mar;20(2):74--80.
\newblock Available from:
  \url{http://www.cmlab.csie.ntu.edu.tw/DSPCourse/reference/Sliding\%20DFT.pdf}.

\bibitem{hkp}
Hertz JA, Krogh AS, Palmer RG.
\newblock Introduction to the Theory of Neural Computation.
\newblock Perseus Books; 1991.

\bibitem{Markowitz:2015ko}
Markowitz JE, Liberti WA, Guitchounts G, Velho T, Lois C, Gardner TJ.
\newblock Mesoscopic Patterns of Neural Activity Support Songbird Cortical
  Sequences.
\newblock PLOS Biology. 2015 June;13(6):e1002158--20.
\newblock Available from:
  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1002158}.

\bibitem{Adelsberger:2005dy}
Adelsberger H, Garaschuk O, Konnerth A.
\newblock {Cortical calcium waves in resting newborn mice.}
\newblock Nature Neuroscience. 2005 August;8(8):988--990.
\newblock Available from:
  \url{http://www.nature.com/neuro/journal/v8/n8/full/nn1502.html}.

\bibitem{Schulz:2012fv}
Schulz K, Sydekum E, Krueppel R, Engelbrecht CJ, Schlegel F, Schr{\"o}ter A,
  et~al.
\newblock {Simultaneous BOLD fMRI and fiber-optic calcium recording in rat
  neocortex}.
\newblock Nature Methods. 2012 May;9(6):597--602.
\newblock Available from:
  \url{http://www.nature.com/nmeth/journal/v9/n6/full/nmeth.2013.html}.

\bibitem{Cui:2013dq}
Cui G, Jun SB, Jin X, Pham MD, Vogel SS, Lovinger DM, et~al.
\newblock {Concurrent activation of striatal direct and indirect pathways
  during action initiation}.
\newblock Nature. 2013 February;494(7436):238--242.
\newblock Available from:
  \url{http://www.nature.com/nature/journal/v494/n7436/full/nature11846.html}.

\bibitem{Adelsberger:2014jd}
Adelsberger H, Zainos A, Alvarez M, Romo R, Konnerth A.
\newblock {Local domains of motor cortical activity revealed by fiber-optic
  calcium recordings in behaving nonhuman primates.}
\newblock Proceedings of the National Academy of Sciences of the United States
  of America. 2014 January;111(1):463--468.
\newblock Available from: \url{http://www.pnas.org/content/111/1/463.full}.

\bibitem{Gunaydin:2014dh}
Gunaydin LA, Grosenick L, Finkelstein JC, Kauvar IV, Fenno LE, Adhikari A,
  et~al.
\newblock {Natural neural projection dynamics underlying social behavior.}
\newblock Cell. 2014 June;157(7):1535--1551.
\newblock Available from:
  \url{http://www.cell.com/cell/fulltext/S0092-8674%2814%2900659-X}.

\bibitem{MandelblatCerf:2014eu}
Mandelblat-Cerf Y, Las L, Denissenko N, Fee M.
\newblock {A role for descending auditory cortical projections in songbird
  vocal learning}.
\newblock Elife. 2014 January;3(0):e02152--e02152.
\newblock Available from: \url{https://elifesciences.org/content/3/e02152}.

\bibitem{Noll1970pitchdetection}
Noll AM.
\newblock Pitch determination of human speech by the harmonic product spectrum,
  the harmonic sum spectrum, and a maximum likelihood estimate.
\newblock In: Proceedings of the symposium on computer processing in
  communications. vol.~19; 1970. p. 779--797.

\end{thebibliography}

\end{document}

### 2016-07-27 main.tex
Begin to incorporate reviewer feedback:
- Most of the tutorial

- Reference other birds, explain why it's ok that we just show 1

- specify "Undirected song"

- LevenbUrg

- Remove fibre photometry from methods (need to remove from results; this is just an intermediate checkin so Nathan can add his bit)

@@
 
-We begin with a few hundred recordings of a given bird's song, in this
-case made inside a sound-isolating chamber in which was mounted a
-recording microphone (Audio-Technica AT803).  The songs are
-time-aligned as described in \cite{Poole2012}.
+We begin with a few hundred recordings of a given bird's song.  Male
+zebra finch song is most highly stereotyped when a female is present
+(``directed song'') and slightly more variable otherwise
+(``undirected''); we train and test on undirected song since this is
+more challenging.  The songs are recorded inside a sound-isolating
+chamber in which was mounted a recording microphone (Audio-Technica
+AT803).  The songs are time-aligned as described in \cite{Poole2012}.
 
@@
 
-\noprint{
-  % I screwed up: the results are without cage noise.  Rather than
-  repeat all measurements, let's just not mention it.  }
-
 With inputs well outside the space on which a neural network has been
 trained, its outputs will be essentially random. In order to reduce
-the false positive rate it is useful to provide negative training
-examples that include silence, cage noise, non-song vocalisations, and
-perhaps songs from other birds.  Although it will depend on the makeup
-of the non-song data, we have found that training with as low as a 1:1
-ratio of non-song to song yields excellent results.
+the false positive rate it is necessary to provide negative training
+examples that include silence, cage noise, wing flapping, non-song
+vocalisations, and perhaps songs from other birds.  Although it will
+depend on the makeup of the non-song data, we have found that training
+with as low as a 1:1 ratio of non-song to song yields excellent
+results on most birds.
 
@@
 
-We used the Levenburg-Marquardt algorithm, which is the MATLAB
+We used the Levenberg-Marquardt algorithm, which is the MATLAB
 toolbox's default.  It is a fast algorithm, but is memory-intensive,
@@
 
-\subsection{Fibre Photometry}
-\label{sec:fp-methods}
-
-In addition to synthetic tests of detector accuracy and 
-latency, we piloted the syllable detector in an experiment
-providing altered auditory feedback (AAF) to zebra finches. AAF, or 
-playback of a perturbing white noise stimulus, is known to disrupt 
-the highly-stereotyped zebra finch song. By probabilistically 
-targeting the AAF to a specific syllable, it is possible 
-to selectively destabilize parts of the song. During this 
-feedback, signal from genetically-encoded calcium indicators was 
-recorded via fibre photometry to identify a neural representation of 
-an error signals associated with the feedback.
-
-The initial experiment design used a DSP with hand-designed filter-banks to
-detect syllables, but these required tuning and were only conducive
-to detecting harmonic stacks. This experiment provided an ideal design
-to pilot our neural-network-based syllable detector.
-
-\subsubsection{Surgery}
-
-GCaMP6s and GCaMP6f were expressed using a lentivirus under the RSV 
-promoter \cite{Markowitz:2015ko}. Injections were targeted under HVC 
-using stimulation of Area X with a bipolar stimulation electrode. 
-After injecting the virus, we implanted a fibre-optic cannula above 
-HVC ($400 \mu m$ diameter, Doric Lenses). The location of the injection 
-was verified histologically after each experiment.
-
-\subsubsection{Photometry}
-
-Fibre photometry \cite{Adelsberger:2005dy,Schulz:2012fv,Cui:2013dq,Adelsberger:2014jd,Gunaydin:2014dh} 
-was performed in N=3 birds at least 2 weeks post-injection using a 
-completely fibre-coupled light path. A 470 nm fibre-coupled LED (Mightex) 
-was sinusoidally driven at 5 kHz by a TDT RX8. The output of the LED 
-was filtered with a GFP excitation filter (Doric Lenses Fluorescence Cube) 
-and coupled to an optical rotary joint (Doric Lenses). The rotary joint 
-was connected via a multimode fibre optic patch cord (400 $\mu$m, .48 NA, 
-Doric Lenses) to the implanted cannula using a brass sleeve (Doric Lenses). 
-LED power was calibrated such that optical power at the tip was 
-20--300 $\mu$W (LED was turned on only during vocalizations, detected using 
-a TDT RX8). Fluorescence was collected through the same patch cord, 
-filtered with a GFP emission filter, and sent to a silicon photomultiplier 
-(SensL MiniSM 30000). The output of the photomultiplier was demodulated 
-using a dual-phase lock-in amplifier (10-ms time-scale, SR830, Stanford 
-Research Systems) and digitized along with the bird's song at 24.414 kHz 
-using a National Instruments data acquisition board (PCIE-6323).
-
-\subsubsection{Syllable detection}
-
-User-selected syllables were detected using the Swift implementation of the
-syllable detector, and the TTL output triggered altered auditory feedback 
-via the TDT RX8.
-
-\subsubsection{Data analysis}
-
-$\Delta F / F0$ was estimated at each point by subtracting and dividing by 
-F0, defined as the 11th percentile computed in a 300-ms sliding window. 
-Data were then smoothed with a 50-ms exponential window. Fluorescence traces
-during singing were normalized by subtracting the trial-average signal, in 
-order to remove song-based modulation.
-
 \subsection{Ethics}
@@
 \label{sec:results}
+
+We have used this detector for song alignment and feedback experiments
+in 5 behaving birds, and for offline analysis on various others.  For
+simplicity we present performance on 6 randomly chosen points in one
+bird's song, but the results presented here are representative of
+real-world performance.\marginpar{Ok?}
 
@@
 
+  \begin{description}
+  \item[Installation:]\hfill
+    \begin{trivlist}
+    \item Clone the Github repository onto your computer.
+    \end{trivlist}
+  \item[Usage:]\hfill
+    \begin{itemize}
+      \item The main programme is {\tt learn\_detector.m}.
+    \item The top section of {\tt learn\_detector.m} contains global configuration variables.
+The important ones are the data directory and the bird name: as supplied, the software looks for {\tt \textless data directory\textgreater/\textless bird name\textgreater/song.mat}.
+Modify as necessary.
+    \item The format for the data file {\tt \textless data directory\textgreater/\textless bird name\textgreater/song.mat} is specified in {\tt birds/align/README.md}: a MATLAB file containing aligned song, nonsong (calls, cage noise, etc), and the sampling frequency.
+    \item The configuration file {\tt \textless data directory\textgreater/\textless bird name\textgreater/params.m} is also described in {\tt README.md}.
+If the file is empty, the programme will display an image showing the average of the song spectrograms (frequency vs.~time), tell the user how to define syllables of interest, and exit.
+If {\tt times\_of\_interest} is defined, the programme will train the neural network.
+{\tt params.m} can also override any of the configuration parameters at the top of {\tt learn\_detector.m}.
+    \item The software will produce a neural network file, {\tt detector\_bird\_times\_\dots.mat} (the ellipses signify other information encoded in the filename).
+It will also produce the audio test file described in Methods.
+    \end{itemize}
+  \end{description}
+
 \item[Runtime:]\hfill
@@
   \item[MATLAB:] {\tt https://github.com/gardner-lab/syllable-detector-matlab}
+    \marginpar{Jeff: Can you flesh this out?}
+    \begin{description}
+    \item[Installation:]\hfill
+      \begin{trivlist}
+      \item Clone the Github repository onto your computer.
+      \end{trivlist}
+    \item[Connection:]\hfill
+    \item[Usage:]\hfill
+    \end{description}
   \item[Swift:] {\tt https://github.com/gardner-lab/syllable-detector-swift}
+    \marginpar{Nathan: Can you flesh this out?}
+    \begin{description}
+    \item[Installation:]\hfill
+      \begin{trivlist}
+      \item Clone the Github repository onto your computer.
+      \end{trivlist}
+    \item[Connection:]\hfill
+      \begin{itemize}
+      \item Connect the microphone through a preamp to {\em FIXME the left channel of the stereo audio input jack on the Mac Mini?}.
+      \item {\em FIXME} How to connect the Arduino?  Add the Arduino programme to your GitHub?
+      \item {\em FIXME} How to connect if not using Arduino?
+      \end{itemize}
+    \item[Usage:]\hfill
+      \begin{itemize}
+      \item Use {\tt convert\_to\_text.m} to convert the Matlab output file to a format easily readable by Swift.
+      \item {\em FIXME} How to start Swift?  Parameters?
+      \end{itemize}
+    \end{description}
   \item[LabVIEW:] {\tt https://github.com/bwpearre/birds/align}
+    \begin{description}
+    \item[Installation:]\hfill
+      \begin{trivlist}
+      \item Clone the Github repository onto your computer.
+      \end{trivlist}
+    \item[Connection:]\hfill
+      \begin{itemize}
+      \item The microphone should be connected through a preamplifier to the data acquisition card, by default on Analog Input 0 (AI0).
+      \item Outputs are on the Digital I/O (DIO) channels.  See the block diagram of {\tt nnfft.vi} for documentation, or just connect to DIO0.
+      \end{itemize}
+    \item[Usage:]\hfill
+      Use LabVIEW to open {\tt nnfft.vi}.  A file
+      chooser is provided in the front panel.  Use it to point to the network file
+      produced by {\tt learn\_detector.m}, and use LabView's ``Run''
+      button to begin detection.
+    \end{description}
   \end{description}

### 2016-07-27 main.tex
Most of the rest of reviewer feedback:
- Clarify inclusion of non-song data in training/test

- Remove apparently-confusing (?!) reference to hillclimbing

- Clarify how we measure latency, jitter, and give example of how to trigger at the very beginning of a syllable?

- "Here's a presentation of one bird; we have run others with similar results."  Ok?

- Quick reference to why this might work with Bengalese finches (correct me?)

- Remove the rest of Jeff's experiment.

Still waiting on new accuracy data, and confirmation that Levenberg-Marquardt is obsolete.

@@
 with as low as a 1:1 ratio of non-song to song yields excellent
-results on most birds.
+results on most birds.  These data are also included in our accuracy tests.
 
@@
 Section~\ref{sec:accuracy}. Since large portions of the cost function
-are flat, hillclimbing is ineffective.  Random-restart hillclimbing
-would work, but a brute-force search requires fractions of a
-second. For the results presented here, we have used $C_n=1$ and
-$\Delta t=10$ ms.
+are flat, we use a brute-force linear search, which requires fractions of a
+second. For the results presented here, we have used
+$\Delta t=10$ ms, and arbitrarily set $C_n=1$.
 
@@
 definition of false-positive rate for data streams that had not been
-segmented into songs.
+segmented into song-sized chunks and thus allowing easy interpretation
+of false-positive rates from calls, cage noise, etc.
 
@@
 
-We evaluate the time taken from the presentation of the target
-syllable to the firing of the detector's TTL pulse. While playing the
-audio test file from any audio playback device, the TTL output from
-the ground-truth channel of the audio output may be used as the
-trigger pulse for an oscilloscope, and compared to the TTL pulse
-produced by the detector implementation, which sees only the birdsong
+We evaluate the time taken from the presentation of each instance of the target
+syllable to the firing of the detector's TTL pulse---for example, if the target trigger point with respect to the aligned songs is at 200 ms, and for a given presentation of the song the detector fires at 203 ms, we take the latency on that trial to be 3 ms.
+
+While playing the
+audio test file from any audio playback device, we used the TTL output from
+the ground-truth channel of the audio output as the
+trigger pulse for an oscilloscope, and compared it to the TTL pulse
+produced by the detector, which sees only the birdsong
 channel of the audio file. For this purpose we used a pulse generator
@@
 
-% Use to say: "Mean latency is then given as the halfway point of that detection 
-% sigmoid."
-%
-% I don't think this description matches the calculation, this description
-% seems to better match a "median" latency. The mean latency is calculated
-% as an average latency of all detection events, which is the same as 
-% fitting a normal distribution to the detection events and taking the mean
-% of the normal distribution (the MLE estimator is the same as the sample mean).
-
 Mean latency is then given as the average time between the 
 canonical signal and the corresponding detection event. 
-It is a helpful number, but not a critical one, since a
+It is a helpful number, but not a critical one, since in zebra finch a
 detector with high but constant latency can often be trained to
-trigger at a point somewhat before the true moment of interest,
-especially given the stereotypy of zebra finch song.  Often more
-important is latency jitter: how much variability is there in the
+trigger at a point somewhat before the true moment of interest (for example, if triggering should occur at the very beginning of a syllable, the detector's time window may include previous syllable).
+Often more
+important is the variability in the
 latency?  We define jitter as the standard deviation of the latency.
@@
 bird's song, but the results presented here are representative of
-real-world performance.\marginpar{Ok?}
+real-world performance in our lab.\marginpar{Ok?}
 
@@
 
-\subsection{Fibre Photometry}
-\label{sec:fp-results}
-
-\fig{fig:fp-audio} shows 1,123 song performances from a single bird. 
-At trial 858, the altered auditory feedback (AAF) was introduced, 
-triggered by the first syllable in the song.
-
-Measurements of the fluorescence from genetically encoded calcium indicators
-in the HVC shelf reveal an AAF-sensitive response as compared with catch trials
-(\fig{fig:fp-results}). This suggests a population of cells in the HVC
-shelf that respond to AAF and may be involved in error signaling, given projections
-from the shelf to AIV, an area area known to be critical for processing of 
-sensory errors during singing \cite{MandelblatCerf:2014eu}.
-
-The neural-network-based syllable detector enabled this experiment
-by allowing low-latency AAF triggered off user-selected syllables, 
-without hand-tuning and without relying on harmonic stacks needed 
-for filter-bank-based syllable detection. The detector will allow
-continued research into the neural encoding of these error signals.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=7cm]{Fig8}
-  \end{center}
-  \caption{Audio amplitude stack plot of the experiment as aligned by
-    the bird's song.  The top panel shows a spectrogram of the song
-    with the triggered white-noise stimulus.  In the bottom panel,
-    each row is average amplitude for a single performance of the song
-    (1123 songs total).  The white-noise stimulus is visible as a
-    bright yellow region after the first syllable in the final quarter
-    of the trials.}
-  \label{fig:fp-audio}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=7cm]{Fig9}
-  \end{center}
-  \caption{Fibre photometry during AAF reveals an AAF-sensitive response
-  (lines indicate mean, shading indicates 95\% confidence interval, bootstrap).
-  Blue are catch trials, red are AAF trials.}
-  \label{fig:fp-results}
-\end{figure}
-
 \section{Discussion}
@@
 although our tests were carried out on songs from that species, it is
-also likely to work well for Bengalese finches.  It offers the
+also likely to work well for Bengalese finches due to their similarly stereotyped syllables.  It offers the
 following benefits:

### 2016-07-28 main.tex
Nothing much.  Interim checkin.

@@
 
-We begin with a few hundred recordings of a given bird's song.  Male
+We begin with a few hundred recordings of a given bird's song, as well as calls, cage noise, and other non-song audio data.  Male
 zebra finch song is most highly stereotyped when a female is present
@@
 (``undirected''); we train and test on undirected song since this is
-more challenging.  The songs are recorded inside a sound-isolating
+more challenging.  Recordings were made inside a sound-isolating
 chamber in which was mounted a recording microphone (Audio-Technica
-AT803).  The songs are time-aligned as described in \cite{Poole2012}.
+AT803).  The songs were time-aligned as described in \cite{Poole2012}.
 

### 2016-07-28 main.tex

@@
   \item[Swift:] {\tt https://github.com/gardner-lab/syllable-detector-swift}
-    \marginpar{Nathan: Can you flesh this out?}
     \begin{description}
@@
       \begin{trivlist}
-      \item Clone the Github repository onto your computer.
+      \item Download the most recent release from the Github repository onto your computer.
       \end{trivlist}
@@
       \begin{itemize}
-      \item Connect the microphone through a preamp to {\em FIXME the left channel of the stereo audio input jack on the Mac Mini?}.
-      \item {\em FIXME} How to connect the Arduino?  Add the Arduino programme to your GitHub?
-      \item {\em FIXME} How to connect if not using Arduino?
+      \item Either connect the microphone directly into the microphone jack or through a 
+      preamp into the audio input jack. If the input jack supports multiple channels (e.g., 
+      audio input jacks are usually stereo), you can connect multiple audio sources and run 
+      different detectors on each channel simultaneously.
+      \item To generate TTL signals via the audio output, use the headphone jack. The 
+      audio output must have at least the same number of channels as the input.
+      \item To generate TTL signals via an Arduino, load the MATLAB arduino IO sketch 
+      onto the Arduino (the sketch is available in the Github repository, and provides a 
+      simple serial protocol for controlling pins). Connect the Arduino 
+      to the computer via USB. The TTL signal for the first input will be available on 
+      pin 7, the second input on pin 8, etc.
       \end{itemize}
@@
       \begin{itemize}
-      \item Use {\tt convert\_to\_text.m} to convert the Matlab output file to a format easily readable by Swift.
-      \item {\em FIXME} How to start Swift?  Parameters?
+      \item Use {\tt convert\_to\_text.m} to convert the Matlab output file to a format 
+      easily readable by Swift. This will generate a text version of the neural network.
+      \item Launch the application downloaded above. It will provide an 
+      menu to select an audio input and either an audio output or an Arduino output (if 
+      detected). Once you select an input and an output, a new window will show all audio 
+      channels associated with the input. For each channel, select a the text-encoded version of 
+      the detector created in the previous step. Once configured, press run to begin 
+      monitoring the inputs.
+      \end{itemize}
+    \item[Customization:]\hfill
+      \begin{itemize}
+      \item You can clone the Github repository and use Xcode to edit the project, if you 
+      are interested in extending or modifying the functionality. To hook into either 
+      the raw input or output, look at the ``ViewControllerProcessor.swift'' file. The 
+      {\tt Processor} class acts as a coordinator, receiving income audio, passing that
+      to the detector and generating appropriate output TTL signals.
       \end{itemize}

### 2016-07-29 main.tex
Misc. changes.  Interim checkin.

--- 2016-07-28 main.tex
@@
 
-The network is trained using MATLAB's neural network toolbox. We tried
+The network is trained using MATLAB's neural network toolbox, with Scaled Conjugate Gradient ({\tt trainscg}). We tried
 a variety of feedforward neural network geometries, from simple
@@
 zebra finch song.
-
-We used the Levenberg-Marquardt algorithm, which is the MATLAB
-toolbox's default.  It is a fast algorithm, but is memory-intensive,
-so many target syllables or high FFT frame rates require a large
-amount of RAM and increase training time to hours. Other training
-algorithms that use less RAM are much slower, and by default they will
-often terminate prematurely due to their performance gradient going to
-0.
 
@@
 reported in the following section, the training set consisted of 1000
-instances of the song of one bird.
-Our test set consisted of an additional 1818 iterations of the bird's
-song.
+instances of the song of one bird, and 2000 song-length samples of
+non-song (cage noise, calls, other birds' song).  Our test set
+consisted of an additional 1818 iterations of the bird's song and 3636
+song-length samples of non-song.
 
 The MATLAB neural network toolbox further divides our ``training'' set
-into internal training, validation, and test sets.  We did not stray
-from the toolbox's default values, and do not discuss MATLAB's
-internal subdivision of our training set.  Because the dataset
-consists of temporally aligned songs, ground truth is available for
-each song (albeit with minor variations due to the alignment process
-used \cite{Poole2012}), and thus the detector can be checked by
-presenting the recorded training songs as well as the canonical
-detection events. To this end, besides the trained network object, our
-learning code produces an audio file consisting of all of the training
-data on the left audio channel and a delta function at each aligned
-moment of target syllable presentation on the right channel. Thus,
-when played on any audio player, the left channel may be provided as
-input to the detector, and the the detector's output pulses may be
-compared against the right channel.
+into internal training, validation, and test sets.  Because we do our
+own testing on a distinct and much larger data set, we instructed the
+toolbox to use 80\% of its training set for training, and the
+remaining 20\% for validation, with no hold-out test data.
+
+Because the data set consists of temporally aligned songs, ground
+truth is available for each song (albeit with minor variations due to
+the alignment process used \cite{Poole2012}), and thus the detector
+can be checked by presenting the recorded training songs as well as
+the canonical detection events. To this end, besides the trained
+network object, our learning code produces an audio file consisting of
+all of the training data on the left audio channel and a delta
+function at each aligned moment of target syllable presentation on the
+right channel. Thus, when played on any audio player, the left channel
+may be provided as input to the detector, and the the detector's
+output pulses may be compared against the right channel.
 
@@
 
-We evaluate the time taken from the presentation of each instance of the target
-syllable to the firing of the detector's TTL pulse---for example, if the target trigger point with respect to the aligned songs is at 200 ms, and for a given presentation of the song the detector fires at 203 ms, we take the latency on that trial to be 3 ms.
-
-While playing the
-audio test file from any audio playback device, we used the TTL output from
-the ground-truth channel of the audio output as the
-trigger pulse for an oscilloscope, and compared it to the TTL pulse
-produced by the detector, which sees only the birdsong
-channel of the audio file. For this purpose we used a pulse generator
-(Philips PM 5715, with a listed latency of 50 ns, jitter of $\leq
-0.1\%$ or 50 ps, whichever is greater) to widen the detector's output
-spike to a number much larger than the jitter ($\sim 100$ ms).  This
-obviates pulse length variability in the output device by essentially
+We evaluate the time taken from the presentation of each instance of
+the target syllable to the firing of the detector's TTL pulse---for
+example, if the target trigger point with respect to the aligned songs
+is at 200 ms, and for a given presentation of the song the detector
+fires at 203 ms, we take the latency on that trial to be 3 ms.
+
+While playing the audio test file from any audio playback device, we
+used the TTL output from the ground-truth channel of the audio output
+as the trigger pulse for an oscilloscope, and compared it to the TTL
+pulse produced by the detector, which sees only the birdsong channel
+of the audio file. For this purpose we used a pulse generator (Philips
+PM 5715, with a listed latency of 50 ns, jitter of $\leq 0.1\%$ or 50
+ps, whichever is greater) to widen the detector's output spike to a
+number much larger than the jitter ($\sim 100$ ms).  This obviates
+pulse length variability in the output device by essentially
 discarding the falling edge of the output pulse.  The oscilloscope is
@@
 probability distribution function (CDF) of the detector's output in
-response to the chosen song event, with individual detection events visible as steps in this curve.
-
-Mean latency is then given as the average time between the 
-canonical signal and the corresponding detection event. 
-It is a helpful number, but not a critical one, since in zebra finch a
-detector with high but constant latency can often be trained to
-trigger at a point somewhat before the true moment of interest (for example, if triggering should occur at the very beginning of a syllable, the detector's time window may include previous syllable).
-Often more
-important is the variability in the
-latency?  We define jitter as the standard deviation of the latency.
+response to the chosen song event, with individual detection events
+visible as steps in this curve.
+
+Mean latency is then given as the average time between the canonical
+signal and the corresponding detection event.  It is a helpful number,
+but not a critical one, since in zebra finch a detector with high but
+constant latency can often be trained to trigger at a point somewhat
+before the true moment of interest (for example, if triggering should
+occur at the very beginning of a syllable, the detector's time window
+may include previous syllable).  Often more important is the
+variability in the latency?  We define jitter as the standard
+deviation of the latency.
 
@@
 
-All procedures were approved by the Institutional Animal Care and Use Committee of Boston University (protocol number 14-029).
+All procedures were approved by the Institutional Animal Care and Use
+Committee of Boston University (protocol number 14-029).
 
@@
 We have used this detector for song alignment and feedback experiments
-in 5 behaving birds, and for offline analysis on various others.  For
+in 5 behaving birds, and for offline analysis on a few others.  For
 simplicity we present performance on 6 randomly chosen points in one
@@
 
-\fig{fig:song} shows the song for which we present accuracy and timing results.  Six
-example target trigger points are shown, at 150, 200, 250, 300, 350,
-and 400 milliseconds after the beginning of the aligned samples, as
-indicated by the red lines, and referred to henceforth as $t^*_1\ldots
-t^*_6$ respectively.  The rectangles show the recognition window (30
-ms, 1--8 kHz) for each instant in the song.
+\fig{fig:song} shows the song for which we present accuracy and timing
+results.  Six example target trigger points are shown, at 150, 200,
+250, 300, 350, and 400 milliseconds after the beginning of the aligned
+samples, as indicated by the red lines, and referred to henceforth as
+$t^*_1\ldots t^*_6$ respectively.  The rectangles show the recognition
+window (30 ms, 1--8 kHz) for each instant in the song.
 
@@
 far from pure tones, but are closer to coloured noise.  They would be
-difficult for a harmonic-stack filter bank to detect, and especially to
-pinpoint in time.  $t^*_3$ (250 ms) and $t^*_6$ (400 ms) are rich in
-overtones and contain downward slides with changing timbres.  $t^*_4$
-(300 ms) occurs near the beginning of a more typical harmonic stack amenable to a variety of
-recognition techniques, although consistently detecting a point
-partway through the syllable demands a detector that can use multiple
-time steps.  $t^*_5$ (350 ms) shows a steady pitch followed by a
-complex changing tone structure.
+difficult for a harmonic-stack filter bank to detect, and especially
+to pinpoint in time.  $t^*_3$ (250 ms) and $t^*_6$ (400 ms) are rich
+in overtones and contain downward slides with changing timbres.
+$t^*_4$ (300 ms) occurs near the beginning of a more typical harmonic
+stack amenable to a variety of recognition techniques, although
+consistently detecting a point partway through the syllable demands a
+detector that can use multiple time steps.  $t^*_5$ (350 ms) shows a
+steady pitch followed by a complex changing tone structure.
 
 An example of the detector's output for three of these syllables is
-shown in \fig{fig:detection_raster}.  The total audio energy of each song is shown as a single grayscale row.  By sorting according to time of
-each syllable's detection, the rest of the song is shown in the
+shown in \fig{fig:detection_raster}.  The total audio energy of each
+song is shown as a single grayscale row.  By sorting according to time
+of each syllable's detection, the rest of the song is shown in the
 context of the timing for that moment.  This gives an intuition of the
-timing variability within each song and across songs (which may be responsible
-for some of our measured jitter).
+timing variability within each song and across songs (which may be
+responsible for some of our measured jitter).
 

### 2016-08-02 main.tex
Add a reference that Sam Sober mentioned in his review--a nice application.

@@
 representation of song against which auditory signals are compared.
-\cite{Tumer2007pitchshift} played a disruptive sound during renditions of a syllable that were
-slightly above (or below) its average pitch, and showed that the
-presence of this apparently random natural variability in songbird
-motor output is used to drive change in the song.  \cite{Andalman2009}
-extended this experiment and showed that AFP produces a corrective
-signal to bias song away from those disruptions.  \cite{Warren2011}
-showed that that AFP transfers this signal to the robust nucleus of
-the arcopallium (RA) using NMDA-receptor--mediated glutamatergic
-transmission. By looking at song recovery after applying the same
-pitch-shift paradigm, \cite{Canopoli2014} showed that the caudal
-medial nidopallium is implicated in memorising or recalling a recent
-song target, but for neither auditory processing nor directed motor
-learning.
+\cite{Tumer2007pitchshift} played a disruptive sound during renditions
+of a syllable that were slightly above (or below) its average pitch,
+and showed that the presence of this apparently random natural
+variability in songbird motor output is used to drive change in the
+song.  \cite{Andalman2009} extended this experiment and showed that
+AFP produces a corrective signal to bias song away from those
+disruptions.  \cite{Charlesworth2011learning} showed that the song
+change was isolated to within roughly 10 milliseconds (ms) of the
+stimulus, and further proposed an elegant model that predicts the
+shape of the learned response and suggests a simple mechanism.
+\cite{Warren2011} showed that that AFP transfers the error signal to
+the robust nucleus of the arcopallium (RA) using
+NMDA-receptor--mediated glutamatergic transmission. By looking at song
+recovery after applying the same pitch-shift paradigm,
+\cite{Canopoli2014} showed that the caudal medial nidopallium is
+implicated in memorising or recalling a recent song target, but for
+neither auditory processing nor directed motor learning.
 
@@
 \cite{Leonardo1999} used groups of IIR filters with hand-tuned logical
-operators.  Their system had a latency of 50 or 100 milliseconds (ms),
+operators.  Their system had a latency of 50 or 100 ms,
 and they do not report on jitter or accuracy.  As access to
@@
 the target syllable to the firing of the detector's TTL pulse---for
-example, if the target trigger point with respect to the aligned songs
-is at 200 ms, and for a given presentation of the song the detector
-fires at 203 ms, we take the latency on that trial to be 3 ms.
+example, if the target trigger point is at 200 ms with respect to the
+corpus of aligned songs, and for a given presentation of the song the
+detector fires at 203 ms, we take the latency on that trial to be 3
+ms.
 
@@
     %\bibliography{birds}
-\begin{thebibliography}{10}
-
-\bibitem{Kao2005}
-Kao MH, Doupe AJ, Brainard MS.
-\newblock Contributions of an avian basal ganglia--forebrain circuit to
-  real-time modulation of song.
-\newblock Letters to Nature. 2005 February;433:638--643.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v433/n7026/abs/nature03127.html}.
-
-\bibitem{Wang2008interhemispheric}
-Wang CZH, Herbst JA, Keller GB, Hahnloser RHR.
-\newblock Rapid Interhemispheric Switching during Vocal Production in a
-  Songbird.
-\newblock PLOS Biology. 2008 October;6(10).
-\newblock Available from:
-  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.0060250}.
-
-\bibitem{Keller2009}
-Keller GB, Hahnloser RHR.
-\newblock Neural processing of auditory feedback during vocal practice in a
-  songbird.
-\newblock Nature. 2009 January;457:187--190.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v457/n7226/abs/nature07467.html}.
-
-\bibitem{Tumer2007pitchshift}
-Tumer EC, Brainard MS.
-\newblock Performance variability enables adaptive plasticity of 'crystallized'
-  adult birdsong.
-\newblock Nature. 2007 December;450:1240--1244.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v450/n7173/abs/nature06390.html}.
-
-\bibitem{Andalman2009}
-Andalman AS, Fee MS.
-\newblock A basal ganglia-forebrain circuit in the songbird biases motor output
-  to avoid vocal errors.
-\newblock Proceedings of the National Academy of Sciences of the United States
-  of America. 2009 July;106(30):12518--12523.
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2709669/}.
-
-\bibitem{Warren2011}
-Warren TL, Tumer EC, Charlesworth JD, Brainard MS.
-\newblock Mechanisms and time course of vocal learning and consolidation in the
-  adult songbird.
-\newblock Journal of Neurophysiology. 2011 October;106(4):1806--1821.
-\newblock Available from:
-  \url{http://jn.physiology.org/content/106/4/1806.full}.
-
-\bibitem{Canopoli2014}
-Canopoli A, Herbst JA, Hahnloser RHR.
-\newblock A Higher Sensory Brain Region Is Involved in Reversing
-  Reinforcement-Induced Vocal Changes in a Songbird.
-\newblock The Journal of Neuroscience. 2014;34(20):7018--7026.
-\newblock Available from:
-  \url{http://www.jneurosci.org/content/34/20/7018.full}.
-
-\bibitem{Leonardo1999}
-Leonardo A, Konishi M.
-\newblock Decrystallization of adult birdsong by perturbation of auditory
-  feedback.
-\newblock Nature. 1999 June;399:466--470.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v399/n6735/abs/399466a0.html}.
-
-\bibitem{Skocik2013}
-Skocik M, Kozhevnikov A.
-\newblock Real-time system for studies of the effects of acoustic feedback on
-  animal vocalizations.
-\newblock Frontiers in Neural Circuits. 2013 January;6(111).
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3539774/}.
-
-\bibitem{Poole2012}
-Poole B, Markowitz JE, Gardner TJ.
-\newblock The Song Must Go On: Resilience of the Songbird Vocal Motor Pathway.
-\newblock PLOS One. 2012 June;7(6).
-\newblock Available from:
-  \url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0038173}.
-
-\bibitem{Jacobsen2003SlidingDFT}
-Jacobsen E, Lyons R.
-\newblock The sliding DFT.
-\newblock IEEE Signal Processing Magazine. 2003 Mar;20(2):74--80.
-\newblock Available from:
-  \url{http://www.cmlab.csie.ntu.edu.tw/DSPCourse/reference/Sliding\%20DFT.pdf}.
-
-\bibitem{hkp}
-Hertz JA, Krogh AS, Palmer RG.
-\newblock Introduction to the Theory of Neural Computation.
-\newblock Perseus Books; 1991.
-
-\bibitem{Markowitz:2015ko}
-Markowitz JE, Liberti WA, Guitchounts G, Velho T, Lois C, Gardner TJ.
-\newblock Mesoscopic Patterns of Neural Activity Support Songbird Cortical
-  Sequences.
-\newblock PLOS Biology. 2015 June;13(6):e1002158--20.
-\newblock Available from:
-  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.1002158}.
-
-\bibitem{Adelsberger:2005dy}
-Adelsberger H, Garaschuk O, Konnerth A.
-\newblock {Cortical calcium waves in resting newborn mice.}
-\newblock Nature Neuroscience. 2005 August;8(8):988--990.
-\newblock Available from:
-  \url{http://www.nature.com/neuro/journal/v8/n8/full/nn1502.html}.
-
-\bibitem{Schulz:2012fv}
-Schulz K, Sydekum E, Krueppel R, Engelbrecht CJ, Schlegel F, Schr{\"o}ter A,
-  et~al.
-\newblock {Simultaneous BOLD fMRI and fiber-optic calcium recording in rat
-  neocortex}.
-\newblock Nature Methods. 2012 May;9(6):597--602.
-\newblock Available from:
-  \url{http://www.nature.com/nmeth/journal/v9/n6/full/nmeth.2013.html}.
-
-\bibitem{Cui:2013dq}
-Cui G, Jun SB, Jin X, Pham MD, Vogel SS, Lovinger DM, et~al.
-\newblock {Concurrent activation of striatal direct and indirect pathways
-  during action initiation}.
-\newblock Nature. 2013 February;494(7436):238--242.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v494/n7436/full/nature11846.html}.
-
-\bibitem{Adelsberger:2014jd}
-Adelsberger H, Zainos A, Alvarez M, Romo R, Konnerth A.
-\newblock {Local domains of motor cortical activity revealed by fiber-optic
-  calcium recordings in behaving nonhuman primates.}
-\newblock Proceedings of the National Academy of Sciences of the United States
-  of America. 2014 January;111(1):463--468.
-\newblock Available from: \url{http://www.pnas.org/content/111/1/463.full}.
-
-\bibitem{Gunaydin:2014dh}
-Gunaydin LA, Grosenick L, Finkelstein JC, Kauvar IV, Fenno LE, Adhikari A,
-  et~al.
-\newblock {Natural neural projection dynamics underlying social behavior.}
-\newblock Cell. 2014 June;157(7):1535--1551.
-\newblock Available from:
-  \url{http://www.cell.com/cell/fulltext/S0092-8674%2814%2900659-X}.
-
-\bibitem{MandelblatCerf:2014eu}
-Mandelblat-Cerf Y, Las L, Denissenko N, Fee M.
-\newblock {A role for descending auditory cortical projections in songbird
-  vocal learning}.
-\newblock Elife. 2014 January;3(0):e02152--e02152.
-\newblock Available from: \url{https://elifesciences.org/content/3/e02152}.
-
-\bibitem{Noll1970pitchdetection}
-Noll AM.
-\newblock Pitch determination of human speech by the harmonic product spectrum,
-  the harmonic sum spectrum, and a maximum likelihood estimate.
-\newblock In: Proceedings of the symposium on computer processing in
-  communications. vol.~19; 1970. p. 779--797.
-
-\end{thebibliography}
+\bibliography{birds}
 

### 2016-08-09 main.tex
Replace fig:accuracies (Fig. 3) with the new one that includes test results for cage noise.  Add a table for syllable-by-syllable variation in accuracies.

@@
   isolated and detected with false negative and false positive rates
-  well under 0.5\% and 0.01\% respectively. The detector can run on a
+  well under 1\% and 0.001\% respectively. The detector can run on a
   stock Mac Mini with triggering delay of less than a millisecond and
@@
 \item Fast: sub-millisecond latencies, with jitter around 2 ms.
-\item Accurate: false positive rates under 0.02\% and false negative rates under 0.5\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
+\item Accurate: false positive rates under 0.001\% and false negative rates under 1\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
 \item State-of-the-art performance with default parameters.
@@
     simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
-    $t^*_6$. The horizontal axis is time relative to the beginning of
+    $t^*_6$, and we do not show the non-response to presentation of non-song. The horizontal axis is time relative to the beginning of
     the aligned song, and the vertical axis is an index for the 2818
@@
 ALSO need to present a picture of the test syllable or etc\dots} is identical across the three detectors.
+
 \noprint{
-  Typical confusion matrices for these syllables are:
+  The means of the false positive and
 
@@
     & pos & neg \\ 
-    \hline  Detected pos & 99.96451\% & 0.000298\%\\ 
+    \hline  Detected pos & 99.34\% & 0.0015\%\\ 
     neg & 0.0355\% & 99.99970\%\\ 
@@
 training.  Each network's accuracy is shown as a single point in each
-chart in \fig{fig:accuracies}.  Some syllables are easier to identify
-reliably than others, allowing detection point choice to impact the
-detector's accuracy.  For difficult syllables, the training process
+chart in \fig{fig:accuracies}, and the means are shown in Table~\ref{table:accuracies}.  Some syllables are easier to identify
+reliably than others, showing that detection point choice can impact the
+detector's accuracy.  As can be seen from \fig{fig:accuracies}, for difficult syllables, the training process
 occasionally yields a network that performs unusually poorly, but it
@@
 
-  \caption{Accuracy variability over 100 different training runs for each
-    of the six test syllables.  Each dot shows the test-set accuracy
-    for an independently trained detector.  Because the horizontal
-    positions have been randomised slightly so as not to occlude same-valued
-    measurements, test syllable is also indicated by colour.}
+  \caption{Accuracy variability over 100 different training runs for
+    each of the six test syllables.  Each dot shows the test-set
+    accuracy for an independently trained detector.  Because the
+    horizontal positions have been randomised slightly so as not to
+    occlude same-valued measurements, test syllable is also indicated
+    by colour.  The means are given in Table~\ref{table:accuracies}.}
   \label{fig:accuracies}
 \end{figure}
+
+\begin{table}
+  \begin{tabular}{c|rr}
+    Target time & \% True positives & \% False positives \\ \hline
+    $t^*_1$ & 99.34 &  0.0015 \\
+    $t^*_2$ & 99.65 &  0.00040 \\
+    $t^*_3$ & 98.98 &  0.0015 \\
+    $t^*_4$ & 99.30 &  0.00093 \\
+    $t^*_5$ & 99.73 &  0.000094 \\
+    $t^*_6$ & 98.31 &  0.0026
+  \end{tabular}
+  \caption{Mean values for the detection accuracies shown in \fig{fig:accuracies}.}
+  \label{table:accuracies}
+\end{table}
 
@@
 \item The detector is accurate. False negative and false positive
-  rates can be well under 0.5\% and 0.01\% respectively, and trading
+  rates can be well under 1\% and 0.001\% respectively, and trading
   these two numbers off against each other is through a single

### 2016-08-10 main.tex
Minor word choice change in Accuracies results.

@@
 training.  Each network's accuracy is shown as a single point in each
-chart in \fig{fig:accuracies}, and the means are shown in Table~\ref{table:accuracies}.  Some syllables are easier to identify
-reliably than others, showing that detection point choice can impact the
-detector's accuracy.  As can be seen from \fig{fig:accuracies}, for difficult syllables, the training process
-occasionally yields a network that performs unusually poorly, but it
+chart in \fig{fig:accuracies}, and the means are shown in Table~\ref{table:accuracies}.  Evidently some syllables are easier to identify
+reliably than others; thus, if the experimental design permits, it can be worthwhile to choose an easily identifiable syllable.
+Furthermore, as can be seen from \fig{fig:accuracies}, for difficult syllables the training process
+occasionally yields a network that performs unusually poorly.  It
 is easy to spot these networks by their performance either as measured

### 2016-08-10 main.tex
Fix fonts on Fig. 1.  Move discussion of why we show only 1 bird to Sec. 2.3 (Quantification).  Other minor textual changes.

@@
 \begin{figure}
+  % Width: 8, Height: 3, Fonts scale: 120%
   \includegraphics[width=\textwidth]{Fig1}
@@
 
+We have used this detector for song alignment and feedback experiments
+in 5 behaving birds, and for offline analysis on a few others.  For
+simplicity, and to test robustness to imperfect choice of ``syllable'', we present performance on 6 randomly chosen points in one
+bird's song.  However, the results presented here are representative of
+real-world performance over many birds in our lab.
+
 We divided the data into training and test sets.  For the results
@@
 into internal training, validation, and test sets.  Because we do our
-own testing on a distinct and much larger data set, we instructed the
-toolbox to use 80\% of its training set for training, and the
+own testing on a distinct data set, we instructed the
+toolbox to use 80\% of its input set for training, and the
 remaining 20\% for validation, with no hold-out test data.
@@
 defining accuracy on a per-song basis, such that a song without the
-target syllable counted as a single true negative.  Computing the
-optimal output thresholds on a per-frame basis resulted in higher
-thresholds and thus a lower false-positive rate, with minimal
-consequences to the false-negative rate, while also providing a valid
+target syllable counted as a single true negave.  Computing the
+optimal output thresholds on a per-frame rather than a per-song basis resulted in higher
+thresholds and thus a slight reduction in both the false-positive and true-positive rates (reversible by increasing $C_n$), while also providing a valid
 definition of false-positive rate for data streams that had not been
@@
 \label{sec:results}
-
-We have used this detector for song alignment and feedback experiments
-in 5 behaving birds, and for offline analysis on a few others.  For
-simplicity we present performance on 6 randomly chosen points in one
-bird's song, but the results presented here are representative of
-real-world performance in our lab.\marginpar{Ok?}
 

### 2016-08-10 main.tex
Change repository locations, some wording.

@@
 
-{\em If accepted, before final submission, all code will be moved to the Gardner Lab github repository, final release versions generated, and DOIs created as described in https://guides.github.com/activities/citable-code/ so that our results may be reproduced and the software deployed.
+{\em If accepted, before final submission, DOIs for the latest code will be created as described in https://guides.github.com/activities/citable-code/ so that our results may be reproduced and the software deployed.
 For now, in lieu of final release DOIs, we provide these links to our development repositories.
@@
 \noindent Our implementation of the syllable
-detector is available under an open-source license {\em (we are still evaluating which open source license is the most suitable for this software):}
+detector is available under an open-source license {\em (Tentatively GPL, but we are still evaluating which open source license is the most suitable for this software):}
 \begin{description}
@@
   \begin{trivlist}
-  \item {\tt https://github.com/bwpearre/birds/align}
+  \item {\tt https://github.com/gardner-lab/syllable-detector-learn}
   \end{trivlist}
@@
   \item[Installation:]\hfill
-    \begin{trivlist}
+    \begin{itemize}
     \item Clone the Github repository onto your computer.
-    \end{trivlist}
+    \item You also need the following packages from MATLAB File Exchange:
+      \begin{description}
+      \item[Significant Figures] by Teck Por Lam.  File ID \#10669.
+      \item[Generate maximally perceptually-distinct colors] by Tim Holy.  File ID \#29702.
+      \end{description}
+    \end{itemize}
   \item[Usage:]\hfill
@@
       \item The main programme is {\tt learn\_detector.m}.
-    \item The top section of {\tt learn\_detector.m} contains global configuration variables.
-The important ones are the data directory and the bird name: as supplied, the software looks for {\tt \textless data directory\textgreater/\textless bird name\textgreater/song.mat}.
-Modify as necessary.
-    \item The format for the data file {\tt \textless data directory\textgreater/\textless bird name\textgreater/song.mat} is specified in {\tt birds/align/README.md}: a MATLAB file containing aligned song, nonsong (calls, cage noise, etc), and the sampling frequency.
-    \item The configuration file {\tt \textless data directory\textgreater/\textless bird name\textgreater/params.m} is also described in {\tt README.md}.
-If the file is empty, the programme will display an image showing the average of the song spectrograms (frequency vs.~time), tell the user how to define syllables of interest, and exit.
-If {\tt times\_of\_interest} is defined, the programme will train the neural network.
-{\tt params.m} can also override any of the configuration parameters at the top of {\tt learn\_detector.m}.
-    \item The software will produce a neural network file, {\tt detector\_bird\_times\_\dots.mat} (the ellipses signify other information encoded in the filename).
-It will also produce the audio test file described in Methods.
+    \item The top section of {\tt learn\_detector.m} contains global
+      configuration variables.  The important ones are the data
+      directory and the bird name: as supplied, the software looks for
+      {\tt \textless data directory\textgreater/\textless bird
+        name\textgreater/song.mat}.  Modify as necessary.
+    \item The format for the data file {\tt \textless data
+      directory\textgreater/\textless bird name\textgreater/song.mat}
+      is specified in {\tt syllable-detector-learn/README.md}: a
+      MATLAB file containing aligned song, nonsong (calls, cage noise,
+      etc), and the sampling frequency.
+    \item The configuration file {\tt \textless data
+      directory\textgreater/\textless bird name\textgreater/params.m}
+      is also described in {\tt README.md}.  If the file is empty, the
+      programme will display an image showing the average of the song
+      spectrograms (frequency vs.~time), tell the user how to define
+      syllables of interest, and exit.  If {\tt times\_of\_interest}
+      is defined, the programme will train the neural network.
+      Optionally, {\tt params.m} can also override any of the
+      configuration parameters at the top of {\tt learn\_detector.m}.
+    \item The software will produce a neural network file, {\tt
+      detector\_bird\_times\_\dots.mat} (the ellipses signify other
+      information encoded in the filename).  It will also produce the
+      audio test file described in Methods.
     \end{itemize}
@@
       \end{itemize}
-    \item[Customization:]\hfill
+    \item[Customisation:]\hfill
       \begin{itemize}
@@
     \end{description}
-  \item[LabVIEW:] {\tt https://github.com/bwpearre/birds/align}
+  \item[LabVIEW:] {\tt https://github.com/gardner-lab/\dots\ syllable-detector-runtime-labview}
     \begin{description}
@@
       \begin{itemize}
-      \item The microphone should be connected through a preamplifier to the data acquisition card, by default on Analog Input 0 (AI0).
+      \item The microphone should be connected, optionally through a preamplifier, to the data acquisition card, by default on Analog Input 0 (AI0).
       \item Outputs are on the Digital I/O (DIO) channels.  See the block diagram of {\tt nnfft.vi} for documentation, or just connect to DIO0.

### 2016-08-12 main.tex

@@
   \item[MATLAB:] {\tt https://github.com/gardner-lab/syllable-detector-matlab}
-    \marginpar{Jeff: Can you flesh this out?}
     \begin{description}
     \item[Installation:]\hfill
-      \begin{trivlist}
+      \begin{itemize}
       \item Clone the Github repository onto your computer.
-      \end{trivlist}
+      \item Usage requires MATLAB with the Data Acquisition Toolbox and the 
+      Parallel Computing Toolbox.
+      \end{itemize}
     \item[Connection:]\hfill
+      \begin{itemize}
+      \item Either connect the microphone directly into the microphone jack or through a 
+      preamp into the audio input jack. The current implementation only supports a single 
+      channel of audio.
+      \item To generate TTL signals via the audio output, use the headphone or audio output
+      jack.
+      \item To generate TTL signals via an Arduino, switch to the {\tt arduino} branch, 
+      load the MATLAB Arduino IO sketch onto the Arduino (the sketch is available in the 
+      Github repository, and provides a simple serial protocol for controlling pins). 
+      Connect the Arduino to the computer via USB. The TTL signal will be available on
+      pin 7.
+      \end{itemize}
     \item[Usage:]\hfill
+      \begin{itemize}
+      \item Switch to the directory of the repository and run {\tt nndetector\_live}. You
+      can optionally specify parameters as string/value pairs. When you run the command,
+      you will be prompted to select an input device, output device and select a ``.mat''
+      file containing the trained detector.
+      \item Note that you may need to tune the input and output buffer based on computer 
+      performance. Try decreasing the size of the buffers until you see warnings about buffer
+      under/overruns, then increase the value slightly. We usually found 2-3ms worked reliably.
+      To specify the buffer sizes, use the {\tt buffer\_size\_input} and 
+      {\tt buffer\_size\_output} parameters.
+      \item Information about the detector is logged to a file called ``detector\_status.log''.
+      \end{itemize}
     \end{description}
@@
       audio output must have at least the same number of channels as the input.
-      \item To generate TTL signals via an Arduino, load the MATLAB arduino IO sketch 
+      \item To generate TTL signals via an Arduino, load the MATLAB Arduino IO sketch 
       onto the Arduino (the sketch is available in the Github repository, and provides a 
@@
       are interested in extending or modifying the functionality. To hook into either 
-      the raw input or output, look at the ``ViewControllerProcessor.swift'' file. The 
+      the raw input or output, look at the ``Processor.swift'' file. The 
       {\tt Processor} class acts as a coordinator, receiving income audio, passing that
@@
 \end{document}
+
+s

### 2016-09-01 main.tex
More minor wording changes.

--- 2016-08-10 main.tex
@@
   isolated and detected with false negative and false positive rates
-  well under 1\% and 0.001\% respectively. The detector can run on a
+  well under 0.5\% and 0.01\% respectively. The detector can run on a
   stock Mac Mini with triggering delay of less than a millisecond and
@@
 homologue of mammalian basal ganglia consisting of a few distinct
-brain areas concerned with the acquisition and learning of song.  For
+brain areas concerned with the production and learning of song.  For
 example, \cite{Kao2005} stimulated the lateral magnocellular nucleus
@@
 
-Despite the power and versatility of such experiments, there is no standard syllable detector.  Considerations include:
+Despite the power and versatility of vocal feedback experiments, there
+is no standard syllable detector.  Desiderata include:
 \begin{description}
@@
 operators.  Their system had a latency of 50 or 100 ms,
-and they do not report on jitter or accuracy.  As access to
+and they did not report on jitter or accuracy.  As access to
 computational resources has improved, approaches have changed: in
 2009, \cite{Andalman2009} still used hand-tuned filters, but ran them
-on a Tucker-Davis Technologies digital signal processor.  They report
+on a Tucker-Davis Technologies digital signal processor.  They reported
 a latency of around 4 ms, but as with other filter-bank techniques, it
@@
 \cite{Keller2009} applied a neural network to a spectral image of
-song.  They report a jitter of 4.3 ms, but further implementation and
+song.  They reported a jitter of 4.3 ms, but further implementation and
 performance details are not available.  In 2011, \cite{Warren2011}
 matched spectral templates to stable portions of syllables in 8-ms
-segments.  They report a jitter of 4.5 ms, and false-negative and
+segments.  They reported a jitter of 4.5 ms, and false-negative and
 false-positive rates of up to 2\% and 4\%, respectively.  Hardware
@@
 latency and jitter of just a few microseconds, and the detection
-computation they use should not much increase that.  Drawbacks are
-that some hand-tuning may still be required, and that they report
+computation they use should not much increase that.  They reported
 false-negative rates around 4\% and 7\% for zebra finches and
@@
 song using a neural network applied to the song spectrogram, and
-outputs a TTL pulse (a brief 5-volt pulse) at the chosen moment. The approach consists of
-three steps:
+outputs a TTL pulse (a brief 5-volt pulse) at the chosen moment. The
+approach consists of three steps:
 
@@
 \item Fast: sub-millisecond latencies, with jitter around 2 ms.
-\item Accurate: false positive rates under 0.001\% and false negative rates under 1\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
+\item Accurate: false positive rates under 0.02\% and false negative rates under 0.5\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
 \item State-of-the-art performance with default parameters.
@@
 (``undirected''); we train and test on undirected song since this is
-more challenging.  Recordings were made inside a sound-isolating
+more commonly studied, as well as more challenging.  Recordings were made inside a sound-isolating
 chamber in which was mounted a recording microphone (Audio-Technica
@@
 \begin{figure}
-  % Width: 8, Height: 3, Fonts scale: 120%
   \includegraphics[width=\textwidth]{Fig1}
@@
 depend on the makeup of the non-song data, we have found that training
-with as low as a 1:1 ratio of non-song to song yields excellent
-results on most birds.  These data are also included in our accuracy tests.
+with as low as a 1:1 ratio of non-song to song---or roughly 10-20
+minutes of non-song audio---yields excellent results on most birds.
+These data are also included in our accuracy tests.
 
@@
 
-% technically, it is LabVIEW
 \subsubsection{LabVIEW}
@@
 While it did not perform as well on our test system as the other
-implementations, the convenience may outweigh the slight timing
+implementations, the convenience may outweigh the timing
 performance penalty for some experiments.  Key to minimising jitter is
@@
 
-Similar to the Swift implementation, it is also possible to modify the
+As with the Swift detector, it is also possible to modify the
 MATLAB implementation to generate the TTL pulse from a microcontroller
@@
 \label{sec:quantify}
-
-We have used this detector for song alignment and feedback experiments
-in 5 behaving birds, and for offline analysis on a few others.  For
-simplicity, and to test robustness to imperfect choice of ``syllable'', we present performance on 6 randomly chosen points in one
-bird's song.  However, the results presented here are representative of
-real-world performance over many birds in our lab.
 
@@
 into internal training, validation, and test sets.  Because we do our
-own testing on a distinct data set, we instructed the
-toolbox to use 80\% of its input set for training, and the
+own testing on a distinct and much larger data set, we instructed the
+toolbox to use 80\% of its training set for training, and the
 remaining 20\% for validation, with no hold-out test data.
@@
 non-detected non-syllable counting as a true negative, we also tried
-defining accuracy on a per-song basis, such that a song without the
-target syllable counted as a single true negave.  Computing the
-optimal output thresholds on a per-frame rather than a per-song basis resulted in higher
-thresholds and thus a slight reduction in both the false-positive and true-positive rates (reversible by increasing $C_n$), while also providing a valid
+defining accuracy on a per-song basis, such that a song-length sample without the
+target syllable counted as a single true negative.  Computing the
+optimal output thresholds on a per-frame basis resulted in higher
+thresholds and thus a lower false-positive rate, while also providing a valid
 definition of false-positive rate for data streams that had not been
@@
 may include previous syllable).  Often more important is the
-variability in the latency?  We define jitter as the standard
+variability in the latency.  We define jitter as the standard
 deviation of the latency.
@@
 
+We have used this detector for song alignment and feedback experiments
+in 5 behaving birds, and for offline analysis on a few others.  For
+simplicity we present performance on 6 randomly chosen points in one
+bird's song, but the results presented here are representative of
+real-world performance in our lab.
+
 \fig{fig:song} shows the song for which we present accuracy and timing
@@
 context of the timing for that moment.  This gives an intuition of the
-timing variability within each song and across songs (which may be
-responsible for some of our measured jitter).
+timing variability within each song and across songs (which is
+responsible for a small amount of our measured jitter, but note that few of the songs are significantly timeshifted).
 
@@
     simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
-    $t^*_6$, and we do not show the non-response to presentation of non-song. The horizontal axis is time relative to the beginning of
+    $t^*_6$. The horizontal axis is time relative to the beginning of
     the aligned song, and the vertical axis is an index for the 2818
@@
 that the detector will see---only the non-song data may be different.
-%Accuracy \marginpar{Should be!
-Verify!
-And/or explain that there is a bug in the LabVIEW implementation, if there in fact is.
-ALSO need to present a picture of the test syllable or etc\dots} is identical across the three detectors.
-
-\noprint{
-  The means of the false positive and
-
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_1$
-    & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.34\% & 0.0015\%\\ 
-    neg & 0.0355\% & 99.99970\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_2$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.92903\% & 0.000596\%\\ 
-    neg & 0.0710\% & 99.99940\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_3$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.68062\% & 0.00194\%\\ 
-    neg & 0.319\% & 99.99806\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_4$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.82257\% & 0.00239\%\\ 
-    neg & 0.177\% & 99.99761\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_5$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.78708\% & 0\\ 
-    neg & 0.213\% & 100\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_6$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.71611\% & 0.00552\%\\ 
-    neg & 0.284\% & 99.99448\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-}
 
@@
 training.  Each network's accuracy is shown as a single point in each
-chart in \fig{fig:accuracies}, and the means are shown in Table~\ref{table:accuracies}.  Evidently some syllables are easier to identify
-reliably than others; thus, if the experimental design permits, it can be worthwhile to choose an easily identifiable syllable.
-Furthermore, as can be seen from \fig{fig:accuracies}, for difficult syllables the training process
-occasionally yields a network that performs unusually poorly.  It
+chart in \fig{fig:accuracies}.  Some syllables are easier to identify
+reliably than others, allowing detection point choice to impact the
+detector's accuracy.  For difficult syllables, the training process
+occasionally yields a network that performs unusually poorly, but it
 is easy to spot these networks by their performance either as measured
@@
 
-  \caption{Accuracy variability over 100 different training runs for
-    each of the six test syllables.  Each dot shows the test-set
-    accuracy for an independently trained detector.  Because the
-    horizontal positions have been randomised slightly so as not to
-    occlude same-valued measurements, test syllable is also indicated
-    by colour.  The means are given in Table~\ref{table:accuracies}.}
+  \caption{Accuracy variability over 100 different training runs for each
+    of the six test syllables.  Each dot shows the test-set accuracy
+    for an independently trained detector.  Because the horizontal
+    positions have been randomised slightly so as not to occlude same-valued
+    measurements, test syllable is also indicated by colour.}
   \label{fig:accuracies}
 \end{figure}
-
-\begin{table}
-  \begin{tabular}{c|rr}
-    Target time & \% True positives & \% False positives \\ \hline
-    $t^*_1$ & 99.34 &  0.0015 \\
-    $t^*_2$ & 99.65 &  0.00040 \\
-    $t^*_3$ & 98.98 &  0.0015 \\
-    $t^*_4$ & 99.30 &  0.00093 \\
-    $t^*_5$ & 99.73 &  0.000094 \\
-    $t^*_6$ & 98.31 &  0.0026
-  \end{tabular}
-  \caption{Mean values for the detection accuracies shown in \fig{fig:accuracies}.}
-  \label{table:accuracies}
-\end{table}
 
@@
 \item The detector is accurate. False negative and false positive
-  rates can be well under 1\% and 0.001\% respectively, and trading
+  rates can be well under 0.5\% and 0.01\% respectively, and trading
   these two numbers off against each other is through a single

### 2016-09-06 main.tex

--- 2016-08-12 main.tex
@@
 simplicity, and to test robustness to imperfect choice of ``syllable'', we present performance on 6 randomly chosen points in one
-bird's song.  However, the results presented here are representative of
+bird's song.  The results presented here are representative of
 real-world performance over many birds in our lab.
@@
 non-detected non-syllable counting as a true negative, we also tried
-defining accuracy on a per-song basis, such that a song without the
-target syllable counted as a single true negave.  Computing the
+defining accuracy on a per-song basis, such that a song-length sample without the
+target syllable counted as a single true negative.  Computing the
 optimal output thresholds on a per-frame rather than a per-song basis resulted in higher

### 2016-09-07 rebuttal.txt
Preparing for submission to PLOS ONE:
- Rewrite cover letter.

- Rebuttal of reviewer comments.

- A few more changes to the document to answer clarification questions.

> Reviewer #1: The authors describe a novel technique for real-time
> identification of individual vocal gestures (syllables) in songbirds
> and provide a range of performance metrics for their new technique. I
> think that this method potentially represents an important advance in
> the execution of a very useful and widely applied method for studying
> the neural basis of vocal production and plasticity. There are however
> a few revisions that would be useful.
> 
> Also, since my comments will likely give me away anyhow, I’ll just go
> ahead and identify myself – this is Sam Sober. Nice work guys!
> 
> First, I’ll comment that a real-time detection paradigm that is fast,
> accurate, and easy to implement/modify is sorely lacking in the field,
> and that the deficiencies of current techniques (in use in my own and
> other labs) both slow down current research and likely constitute a
> barrier to entry to other labs that might otherwise make important
> contributions. So – making this method widely available and, just as
> importantly, easy to implement, would be a great step forward.
> 
> Next, here are my major comments/suggestions:
> 
> 1. Whether or not people actually use this technique is going to
> depend strongly on ease of implementation. My understanding is that
> users need not be experts in implementing MATLAB’s neural network
> toolbox (and other computational steps) to use this method, however
> the level of technical expertise required of a user is not clear in
> the manuscript, especially with respect to the various
> MATLAB/Swift/LabView realtime implementations. Specifically, although
> the computational steps are well described, it is somewhat opaque how
> much user effort is involved in setting up and running the detection
> algorithm. I suggest adding a figure/table/supplemental document that
> clarifies all required user inputs and any additional code writing
> that a user must perform. This would both constitute a compact users
> guide, outlining all steps required, and also would distinguish
> “optional” steps (resetting number of hidden units per syllable, etc.)
> where users could use default values, from “required” steps (choosing
> detection points, etc.).

We try to provide a complete description of the method, but since we
are also distributing our implementation, this makes sense.  Therefore
we have added installation/configuration instructions in Appendix A.

> 2. The discussion of latency/timing is somewhat confusing, and I
> believe it omits what I (and other investigators) consider to be the
> most crucial metric of detection performance – the variation in
> latency from the onset of the targeted syllable to the firing of the
> TTL pulse.

We had already characterised the timing from the completion of the
target recognition region to the firing of the TTL pulse--this is what
we define as "Jitter" (line 27, Section Quantification.Timing
(beginning around line 342), and also extensively throughout the
document).  We have tried to clean up those references and clarify
our description.

> Charlesworth et al. (“Learning the microstructure of
> successful behavior”) showed that learning after white noise targeting
> is temporally localized at less than 10 msec timescale. Furthermore,
> in my own and other labs, the key predictor of whether online
> targeting will successfully drive learning is the standard deviation
> of detections/triggers relative to syllable onset, with syllable onset
> defined by some measure of sound amplitude for each individual
> iteration of the syllable. I believe (and please correct me if I’m
> wrong) that the song alignment technique used here does not provide
> this statistic on a syllable-by-syllable basis.

((1a)) That is correct, but the alignment does appear to be
sufficiently accurate:

We refer to this syllable-by-syllable adjustment to the alignment as
micro-realignment.  Even without that micro-realignment, jitter was
generally around 2 ms, as can be seen throughout Section 3.2,
especially Figs. 4-7.  Since this is well within the 10-ms window
suggested by the literature and indeed barely above the theoretically
optimal jitter given our detection parameters, we decided that the
improvement that could be achieved by micro-realigning song was
unimportant.

> So – please provide this measure where appropriate in Results. I
> suspect that the jitter measurements for real-time performance will
> not be much different with this metric than what is currently
> reported, but it would be very useful nonetheless.

We believe that we have provided a sufficient analysis of jitter, and
that our results are good enough to obviate the improvement you
suggest.

> 3. The authors should quantify additional examples of real-time
> performance. Most analysis focuses on the song of a single bird, and
> real-time latency performance is only reported for detection point
> t4 (which, given the strong harmonic content of t4, I suspect is the
> easiest syllable to target). The authors should provide similar
> performance metrics for the other detection times.

Actually, we already had done so: see Fig. 5 (and section 3.2.syllable
choice).

> Also, the authors should provide performance metrics for at least
> one additional bird;

((1b)) We have achieved similar results on quite a few birds, and
present only one bird simply for reasons of conciseness.  Since the 6
target points in that bird's song are essentially random (50-ms
intervals in a song that includes a range of typical sound types and
does not have any 50-ms periodicity), we just added a comment that
these results are indeed representative.

> although the three syllables shown in Figure 1 look like fairly
> typical ZF syllables, a method such as this one should be shown to
> work on multiple syllables/animals. This additional bird could be the
> zebra finch shown in Fig 8-9. Alternately, given that many researchers
> use Bengalese finches in real-time detection experiments,
> demonstrating the efficacy of the method in a Bengalese would be
> useful.

It would.  But we don't have a convenient data set, and we feel that
it is more important to get the technique out there quickly than to
continue to characterise it in yet more detail.

> 4. I am ambivalent about the use of the fiber photometry data in this
> paper. Although this looks like a great technical achievement, I see
> two problems with including it here. First is that it is not actually
> necessary to demonstrate that the targeting software works – this
> could be accomplished just by quantifying the accuracy of the white
> noise hits shown in Figure 8, or by showing that white noise targeting
> using the new method successfully drives learning (indeed, driving
> learning is what most potential users of this method will likely care
> about). Second is that for the photometry data to be presented, more
> analysis is needed. How precisely was the GCaMP6s/f confined to HVC
> shelf versus HVC or more ventral areas? If there was spillover, how
> would this affect interpretation of the data shown in Fig. 9? If
> individual animals were co-infected with GCaMP6s and GCaMP6f, which
> expressed more robustly and/or contributed more to the signal?
>
> In general, it seems odd to use photometry – itself a methodological
> innovation in songbirds – to test the implementation of another
> innovative method. I think the authors should either (1) provide a
> more detailed analysis of the photometry technique, (2) confine
> analysis to real-time implementation of white noise (WN) targeting,
> without photometery data, or (3) demonstrate that pitch-contingent WN
> drives vocal learning with the new software. Among these options I
> think (3) would be the most useful, but I leave this up to the Editor
> and the authors.

((1c)) These are very good questions.  We wanted to throw in a
real-world test case, but you are right that it is not the most
appropriate one, and our other experimental test cases are far from
publication as yet.  So we've removed this section.

> Minor comments:
> 
> Line 318-319 – As I mentioned above, I find the discussion of
> timing/latency a bit confusing. For example – “time taken from the
> presentation of the target syllable to the firing of the TTL pulse” –
> I can’t tell whether the first time point is the onset time of the
> target syllable on a trial-by-trial basis or something averaged across
> songs with variable timing.

See my (1a) above.  I hope we've clarified this.

> Figure 3 – Although the “correct detection” performance is great for
> all syllables, I notice that it’s a bit worse for t3 and t6. Is this
> because these two syllables, which are acoustically very similar, are
> mistaken for each other? I’m guessing that the LabView code that my
> lab currently uses would never be able to distinguish between these,
> so the fact that performance sinks to “only” ~99% for these is a major
> strength of the technique, and you might want to point this out.

Yes, the software sometimes confuses between these two.  Noted.

> Please clarify what data/parameters are saved by the current version
> of the software. Most users will want the software to save detailed
> records of acoustic input, trigger/detection times, time-dependent
> evaluation of similarity, etc. for each song trial. If no such data
> are saved, it would be useful to indicate (either in the manuscript,
> in the code, or both) how this could be added.

None--we use other software for that.  But the new user's guide
(Appendix A) now includes a pointer for where to go to make such
modifications for the Swift implementation.

> Reviewer #2: The authors have performed a useful study to examine the
> suitability of diverse hardware platforms for real time vocal syllable
> detection. Many solutions to this problem exist already but the
> current study seemingly outperforms these. Making the software open
> access will be particularly useful for researchers in the field.
> 
> The strength of the study is that their solution achieves low latency
> and low jitter. For detecting target points and delivering TTL Pulses
> during song they implemented a neural network detector on 3 platforms:
> LabView (hard to debug), MATLAB (slowest), and Swift+Serial (fast, MAC
> based).  Their approach for syllable detection is a neural network
> (with 2 layers) that requires relatively little computation
> time. Overall the work would be outstanding if the authors had also
> evaluated an FPGA solution, but presumably such an approach is highly
> demanding on the programming side.
> 
> The main concern with the presentation of the work is that the
> classification could be better evaluated. First of all, they only
> train and test on six target points of only one song.

Yes.  See (1b) above.

> Most
> importantly, they do not do a life bird test.

I'm not sure what this means.  The fibre photometry test is a live
bird test (and ongoing research in our lab uses the detector with
great success), but of course all quantitative tests must be done on
recordings in order to systematically analyse accuracies of various
detectors.

> In fact, they haven't included calls, wing flaps, and other noises
> in their training or testing sets.  Thus, it seems there is no
> realistic evaluation of their system.

That is an excellent point!  We have remedied this.  See the paragraph
around line 160.  Also, the new Results/Accuracy graph (Fig. 3) shows
accuracy figures including lots of cage noise samples (about half an
hour's worth).  It is slightly different from the previous version of
Fig. 3, since the introduction of cage noise decreases the ratio of
positive/negative examples and we have not adjusted the false positive
vs. false negative cost parameter (which was arbitrarily set at 1).
Consequently correct detections and incorrect detections have both
somewhat decreased from our earlier results.

> Minor:
> 
> Their method requires to find an "optimal threshold" for the output of
> the network but they don't say how they choose it. Perhaps cross
> validation could help (their network does regression, but of course
> the purpose is to output a binary signal, so the evaluation needs to
> be binary).

We do say how we choose it: see Sec. 2.1, "Computing optimal output
thresholds".

> The paper has quite a few sections on fiber photometry, but I find
> these section quite irrelevant to the main point of the paper, which
> is to detect syllables in an audio recording. Perhaps the photometry
> could be left out?

See (1c) above.

> How robust is the method to errors in the training set?

There is an infinite space of possible errors in the training set, so
this question is impossible to answer.  However, no effort was made to
compensate for errors in the alignment tool, and the results aren't
bad, so I guess it depends on what kinds of errors you're worried
about.

> the writing and presentation could be improved.

Always, but that is too vague to address.

> e.g. fig 1 increase font size of labels

Thanks.  Done.

> Line 189: It takes so long to train a network with just 2 to 4 hidden
> neurons?

Big data sets!  But we changed the training algorithm and now things
are about 10 times faster.

> maybe reducing the training data would help (reduce the
> negative targets only).

This directly contradicts your request that we add cage noise,
flapping, etc.  Luckily, training time shouldn't matter--it's online
performance that counts, right?  However, the change to scaled
conjugate gradient training has sped things up considerably.

> Line 203: How was Cn chosen?

Arbitrarily, as noted in the text.

> Line 201: gradient descent, not hillclimbing (on cost). cross
> validation.

My comment applied to either one, but that text (it just described
something we didn't do) was unnecessary, so I removed it.

> Line 292: needs real data including all kinds of sounds present in the
> cage

Done.

> Line 428: robustness to these variables needs to be documented.

We documented robustness to many choices.  There are always more, but
this was supposed to be a 5-page paper.  We have to stop somewhere.

> Line 238: what is no user loads?

No loads from users (i.e. operating system processes only).

> Fig. 8, what are the dark lines intermingled between the yellow lines
> at the bottom quarter of the figure? false negatives?

Removed as per (1c).

> Reviewer #3: This paper describes a program to detect syllable in ZF
> songs directed at the ZF brain and learning community. The main method
> relies on neural network learning on the zscored spectrogram and
> optimization techniques to diminish latency and detection time
> variations.
> 
> I cannot recommend the paper for publication for the following reasons:
> 
> - there does not seem to be any real and documented innovation here to
>   justify a publication - it seems more like several enhancements
>   designed for a specific use in a neuroscience lab

Agreed.  But we feel it is useful to the community (as do the other
two reviewers, as well as several of our colleagues), and various
researchers have requested that we publish it.

> - which lead to my second reason: the tool and the data required is
>   too narrowly focused on one particular study on ZF. My contention is
>   that this would be usable for a very limited number of teams.

We agree with you, although several have noted that this model is
extremely powerful and the absence of tools such as this one is a
barrier to entry into the field.

> Several points:
> 
> a) From a machine learning community point of view the article is
> extremely frustrating ; nothing is done in the statistics of syllables
> to justify the architecture and its hyper parameters.

Agreed.  However, it is not a machine learning paper, but rather a
paper that happens to use a simple machine learning tool to build
something useful.

> b) also the dataset is super standardized and it should have been
> tested on others labs data and why not different 'strains' of ZF
> (American, European and Wild type).

Yes, more birds/species would always be better, but see (1b).

> In other word I highly doubt that the authors' system can be
> generalised (even though they claim it can be extended to another
> species - without providing any insights on why this would be true).

Some insight added, but we agree that it's a tool for a specific
community.  As is any tool.

> c) the training is done on around 3k song annotated and aligned and
> never provide the readers and potential user how to obtain such
> dataset.

((3a)) This is incorrect.  We cited the article in which alignment
software is described, and we linked to the alignment software.  The
only step that we have not described is using a recording device,
which is outside the scope of the paper.

> With my own experience on working with birds in social context, such
> dataset is extremely difficult to obtain without very invasive and
> constrained method.

Yup.

> In the present manuscript, it has been obtained, presumably, by
> letting a bird to its own device in a small cage for hours to
> extract songs. Many experiments do not want such harsh treatment.

Without isolating a bird, isolating a song requires an entirely
different set of techniques, which would also make for interesting
research.  It sounds like we proposed a solution to Problem A, but you
wish we'd solved Problem B instead.

> d) Also here to digress a bit, it is not specified if these are
> directed and undirected songs (or both). The distinction can be of
> import because they have different syllable statistics and of course
> this should import on the training and/or the prediction.

Good point.  Now specified.

> e) It is also assume that the user has means to extract / align song
> automatically event though this is never mentioned in the manuscript.

Wrong.  See (3a).

> f) the experiment with Fibre Photometry is really nice but is a
> distraction from the main paper and brings nothing to the story

Right.  See (1c).

> g) very minor: its Levenberg (not Levenburg) and it is a very unusual
> choice here (usually some kind of enhanced backpropagation are used)
> probably related to the black-boxy nature of the matlab toolbox

Thanks!  It's a good point.  Levenberg-Marquardt is the default for
MATLAB's toolbox, and we wanted to use as many defaults as we could
(preserving the black-boxy nature of third-party tools as much as
possible in order to emphasise that the specific choice probably
doesn't really matter very much--hopefully any regression tool would
work similarly).  But it did turn out in the end that Scaled Conjugate
Gradient produced similar results faster.

> in short, all these criticisms in fact converge to the main weakness:
> this paper/software is directed to a very specific community both in
> scope and type of data needed for it to work and the authors should
> aim to a journal directed at this community.

Yes.

### 2016-09-07 main.tex
Preparing for submission to PLOS ONE:
- Rewrite cover letter.

- Rebuttal of reviewer comments.

- A few more changes to the document to answer clarification questions.

@@
 
-We evaluate the time taken from the presentation of each instance of
-the target syllable to the firing of the detector's TTL pulse---for
-example, if the target trigger point is at 200 ms with respect to the
-corpus of aligned songs, and for a given presentation of the song the
-detector fires at 203 ms, we take the latency on that trial to be 3
-ms.
+We evaluate the time taken from the complete presentation of each
+instance of the target syllable to the firing of the detector's TTL
+pulse.  For example, if the recognition region is 30 ms long and the
+target trigger point is at 200 ms with respect to the corpus of
+aligned songs, then when the detector has seen the region from
+170--200 ms, it should recognise the syllable.  If for a given
+presentation of the song the detector fires at 203 ms, we take the
+latency on that trial to be 3 ms.
 
@@
 \fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
-songs were always 100\%).  For each test syllable we trained
-100 different
-networks, which differ from each other in the random initialisation of
-network weights and on which random subset of our songs was used for
-training.  Each network's accuracy is shown as a single point in each
-chart in \fig{fig:accuracies}, and the means are shown in Table~\ref{table:accuracies}.  Evidently some syllables are easier to identify
-reliably than others; thus, if the experimental design permits, it can be worthwhile to choose an easily identifiable syllable.
-Furthermore, as can be seen from \fig{fig:accuracies}, for difficult syllables the training process
-occasionally yields a network that performs unusually poorly.  It
-is easy to spot these networks by their performance either as measured
-as part of the training process or on the test audio file, and
-re-train when necessary.
+songs were always 100\%).  For each test syllable we trained 100
+different networks, which differ from each other in the random
+initialisation of network weights and on which random subset of our
+songs was used for training.  Each network's accuracy is shown as a
+single point in each chart in \fig{fig:accuracies}, and the means are
+shown in Table~\ref{table:accuracies}.  Evidently some syllables are
+easier to identify reliably than others (for example, $t^*_3$ and
+$t^*_6$ are occasionally confused due to their similar appearance, as
+could be guessed from \fig{fig:song}); thus, if the experimental
+design permits, it can be worthwhile to choose an easily identifiable
+syllable.  Furthermore, as can be seen from \fig{fig:accuracies}, for
+difficult syllables the training process occasionally yields a network
+that performs unusually poorly.  It is easy to spot these networks by
+their performance either as measured as part of the training process
+or on the test audio file, perhaps taking the best of a few training
+runs.
 
@@
     \item Clone the Github repository onto your computer.
-    \item You also need the following packages from MATLAB File Exchange:
+    \item You also need the following packages from MATLAB File Exchange somewhere in your MATLAB path:
       \begin{description}
@@
       MATLAB file containing aligned song, nonsong (calls, cage noise,
-      etc), and the sampling frequency.
+      etc), and the audio sampling frequency.
     \item The configuration file {\tt \textless data
@@
     \item[Usage:]\hfill
-      Use LabVIEW to open {\tt nnfft.vi}.  A file
-      chooser is provided in the front panel.  Use it to point to the network file
+      \begin{itemize}
+      \item Use LabVIEW to open {\tt nnfft.vi}.
+        \item A file chooser is provided in the front panel of {\tt nnfft.vi}.  Use it to point to the network file
       produced by {\tt learn\_detector.m}, and use LabView's ``Run''
       button to begin detection.
+      \end{itemize}
     \end{description}
@@
 %\bibliography{birds}
-\bibliography{birds}
+\begin{thebibliography}{10}
+
+\bibitem{Kao2005}
+Kao MH, Doupe AJ, Brainard MS.
+\newblock Contributions of an avian basal ganglia--forebrain circuit to
+  real-time modulation of song.
+\newblock Letters to Nature. 2005 February;433:638--643.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v433/n7026/abs/nature03127.html}.
+
+\bibitem{Wang2008interhemispheric}
+Wang CZH, Herbst JA, Keller GB, Hahnloser RHR.
+\newblock Rapid Interhemispheric Switching during Vocal Production in a
+  Songbird.
+\newblock PLOS Biology. 2008 October;6(10).
+\newblock Available from:
+  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.0060250}.
+
+\bibitem{Keller2009}
+Keller GB, Hahnloser RHR.
+\newblock Neural processing of auditory feedback during vocal practice in a
+  songbird.
+\newblock Nature. 2009 January;457:187--190.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v457/n7226/abs/nature07467.html}.
+
+\bibitem{Tumer2007pitchshift}
+Tumer EC, Brainard MS.
+\newblock Performance variability enables adaptive plasticity of 'crystallized'
+  adult birdsong.
+\newblock Nature. 2007 December;450:1240--1244.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v450/n7173/abs/nature06390.html}.
+
+\bibitem{Andalman2009}
+Andalman AS, Fee MS.
+\newblock A basal ganglia-forebrain circuit in the songbird biases motor output
+  to avoid vocal errors.
+\newblock Proceedings of the National Academy of Sciences of the United States
+  of America. 2009 July;106(30):12518--12523.
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2709669/}.
+
+\bibitem{Charlesworth2011learning}
+Charlesworth JD, Tumer EC, Warren TL, Brainard MS.
+\newblock Learning the microstructure of successful behavior.
+\newblock Nature Neuroscience. 2011;14(3):373--380.
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3045469/}.
+
+\bibitem{Warren2011}
+Warren TL, Tumer EC, Charlesworth JD, Brainard MS.
+\newblock Mechanisms and time course of vocal learning and consolidation in the
+  adult songbird.
+\newblock Journal of Neurophysiology. 2011 October;106(4):1806--1821.
+\newblock Available from:
+  \url{http://jn.physiology.org/content/106/4/1806.full}.
+
+\bibitem{Canopoli2014}
+Canopoli A, Herbst JA, Hahnloser RHR.
+\newblock A Higher Sensory Brain Region Is Involved in Reversing
+  Reinforcement-Induced Vocal Changes in a Songbird.
+\newblock The Journal of Neuroscience. 2014;34(20):7018--7026.
+\newblock Available from:
+  \url{http://www.jneurosci.org/content/34/20/7018.full}.
+
+\bibitem{Leonardo1999}
+Leonardo A, Konishi M.
+\newblock Decrystallization of adult birdsong by perturbation of auditory
+  feedback.
+\newblock Nature. 1999 June;399:466--470.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v399/n6735/abs/399466a0.html}.
+
+\bibitem{Skocik2013}
+Skocik M, Kozhevnikov A.
+\newblock Real-time system for studies of the effects of acoustic feedback on
+  animal vocalizations.
+\newblock Frontiers in Neural Circuits. 2013 January;6(111).
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3539774/}.
+
+\bibitem{Poole2012}
+Poole B, Markowitz JE, Gardner TJ.
+\newblock The Song Must Go On: Resilience of the Songbird Vocal Motor Pathway.
+\newblock PLOS One. 2012 June;7(6).
+\newblock Available from:
+  \url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0038173}.
+
+\bibitem{Jacobsen2003SlidingDFT}
+Jacobsen E, Lyons R.
+\newblock The sliding DFT.
+\newblock IEEE Signal Processing Magazine. 2003 Mar;20(2):74--80.
+\newblock Available from:
+  \url{http://www.cmlab.csie.ntu.edu.tw/DSPCourse/reference/Sliding\%20DFT.pdf}.
+
+\bibitem{hkp}
+Hertz JA, Krogh AS, Palmer RG.
+\newblock Introduction to the Theory of Neural Computation.
+\newblock Perseus Books; 1991.
+
+\bibitem{Noll1970pitchdetection}
+Noll AM.
+\newblock Pitch determination of human speech by the harmonic product spectrum,
+  the harmonic sum spectrum, and a maximum likelihood estimate.
+\newblock In: Proceedings of the symposium on computer processing in
+  communications. vol.~19; 1970. p. 779--797.
+
+\end{thebibliography}
 
 \end{document}
-
-s

### 2016-09-07 rebuttal.txt
New file: show differences between PLOS CompBio submission and this one.
- Made using latexdiff, and then some errors fixed.

- Also include PDF.

@@
 
-> Charlesworth et al. (“Learning the microstructure of
-> successful behavior”) showed that learning after white noise targeting
-> is temporally localized at less than 10 msec timescale. Furthermore,
-> in my own and other labs, the key predictor of whether online
-> targeting will successfully drive learning is the standard deviation
-> of detections/triggers relative to syllable onset, with syllable onset
+> Charlesworth et al. (“Learning the microstructure of successful
+> behavior”) showed that learning after white noise targeting is
+> temporally localized at less than 10 msec timescale. Furthermore, in
+> my own and other labs, the key predictor of whether online targeting
+> will successfully drive learning is the standard deviation of
+> detections/triggers relative to syllable onset, with syllable onset
 > defined by some measure of sound amplitude for each individual
@@
 accuracy figures including lots of cage noise samples (about half an
-hour's worth).  It is slightly different from the previous version of
-Fig. 3, since the introduction of cage noise decreases the ratio of
-positive/negative examples and we have not adjusted the false positive
-vs. false negative cost parameter (which was arbitrarily set at 1).
-Consequently correct detections and incorrect detections have both
+hour's worth).  Correct detections and incorrect detections have both
 somewhat decreased from our earlier results.
@@
 
-Big data sets!  But we changed the training algorithm and now things
-are about 10 times faster.
+Big data sets, and we made little effort to optimise offline
+(training) speed!  But we changed the training algorithm and now
+things are about 10 times faster.
 
@@
 
-Arbitrarily, as noted in the text.
+Arbitrarily.  Now noted in the text.
 

### 2016-09-07 main.tex
Take out last (?) reference to Fibre Photometry experiment.

@@
 \textsuperscript{b}~contributed the Swift implementation and its description herein, helped debug the training code, and measured the timing for most of the detectors.
-\textsuperscript{c}~performed the fibre photometry experiment, wrote several types of filter-bank software against which we compared, made suggestions for the neural network detector, and reformatted the MATLAB code for distribution.
+\textsuperscript{c}~wrote several types of filter-bank software against which we compared, made suggestions for the neural network detector, and reformatted the MATLAB code for distribution.
 \textsuperscript{d}~provided a stream of suggested techniques, metrics, and goals.
@@
 instance of the target syllable to the firing of the detector's TTL
-pulse.  For example, if the recognition region is 30 ms long and the
-target trigger point is at 200 ms with respect to the corpus of
-aligned songs, then when the detector has seen the region from
-170--200 ms, it should recognise the syllable.  If for a given
-presentation of the song the detector fires at 203 ms, we take the
-latency on that trial to be 3 ms.
+pulse.  For example, if the target trigger point is at 200 ms with
+respect to the corpus of aligned songs, and if the detection region is
+30 ms long, then when the detector has seen the region from 170--200
+ms, it should recognise the syllable.  If for a given presentation of
+the song the detector fires at 203 ms, we take the latency on that
+trial to be 3 ms.
 
@@
 that the detector will see---only the non-song data may be different.
+%Accuracy \marginpar{Should be!
+Verify!
+And/or explain that there is a bug in the LabVIEW implementation, if there in fact is.
+ALSO need to present a picture of the test syllable or etc \dots} is identical across the three detectors.
+\noprint{
+  Typical confusion matrices for these syllables are:
+
+  \vspace{8pt}\par\noindent
+  \begin{tabular}{r|cc}
+    $t^*_1$
+    & \multicolumn{2}{c}{True} \\ 
+    & pos & neg \\ 
+    \hline  Detected pos & 99.96451\% & 0.000298\%\\ 
+    neg & 0.0355\% & 99.99970\%\\ 
+  \end{tabular}
+  \vspace{8pt}\par\noindent
+  \begin{tabular}{r|cc}
+    $t^*_2$ & \multicolumn{2}{c}{True} \\ 
+    & pos & neg \\ 
+    \hline  Detected pos & 99.92903\% & 0.000596\%\\ 
+    neg & 0.0710\% & 99.99940\%\\ 
+  \end{tabular}
+  \vspace{8pt}\par\noindent
+  \begin{tabular}{r|cc}
+    $t^*_3$ & \multicolumn{2}{c}{True} \\ 
+    & pos & neg \\ 
+    \hline  Detected pos & 99.68062\% & 0.00194\%\\ 
+    neg & 0.319\% & 99.99806\%\\ 
+  \end{tabular}
+  \vspace{8pt}\par\noindent
+  \begin{tabular}{r|cc}
+    $t^*_4$ & \multicolumn{2}{c}{True} \\ 
+    & pos & neg \\ 
+    \hline  Detected pos & 99.82257\% & 0.00239\%\\ 
+    neg & 0.177\% & 99.99761\%\\ 
+  \end{tabular}
+  \vspace{8pt}\par\noindent
+  \begin{tabular}{r|cc}
+    $t^*_5$ & \multicolumn{2}{c}{True} \\ 
+    & pos & neg \\ 
+    \hline  Detected pos & 99.78708\% & 0\\ 
+    neg & 0.213\% & 100\%\\ 
+  \end{tabular}
+  \vspace{8pt}\par\noindent
+  \begin{tabular}{r|cc}
+    $t^*_6$ & \multicolumn{2}{c}{True} \\ 
+    & pos & neg \\ 
+    \hline  Detected pos & 99.71611\% & 0.00552\%\\ 
+    neg & 0.284\% & 99.99448\%\\ 
+  \end{tabular}
+  \vspace{8pt}\par\noindent
+}
 

### 2016-09-08 rebuttal.txt
Slightly nicer to reviewer 3.
Suggestions for editors.

@@
 We try to provide a complete description of the method, but since we
-are also distributing our implementation, this makes sense.  Therefore
-we have added installation/configuration instructions in Appendix A.
+are also distributing our implementation, it seems like a great idea
+to do as you suggest.  Therefore we have added
+installation/configuration instructions in Appendix A.
 
@@
 
-We had already characterised the timing from the completion of the
-target recognition region to the firing of the TTL pulse--this is what
-we define as "Jitter" (line 27, Section Quantification.Timing
-(beginning around line 342), and also extensively throughout the
-document).  We have tried to clean up those references and clarify
-our description.
+We had extensively characterised the timing from the completion of the
+target recognition region to the firing of the TTL pulse.  We have
+tried to clean up those references and clarify our description: if the
+detection point is at the onset of a syllable, then our results apply
+(see the paragraph at line 361 in the manuscript, or the paragraph at
+line 375 in the diff document).
 
@@
 We believe that we have provided a sufficient analysis of jitter, and
-that our results are good enough to obviate the improvement you
-suggest.
+that, while the improvement you suggest could only decrease our jitter
+measurement somewhat, the performance without that enhancement is
+almost as good as it can be.
 
@@
 
-Actually, we already had done so: see Fig. 5 (and section 3.2.syllable
+Actually, we had already done so: see Fig. 5 (and section 3.2.syllable
 choice).
@@
 > the writing and presentation could be improved.
-
-Always, but that is too vague to address.
-
 > e.g. fig 1 increase font size of labels
@@
 
-This directly contradicts your request that we add cage noise,
-flapping, etc.  Luckily, training time shouldn't matter--it's online
-performance that counts, right?  However, the change to scaled
-conjugate gradient training has sped things up considerably.
+To a first approximation, this contradicts your request that we add
+cage noise, flapping, etc.  Luckily, training time shouldn't
+matter--it's online performance that counts, right?  However, the
+change to scaled conjugate gradient training has sped things up
+considerably.
 
@@
 
-Arbitrarily.  Now noted in the text.
+Arbitrarily.  Now noted in the text.  Thanks.
 
@@
 
-My comment applied to either one, but that text (it just described
-something we didn't do) was unnecessary, so I removed it.
+My comment applied to either one, but that text, which just described
+something we didn't do, was unnecessary, so I removed it.
 
@@
 
-No loads from users (i.e. operating system processes only).
+No loads from users (i.e. operating system processes only).  Clarified
+in the text.
 
@@
 > constrained method.
-
-Yup.
-
+>
 > In the present manuscript, it has been obtained, presumably, by
@@
 doesn't really matter very much--hopefully any regression tool would
-work similarly).  But it did turn out in the end that Scaled Conjugate
-Gradient produced similar results faster.
+work roughly similarly!).  But it did turn out in the end that Scaled
+Conjugate Gradient produced similar results faster.
 

### 2016-09-08 main.tex
Fix minor typos
Slightly nicer to reviewer 3.
Suggestions for editors.

@@
 homologue of mammalian basal ganglia consisting of a few distinct
-brain areas concerned with the production and learning of song.  For
+brain areas concerned with the learning and production of song.  For
 example, \cite{Kao2005} stimulated the lateral magnocellular nucleus
@@
 \item Fast: sub-millisecond latencies, with jitter around 2 ms.
-\item Accurate: false positive rates under 0.001\% and false negative rates under 1\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
+\item Accurate: false negative rates under 1\% and false positive rates under 0.001\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
 \item State-of-the-art performance with default parameters.
@@
 systems' default schedulers and process priorities, running typical
-operating system daemon processes but no user loads.  The computers
-had ample memory resources.
+operating system daemon processes but no loads from user processes.
+The computers had ample memory resources.
 
@@
       detected). Once you select an input and an output, a new window will show all audio 
-      channels associated with the input. For each channel, select a the text-encoded version of 
+      channels associated with the input. For each channel, select the text-encoded version of 
       the detector created in the previous step. Once configured, press run to begin 
@@
       the raw input or output, look at the ``Processor.swift'' file. The 
-      {\tt Processor} class acts as a coordinator, receiving income audio, passing that
+      {\tt Processor} class acts as a coordinator, receiving incoming audio, passing that
       to the detector and generating appropriate output TTL signals.

### 2016-09-09 rebuttal.txt
Incorporate comments from Sanne on making everything more kind to the foolish reviewers.
I did not update the diff doc yet.

@@
 especially Figs. 4-7.  Since this is well within the 10-ms window
-suggested by the literature and indeed barely above the theoretically
-optimal jitter given our detection parameters, we decided that the
-improvement that could be achieved by micro-realigning song was
-unimportant.
+suggested by the literature, we decided that the improvement that
+could be achieved by micro-realigning song was unimportant.  Indeed,
+our measured jitter is barely above the theoretically optimal jitter
+given our detection parameters, so the average micro-realignment error
+cannot be significant.
 
@@
 that, while the improvement you suggest could only decrease our jitter
-measurement somewhat, the performance without that enhancement is
-almost as good as it can be.
+measurement, the performance without that enhancement is good enough
+that the improvement we might see is not very important.
 
@@
 
-Actually, we had already done so: see Fig. 5 (and section 3.2.syllable
-choice).
+Actually, we had already done so: see Fig. 5 (and section 3.2,
+"syllable choice").
 
@@
 does not have any 50-ms periodicity), we just added a comment that
-these results are indeed representative.
+these results are indeed representative (e.g. the beginning of Section
+2.3).
 
@@
 
-It would.  But we don't have a convenient data set, and we feel that
-it is more important to get the technique out there quickly than to
-continue to characterise it in yet more detail.
+It would.  But we don't currently have data sets for other species,
+and we feel that it is more important to get the technique out there
+quickly than to continue to characterise it in more detail.
 
@@
 
-Yes.  See (1b) above.
+We discuss one song in the paper, mostly for convenience, but we've
+added results are representative of what we see on many birds.  See
+(1b) above.
 
-> Most
-> importantly, they do not do a life bird test.
+> Most importantly, they do not do a life bird test.
 
@@
 bird test (and ongoing research in our lab uses the detector with
-great success), but of course all quantitative tests must be done on
-recordings in order to systematically analyse accuracies of various
-detectors.
+great success), but all quantitative tests must be done on recordings
+in order to systematically analyse accuracies of various detectors.
 
@@
 
-That is an excellent point!  We have remedied this.  See the paragraph
-around line 160.  Also, the new Results/Accuracy graph (Fig. 3) shows
-accuracy figures including lots of cage noise samples (about half an
-hour's worth).  Correct detections and incorrect detections have both
-somewhat decreased from our earlier results.
+That is an excellent point!  We have remedied this.  See lines 94, 163
+and 322 in the diff document.  Also, the new Results/Accuracy graph
+(Fig. 3) shows accuracy figures including lots of cage noise samples
+(about half an hour's worth, as now described).  Correct detections
+and incorrect detections have both somewhat decreased from our earlier
+results.
 
@@
 
-See (1c) above.
+We've removed it.  See (1c) above.
 
@@
 There is an infinite space of possible errors in the training set, so
-this question is impossible to answer.  However, no effort was made to
-compensate for errors in the alignment tool, and the results aren't
-bad, so I guess it depends on what kinds of errors you're worried
-about.
+this question is impossible to answer well.  However, no effort was
+made to compensate for errors in the alignment tool, and the results
+aren't bad, so I guess it depends on what kinds of errors you're
+worried about.
 
@@
 
-Thanks.  Done.
+Thanks.  We've increased the font size and improved the writing.
 
@@
 Big data sets, and we made little effort to optimise offline
-(training) speed!  But we changed the training algorithm and now
+(training) speed.  But we changed the training algorithm and now
 things are about 10 times faster.
@@
 
-To a first approximation, this contradicts your request that we add
-cage noise, flapping, etc.  Luckily, training time shouldn't
-matter--it's online performance that counts, right?  However, the
-change to scaled conjugate gradient training has sped things up
-considerably.
+We preferred your other suggestion--that we increase the negative
+targets (adding cage noise, etc).  Training speed shouldn't matter
+much compared to online performance, but the change to scaled
+conjugate gradient training has sped things up considerably.
 
@@
 
-We documented robustness to many choices.  There are always more, but
+We documented robustness to some choices.  There are always more, but
 this was supposed to be a 5-page paper.  We have to stop somewhere.
@@
 
-Agreed.  But we feel it is useful to the community (as do the other
-two reviewers, as well as several of our colleagues), and various
-researchers have requested that we publish it.
+Although there is no machine learning innovation, we believe that the
+combination of making available a useful tool and fairly thoroughly
+documenting its performance constitute a significant contribution to
+the field.
 
@@
 We agree with you, although several have noted that this model is
-extremely powerful and the absence of tools such as this one is a
-barrier to entry into the field.
+extremely powerful and the absence of readily available tools such as
+this one is a barrier to entry into precisely timed zebra finch
+experiments.
 
@@
 
-Yes, more birds/species would always be better, but see (1b).
+Yes, more birds/species would always be better, but see (1b).  The
+characteristic variation in song between different 'strains' appears
+to be much less than the variation in song between different birds
+within the one strain that we have tested on, or even within the
+different moments in the song of one bird that we have used as our
+example.
 
@@
 
-((3a)) This is incorrect.  We cited the article in which alignment
-software is described, and we linked to the alignment software.  The
-only step that we have not described is using a recording device,
-which is outside the scope of the paper.
+This is incorrect.  We cited the article in which alignment software
+is described, and we linked to the alignment software.  The only step
+that we have not described is using a recording device, which is
+outside the scope of the paper.  However, if you are curious, there is
+a nice tutorial at
+http://soundanalysispro.com/manual-1/chapter-2-installation-of-sap2/hardware-installation-building-training-boxes
 
@@
 > dataset is extremely difficult to obtain without very invasive and
-> constrained method.
->
-> In the present manuscript, it has been obtained, presumably, by
-> letting a bird to its own device in a small cage for hours to
-> extract songs. Many experiments do not want such harsh treatment.
+> constrained method.  In the present manuscript, it has been
+> obtained, presumably, by letting a bird to its own device in a small
+> cage for hours to extract songs. Many experiments do not want such
+> harsh treatment.
 
-Without isolating a bird, isolating a song requires an entirely
-different set of techniques, which would also make for interesting
-research.  It sounds like we proposed a solution to Problem A, but you
-wish we'd solved Problem B instead.
+While learning to detect syllables of specific birds using samples
+obtained in the wild would be an interesting research problem, it is
+outside the scope of the current paper.  Instead, we only claim to
+address a need encountered in a great deal of zebra finch research,
+using common protocols.
 
@@
 
-Good point.  Now specified.
+Good point.  Now specified (undirected) in Section 2.1.
 
 > e) It is also assume that the user has means to extract / align song
-> automatically event though this is never mentioned in the manuscript.
+> automatically event though this is never mentioned in the
+> manuscript.
 
-Wrong.  See (3a).
+Is this the same point as your (c)?
 
@@
 
-Right.  See (1c).
+We've removed it (see (1c)).
 
@@
 
-> in short, all these criticisms in fact converge to the main weakness:
-> this paper/software is directed to a very specific community both in
-> scope and type of data needed for it to work and the authors should
-> aim to a journal directed at this community.
+> in short, all these criticisms in fact converge to the main
+> weakness: this paper/software is directed to a very specific
+> community both in scope and type of data needed for it to work and
+> the authors should aim to a journal directed at this community.
 
-Yes.
+We believe that the paper presents a valuable technique and a far more
+thorough characterisation of finch syllable detection accuracy and
+timing than has previously been available.  It is our hope that the
+performance, ease of use, simplicity, and low hardware cost of this
+method will speed the development of a small but growing field.

### 2016-09-14 rebuttal.txt
Clean up rebuttal.
Add reference describing how to record birds.  Sigh.

Sanne's suggestions for editor.

Nathan's new Tag for the Swift detector.

New document for suggested reviewers.

@@
 There is an infinite space of possible errors in the training set, so
-this question is impossible to answer well.  However, no effort was
-made to compensate for errors in the alignment tool, and the results
-aren't bad, so I guess it depends on what kinds of errors you're
-worried about.
+this question is impossible to answer thoroughly.  However, no effort
+was made to compensate for errors in the alignment tool, and the
+results aren't bad, so I guess it depends on what kinds of errors
+you're worried about.
 
@@
 Yes, more birds/species would always be better, but see (1b).  The
-characteristic variation in song between different 'strains' appears
-to be much less than the variation in song between different birds
-within the one strain that we have tested on, or even within the
-different moments in the song of one bird that we have used as our
-example.
+characteristic variation in song between different 'strains' is much
+less than the variation in song between different birds within the one
+strain that we have tested on, or even within the different moments in
+the song of one bird that we have used as our example.
 
@@
 Some insight added, but we agree that it's a tool for a specific
-community.  As is any tool.
+community.  (As is any tool.)
 
@@
 
-This is incorrect.  We cited the article in which alignment software
-is described, and we linked to the alignment software.  The only step
-that we have not described is using a recording device, which is
-outside the scope of the paper.  However, if you are curious, there is
-a nice tutorial at
-http://soundanalysispro.com/manual-1/chapter-2-installation-of-sap2/hardware-installation-building-training-boxes
+We cited the article in which alignment software is described, and we
+linked to the alignment software.  The only step that we had not
+described is using a recording device, which is outside the scope of
+the paper.  However, I've added a reference to a description of a
+common technique
+(http://soundanalysispro.com/manual-1/chapter-2-installation-of-sap2/hardware-installation-building-training-boxes).
 
@@
 outside the scope of the current paper.  Instead, we only claim to
-address a need encountered in a great deal of zebra finch research,
-using common protocols.
+address a need often encountered in zebra finch research using common
+protocols.
 
@@
 doesn't really matter very much--hopefully any regression tool would
-work roughly similarly!).  But it did turn out in the end that Scaled
+work roughly similarly).  But it did turn out in the end that Scaled
 Conjugate Gradient produced similar results faster.
@@
 performance, ease of use, simplicity, and low hardware cost of this
-method will speed the development of a small but growing field.
+method will significantly lower the barrier to entry to a small but
+growing field.

### 2016-09-14 main.tex
Clean up rebuttal.
Add reference describing how to record birds.  Sigh.

Sanne's suggestions for editor.

Nathan's new Tag for the Swift detector.

New document for suggested reviewers.

@@
 
-We begin with a few hundred recordings of a given bird's song, as well as calls, cage noise, and other non-song audio data.  Male
-zebra finch song is most highly stereotyped when a female is present
-(``directed song'') and slightly more variable otherwise
-(``undirected''); we train and test on undirected song since this is
-more commonly studied, as well as more challenging.  Recordings were made inside a sound-isolating
-chamber in which was mounted a recording microphone (Audio-Technica
-AT803).  The songs were time-aligned as described in \cite{Poole2012}.
+We begin with a few hundred recordings of a given bird's song, as well
+as calls, cage noise, and other non-song audio data.  Male zebra finch
+song is most highly stereotyped when a female is present (``directed
+song'') and slightly more variable otherwise (``undirected''); we
+train and test on undirected song since this is more commonly studied,
+as well as more challenging.  Recordings were made inside a
+sound-isolating chamber in which was mounted a recording microphone
+(Audio-Technica AT803), using methods similar to those described in \cite{Tchernichovski2011SAP}, Chapter 2.  The songs were
+time-aligned as described in \cite{Poole2012}.
 
@@
     \end{description}
-  \item[Swift:] {\tt https://github.com/gardner-lab/syllable-detector-swift}
+  \item[Swift:] {\tt https://github.com/gardner-lab/\dots\ syllable-detector-swift/releases/tag/0.3.1}
     \begin{description}
@@
 
+\bibitem{Tchernichovski2011SAP}
+Tchernichovski O. Sound Analysis Pro User Manual; 2011.
+\newblock Available from: \url{http://soundanalysispro.com/manual-1}.
+
 \bibitem{Jacobsen2003SlidingDFT}

### 2016-11-02 main.tex
Actually submitted this version.
Some changes requested by the editor... not done yet.

@@
 of 2--6 syllables, with longer songs taking on the order of a
-second. The song may be repeated hundreds of times per day, is almost
-identical each time, and several brain areas reflect this consistency
-in highly stereotyped neural firing patterns. This consistency makes
+second. The song may be repeated hundreds of times per day, and is almost
+identical each time.  Several brain areas reflect this consistency
+in highly stereotyped neural firing patterns, which makes
 the zebra finch one of the most popular models for the study of the
@@
   \includegraphics[width=\textwidth]{Fig1}
-  \caption{This image was made by superposing the spectra of our 2818
-    aligned songs.  Our example detection points, $t^*_1\ldots t^*_6$,
-    are shown as red lines, with the recognition regions (30 ms
-    $\times$ 1--8 kHz) marked as rectangles.}
+  \caption{The spectrogram of a typical song, used as the example
+    throughout this paper.  This image was made by superposing the
+    spectra of our 2818 aligned songs.  Our example detection points,
+    $t^*_1\ldots t^*_6$, are shown as red lines, with the recognition
+    regions (30 ms $\times$ 1--8 kHz) marked as rectangles.}
   \label{fig:song}
@@
 
-Because the data set consists of temporally aligned songs, ground
-truth is available for each song (albeit with minor variations due to
-the alignment process used \cite{Poole2012}), and thus the detector
-can be checked by presenting the recorded training songs as well as
-the canonical detection events. To this end, besides the trained
+Because the data set consists of temporally aligned songs, the moment at which detection should occur during each song sample is available (albeit with minor variations due to
+the alignment process used \cite{Poole2012}).  The detector
+can be checked by comparing its response to the recorded training song samples against
+these ``canonical'' detection events.  To this end, besides the trained
 network object, our learning code produces an audio file consisting of
@@
 
-Mean latency is then given as the average time between the canonical
-signal and the corresponding detection event.  It is a helpful number,
+We define latency as the time between the training target (during our tests, this is indicated by the canonical
+signal in the test file) and the corresponding detection event.  It is a helpful number,
 but not a critical one, since in zebra finch a detector with high but
@@
 before the true moment of interest (for example, if triggering should
-occur at the very beginning of a syllable, the detector's time window
-may include previous syllable).  Often more important is the
-variability in the latency.  We define jitter as the standard
-deviation of the latency.
+occur at the very beginning of a syllable, the detector may be trained to recognise the previous syllable and the gap thereafter).
+Usually the variability in latency is the more important number.
+We define jitter as the standard deviation of detection latency over the test songs.
 
@@
 The first two triggers, $t^*_1$ at 150 ms and $t^*_2$ at 200 ms, are
-far from pure tones, but are closer to coloured noise.  They would be
+far from pure tones, but are closer to bandpass-filtered white noise.  They would be
 difficult for a harmonic-stack filter bank to detect, and especially

### 2017-01-11 main.tex
Add birds B, C, D to Quantification section. Add dataset sizes in a table.

@@
 
-We have used this detector for song alignment and feedback experiments
-in 5 behaving birds, and for offline analysis on a few others.  For
-simplicity, and to test robustness to imperfect choice of ``syllable'', we present performance on 6 randomly chosen points in one
-bird's song.  The results presented here are representative of
-real-world performance over many birds in our lab.
-
-We divided the data into training and test sets.  For the results
-reported in the following section, the training set consisted of 1000
-instances of the song of one bird, and 2000 song-length samples of
-non-song (cage noise, calls, other birds' song).  Our test set
-consisted of an additional 1818 iterations of the bird's song and 3636
-song-length samples of non-song.
+We measure accuracy and characterise timing for 6 randomly chosen
+points in one bird's song (Bird A), and present accuracy results for
+three more birds (B, C, D) (timing does not appreciably vary among
+birds). The results presented here are representative of real-world
+performance in feedback experiments on many birds in our lab.
+
+% A:
+% B: LNY46 0.28s
+% C: LNY42 0.38s
+% D: LNY4RB 0.25s
+
+Data consisted of extracted, aligned songs, and additional song-length samples of non-song (cage noise, calls, other birds' song).
+These were divided into training and test sets.
+The sizes of our data sets are given in Table~\ref{table:datasizes}.
+
+\begin{table}
+  \begin{tabular}{c|cccc}
+    Bird & Training songs & Test songs & Training non-song & Test non-song \\ \hline
+    A & 1000 & 1818 & 1000 & 3636 \\
+    B & 1000 & 1533 & 158 & 242 \\
+    C & 1000 & 1245 & 178 & 222 \\
+    D & 1000 & 1614 & 153 & 247
+  \end{tabular}
+  \caption{Sizes of our training and test sets.}
+  \label{table:datasizes}
+\end{table}
 
@@
 
-Because the data set consists of temporally aligned songs, the moment at which detection should occur during each song sample is available (albeit with minor variations due to
-the alignment process used \cite{Poole2012}).  The detector
-can be checked by comparing its response to the recorded training song samples against
-these ``canonical'' detection events.  To this end, besides the trained
+Because the data set consists of temporally aligned songs, the moment
+at which detection should occur during each song sample is available
+(albeit with minor variations due to the alignment process used
+\cite{Poole2012}).  The detector can be checked by comparing its
+response to the recorded training song samples against these
+``canonical'' detection events.  To this end, besides the trained
 network object, our learning code produces an audio file consisting of

### 2017-01-24 main.tex
Some changes for editorial comment #2 (compensate for their citation style's failings).

@@
 
-This consistency allows a variety of experiments if precise moments in
-song can reliably be detected quickly enough to trigger other
-apparatus during singing.  A common area of study with this
-song-triggering technique is the anterior forebrain pathway (AFP), a
-homologue of mammalian basal ganglia consisting of a few distinct
-brain areas concerned with the learning and production of song.  For
-example, \cite{Kao2005} stimulated the lateral magnocellular nucleus
+If precise moments in song can reliably be detected quickly enough to
+trigger other apparatus during singing, then this consistency of
+behaviour allows a variety of experiments.  A common area of study
+with this song-triggering technique is the anterior forebrain pathway
+(AFP), a homologue of mammalian basal ganglia consisting of a few
+distinct brain areas concerned with the learning and production of
+song.  For example, stimulation of the lateral magnocellular nucleus
 of the anterior nidopallium (LMAN)---the output nucleus of the
-AFP---at precisely timed moments during song and showed that this area
-repeatably controls specific variables in song output.
-\cite{Wang2008interhemispheric} stimulated LMAN and the high vocal
-centre (HVC) in one hemisphere or the other, and found that control of
-song rapidly switches between hemispheres.  \cite{Keller2009}
-investigated the response of cells in field L and the caudolateral
-mesopallium during song with and without feedback perturbations as
-well as external sounds, and concluded that these regions may hold a
-representation of song against which auditory signals are compared.
-\cite{Tumer2007pitchshift} played a disruptive sound during renditions
-of a syllable that were slightly above (or below) its average pitch,
-and showed that the presence of this apparently random natural
+AFP---at precisely timed moments during song shows that this area
+controls specific variables in song output \cite{Kao2005}.
+Song-synchronised stimulation of LMAN and the high vocal centre (HVC)
+in one hemisphere or the other showed that control of song rapidly
+switches between hemispheres \cite{Wang2008interhemispheric}.
+Feedback experiments have shown that Field L and the caudolateral
+mesopallium may hold a representation of song against which auditory
+signals are compared \cite{Keller2009}.  Renditions of a syllable that
+were disrupted by white noise when they were slightly above (or below)
+its average pitch showed that the apparently random natural
 variability in songbird motor output is used to drive change in the
-song.  \cite{Andalman2009} extended this experiment and showed that
-AFP produces a corrective signal to bias song away from those
-disruptions.  \cite{Charlesworth2011learning} showed that the song
-change was isolated to within roughly 10 milliseconds (ms) of the
-stimulus, and further proposed an elegant model that predicts the
-shape of the learned response and suggests a simple mechanism.
+song \cite{Tumer2007pitchshift}. Furthermore, AFP produces a
+corrective signal to bias song away from those disruptions
+\cite{Andalman2009}. The
+song change is isolated to within roughly 10 milliseconds (ms) of the
+stimulus , and further proposed an elegant model that predicts the
+shape of the learned response and suggests a simple mechanism \cite{Charlesworth2011learning}.
+
 \cite{Warren2011} showed that that AFP transfers the error signal to
@@
 Despite the power and versatility of vocal feedback experiments, there
-is no standard syllable detector.  Desiderata include:
+is no standard syllable detector.  Desiderata for such a detector include:
 \begin{description}
@@
 \begin{enumerate}
-\item Record and align a corpus of training songs.  The technique has been published in \cite{Poole2012}.
+\item Record and align a corpus of training songs.
+The technique has been published in \cite{Poole2012}.
+200 songs generally yield acceptable results, but here we standardise on 1000-song training sets.
 \item Choose one or more instants in the song that should create a trigger event, and train a
-  neural network to recognise them. This step is carried out offline.  While any neural network software would produce equivalent results, we used MATLAB 2015b's neural network toolbox.
+  neural network to recognise them. This step is carried out offline.  While any neural network software would produce similar results, we used MATLAB 2015b's neural network toolbox.
 \item Once trained and saved, the neural network is used by a realtime
-  detection program that listens to an audio signal and indicates detection of the target
+  detection programme that listens to an audio signal and indicates detection of the target
   syllables via a TTL pulse.  We present three implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
@@
 \item Fast: sub-millisecond latencies, with jitter around 2 ms.
-\item Accurate: false negative rates under 1\% and false positive rates under 0.001\% for a variety of syllables.  Balancing these rates depends on a single relative-cost parameter.
+\item Accurate: false negative rates under 1\% and false positive rates under 0.001\% for a variety of syllables.
 \item State-of-the-art performance with default parameters.
+\item Requires almost no programming experience.
 \item Runs on inexpensive hardware.
@@
 processor at 3.7GHz (a relatively low-end machine) with 32 gigabytes
-of RAM, running Microsoft Windows 8.1 Pro.
-
-This implementation has several drawbacks: it requires Windows and
-expensive hardware and software and due to the programming language it
+of RAM, running Microsoft Windows 8.1 Pro and Windows 10.
+
+This implementation has several drawbacks: it requires expensive hardware and software from National Instruments (a data acquisition card and LabVIEW), Windows (we use hardware features of LabVIEW that are unavailable on MacOS or Linux), and due to the programming language it
 is difficult to modify and debug---indeed, a persistent bug in our
@@
 the other detector implementations on some syllables. However, our
-test configuration usually achieved excellent performance, and further
+test configuration proved itself capable of excellent performance, and further
 gains should be possible if the implementation were retargeted onto
@@
 % A:
-% B: LNY46 0.28s
+% B: LNY46 0.28s @ 140ms
 % C: LNY42 0.38s
@@
     A & 1000 & 1818 & 1000 & 3636 \\
-    B & 1000 & 1533 & 158 & 242 \\
-    C & 1000 & 1245 & 178 & 222 \\
-    D & 1000 & 1614 & 153 & 247
+    B & 1000 & 1530 & 1428 & 2185 \\
+    C & 1000 & 1245 & 990 & 1233 \\
+    D & 1000 & 1614 &  & 247
   \end{tabular}

### 2017-01-26 main.tex
Further edits to move references to end of sentences

@@
 were disrupted by white noise when they were slightly above (or below)
-its average pitch showed that the apparently random natural
+the syllable's average pitch showed that the apparently random natural
 variability in songbird motor output is used to drive change in the
@@
 song change is isolated to within roughly 10 milliseconds (ms) of the
-stimulus , and further proposed an elegant model that predicts the
-shape of the learned response and suggests a simple mechanism \cite{Charlesworth2011learning}.
-
-\cite{Warren2011} showed that that AFP transfers the error signal to
+stimulus, and the shape of the learned response can be predicted by
+a simple mechanism \cite{Charlesworth2011learning}.
+
+AFP transfers the error signal to
 the robust nucleus of the arcopallium (RA) using
-NMDA-receptor--mediated glutamatergic transmission. By looking at song
+NMDA-receptor--mediated glutamatergic transmission \cite{Warren2011}. By looking at song
 recovery after applying the same pitch-shift paradigm,
-\cite{Canopoli2014} showed that the caudal medial nidopallium is
+the caudal medial nidopallium is
 implicated in memorising or recalling a recent song target, but for
-neither auditory processing nor directed motor learning.
+neither auditory processing nor directed motor learning \cite{Canopoli2014}.
 
@@
 been documented or characterised in detail.  In 1999,
-\cite{Leonardo1999} used groups of IIR filters with hand-tuned logical
-operators.  Their system had a latency of 50 or 100 ms,
-and they did not report on jitter or accuracy.  As access to
+detection was achieved by a group of IIR filters with hand-tuned logical
+operators \cite{Leonardo1999}.  The system had a latency of 50 or 100 ms,
+and no reported data on jitter or accuracy.  As access to
 computational resources has improved, approaches have changed: in
-2009, \cite{Andalman2009} still used hand-tuned filters, but ran them
-on a Tucker-Davis Technologies digital signal processor.  They reported
-a latency of around 4 ms, but as with other filter-bank techniques, it
+2009, detection still used hand-tuned filters, but implemented
+on a Tucker-Davis Technologies digital signal processor \cite{Andalman2009}.  This
+achieved a latency of around 4 ms, but as with other filter-bank techniques, it
 is not strictly a syllable detector but rather a pitch and timbre
@@
 based on our experience with a similar technique, accurate.
-\cite{Keller2009} applied a neural network to a spectral image of
-song.  They reported a jitter of 4.3 ms, but further implementation and
-performance details are not available.  In 2011, \cite{Warren2011}
-matched spectral templates to stable portions of syllables in 8-ms
-segments.  They reported a jitter of 4.5 ms, and false-negative and
-false-positive rates of up to 2\% and 4\%, respectively.  Hardware
+Also in 2009, another group applied a neural network to a spectral image of
+song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
+performance details are not available.  In 2011, a detector was implemented by
+matching spectral templates to stable portions of syllables in 8 ms
+segments \cite{Warren2011}.  This detector achieved a jitter of 4.5 ms, and false-negative and
+false-positive rates of up to 2\% and 4\% respectively.  Hardware
 requirements and ease of use were not reported.  In 2013,
-\cite{Skocik2013} described in detail a system that matches spectral
-images of template syllables using a correlation coefficient.  With a
+a group described in detail a system that matches spectral
+images of template syllables using a correlation coefficient \cite{Skocik2013}.  With a
 fast desktop (Intel i7 six-core) running Linux and equipped with a

### 2017-01-26 main.tex
Minor edits related to the reviewer feedback

@@
   detection programme that listens to an audio signal and indicates detection of the target
-  syllables via a TTL pulse.  We present three implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
+  syllables via a TTL pulse.  We present three detection implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
 \end{enumerate}
@@
 \cite{Poole2012}).  The detector can be checked by comparing its
-response to the recorded training song samples against these
+response to the recorded training songs against these
 ``canonical'' detection events.  To this end, besides the trained
@@
 signal in the test file) and the corresponding detection event.  It is a helpful number,
-but not a critical one, since in zebra finch a detector with high but
+but not a critical one; due to the stereotyped song of the zebra finch, a detector with high but
 constant latency can often be trained to trigger at a point somewhat
 before the true moment of interest (for example, if triggering should
-occur at the very beginning of a syllable, the detector may be trained to recognise the previous syllable and the gap thereafter).
-Usually the variability in latency is the more important number.
-We define jitter as the standard deviation of detection latency over the test songs.
+occur at the very beginning of a syllable, the detector may be trained 
+to recognise the previous syllable and the gap thereafter).  Usually the 
+variability in latency is the more important number.  We define jitter as 
+the standard deviation of detection latency, measured over the test songs.
 

### 2017-01-27 main.tex
Clean up don't-use-citation-as-sentence-subject changes.

@@
 behaviour allows a variety of experiments.  A common area of study
-with this song-triggering technique is the anterior forebrain pathway
+with song-triggered experiments is the anterior forebrain pathway
 (AFP), a homologue of mammalian basal ganglia consisting of a few
@@
 of the anterior nidopallium (LMAN)---the output nucleus of the
-AFP---at precisely timed moments during song shows that this area
+AFP---at precisely timed moments during song showed that this area
 controls specific variables in song output \cite{Kao2005}.
@@
 corrective signal to bias song away from those disruptions
-\cite{Andalman2009}. The
-song change is isolated to within roughly 10 milliseconds (ms) of the
-stimulus, and the shape of the learned response can be predicted by
-a simple mechanism \cite{Charlesworth2011learning}.
-
-AFP transfers the error signal to
-the robust nucleus of the arcopallium (RA) using
-NMDA-receptor--mediated glutamatergic transmission \cite{Warren2011}. By looking at song
-recovery after applying the same pitch-shift paradigm,
-the caudal medial nidopallium is
-implicated in memorising or recalling a recent song target, but for
-neither auditory processing nor directed motor learning \cite{Canopoli2014}.
+\cite{Andalman2009}. The song change is isolated to within roughly 10
+milliseconds (ms) of the stimulus, and the shape of the learned
+response can be predicted by a simple mechanism
+\cite{Charlesworth2011learning}. AFP transfers the error signal to the
+robust nucleus of the arcopallium (RA) using NMDA-receptor--mediated
+glutamatergic transmission \cite{Warren2011}. The course of song
+recovery after applying such a pitch-shift paradigm showed that the
+caudal medial nidopallium is implicated in memorising or recalling a
+recent song target, but in neither auditory processing nor directed
+motor learning \cite{Canopoli2014}.
 
@@
 operators \cite{Leonardo1999}.  The system had a latency of 50 or 100 ms,
-and no reported data on jitter or accuracy.  As access to
+and jitter or accuracy were not reported.  As access to
 computational resources has improved, approaches have changed: in
-2009, detection still used hand-tuned filters, but implemented
-on a Tucker-Davis Technologies digital signal processor \cite{Andalman2009}.  This
-achieved a latency of around 4 ms, but as with other filter-bank techniques, it
+2009, hand-tuned filters were implemented on a Tucker-Davis Technologies digital signal processor, bringing latency down to around 4 ms \cite{Andalman2009}.
+But as with other filter-bank techniques, it
 is not strictly a syllable detector but rather a pitch and timbre
@@
 based on our experience with a similar technique, accurate.
-Also in 2009, another group applied a neural network to a spectral image of
+2009 also saw the application of a neural network to a spectral image of
 song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
-performance details are not available.  In 2011, a detector was implemented by
-matching spectral templates to stable portions of syllables in 8 ms
+performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8-ms
 segments \cite{Warren2011}.  This detector achieved a jitter of 4.5 ms, and false-negative and
 false-positive rates of up to 2\% and 4\% respectively.  Hardware
-requirements and ease of use were not reported.  In 2013,
-a group described in detail a system that matches spectral
-images of template syllables using a correlation coefficient \cite{Skocik2013}.  With a
+requirements and ease of use were not reported.  In 2013, spectral
+images of template syllables were compared to song using a correlation coefficient \cite{Skocik2013}.  With a
 fast desktop (Intel i7 six-core) running Linux and equipped with a
@@
 Bengalese finches respectively, measured on a small dataset.  In much
-other work, an allusion is made to a syllable detector, but no further
-information is provided.
+other work, a syllable detector is alluded to, but not described.
 

### 2017-02-15 main.tex
Response2 to the editor. A little argumentative, isn't it? Hmmm. Should fix that.
Added Nathan's birds to accuracy table. Fixed dataset size info. Minor wording changes.

Including diff.pdf

--- 2017-01-26 main.tex
@@
 stimulus, and the shape of the learned response can be predicted by
-a simple mechanism \cite{Charlesworth2011learning}.
-
-AFP transfers the error signal to
+a simple mechanism \cite{Charlesworth2011learning}. AFP transfers the error signal to
 the robust nucleus of the arcopallium (RA) using
@@
 These were divided into training and test sets.
-The sizes of our data sets are given in Table~\ref{table:datasizes}.
-
-\begin{table}
-  \begin{tabular}{c|cccc}
-    Bird & Training songs & Test songs & Training non-song & Test non-song \\ \hline
-    A & 1000 & 1818 & 1000 & 3636 \\
-    B & 1000 & 1530 & 1428 & 2185 \\
-    C & 1000 & 1245 & 990 & 1233 \\
-    D & 1000 & 1614 &  & 247
-  \end{tabular}
-  \caption{Sizes of our training and test sets.}
-  \label{table:datasizes}
-\end{table}
+For Bird A, we used 1000 songs and 1000 non-song samples for training, and the remaining 1818 songs and 3636 nonsongs for testing.
+For birds B, C, and D, we used 1000 songs and 1000 non-songs for training, and another 1000 of each for testing.
+
+%\begin{table}
+% \begin{tabular}{c|cccc}
+%    Bird & Training songs & Test songs & Training non-song & Test non-song \\ %\hline
+ %   A & 1000 & 1818 & 1000 & 3636 \\
+  %  B & 1000 & 1000 & 1000 & 1000 \\
+ %   C & 1000 & 1000 & 1000 & 1000 \\
+%    D & 1000 & 1000 &  1000 & 1000
+%  \end{tabular}
+%  \caption{Sizes of our training and test sets.}
+%  \label{table:datasizes}
+%\end{table}
 
@@
 \cite{Poole2012}).  The detector can be checked by comparing its
-response to the recorded training songs against these
-``canonical'' detection events.  To this end, besides the trained
-network object, our learning code produces an audio file consisting of
-all of the training data on the left audio channel and a delta
-function at each aligned moment of target syllable presentation on the
-right channel. Thus, when played on any audio player, the left channel
-may be provided as input to the detector, and the the detector's
-output pulses may be compared against the right channel.
+response to the recorded training songs against these ``canonical''
+detection events.  To this end, besides the trained network object,
+our learning code produces an audio file consisting of all of the
+training data on the left audio channel and a delta function at each
+aligned moment of target syllable presentation on the right
+channel. Thus, when played on any audio player, the left channel may
+be provided as input to the detector, and the the detector's output
+pulses may be compared against the ``ground truth'' given by the
+canonical detection event provided by the right channel.
 
@@
 \begin{table}
-  \begin{tabular}{c|rr}
-    Target time & \% True positives & \% False positives \\ \hline
-    $t^*_1$ & 99.34 &  0.0015 \\
-    $t^*_2$ & 99.65 &  0.00040 \\
-    $t^*_3$ & 98.98 &  0.0015 \\
-    $t^*_4$ & 99.30 &  0.00093 \\
-    $t^*_5$ & 99.73 &  0.000094 \\
-    $t^*_6$ & 98.31 &  0.0026
+  \begin{tabular}{rc|rr}
+    Bird & Target time & \% True positives & \% False positives \\ \hline
+    A&$ t^*_1$ & 99.72 &  0.0025 \\
+    &$ t^*_2$ & 99.81 &  0.00053 \\
+    &$ t^*_3$ & 99.78 &  0.00062 \\
+    &$ t^*_4$ & 99.69 &  0.00082 \\
+    &$ t^*_5$ & 99.75 &  0.00028 \\
+    &$ t^*_6$ & 99.54 &  0.0012 \\
+    B&$ t^*$ & 99.37 & 0.0039 \\
+    C&$ t^*$ & 99.13 & 0.0011 \\
+    D&$ t^*$ & 99.34 &0.0026 \\
   \end{tabular}
@@
 
-%\bibliography{birds}
-\begin{thebibliography}{10}
-
-\bibitem{Kao2005}
-Kao MH, Doupe AJ, Brainard MS.
-\newblock Contributions of an avian basal ganglia--forebrain circuit to
-  real-time modulation of song.
-\newblock Letters to Nature. 2005 February;433:638--643.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v433/n7026/abs/nature03127.html}.
-
-\bibitem{Wang2008interhemispheric}
-Wang CZH, Herbst JA, Keller GB, Hahnloser RHR.
-\newblock Rapid Interhemispheric Switching during Vocal Production in a
-  Songbird.
-\newblock PLOS Biology. 2008 October;6(10).
-\newblock Available from:
-  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.0060250}.
-
-\bibitem{Keller2009}
-Keller GB, Hahnloser RHR.
-\newblock Neural processing of auditory feedback during vocal practice in a
-  songbird.
-\newblock Nature. 2009 January;457:187--190.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v457/n7226/abs/nature07467.html}.
-
-\bibitem{Tumer2007pitchshift}
-Tumer EC, Brainard MS.
-\newblock Performance variability enables adaptive plasticity of 'crystallized'
-  adult birdsong.
-\newblock Nature. 2007 December;450:1240--1244.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v450/n7173/abs/nature06390.html}.
-
-\bibitem{Andalman2009}
-Andalman AS, Fee MS.
-\newblock A basal ganglia-forebrain circuit in the songbird biases motor output
-  to avoid vocal errors.
-\newblock Proceedings of the National Academy of Sciences of the United States
-  of America. 2009 July;106(30):12518--12523.
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2709669/}.
-
-\bibitem{Charlesworth2011learning}
-Charlesworth JD, Tumer EC, Warren TL, Brainard MS.
-\newblock Learning the microstructure of successful behavior.
-\newblock Nature Neuroscience. 2011;14(3):373--380.
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3045469/}.
-
-\bibitem{Warren2011}
-Warren TL, Tumer EC, Charlesworth JD, Brainard MS.
-\newblock Mechanisms and time course of vocal learning and consolidation in the
-  adult songbird.
-\newblock Journal of Neurophysiology. 2011 October;106(4):1806--1821.
-\newblock Available from:
-  \url{http://jn.physiology.org/content/106/4/1806.full}.
-
-\bibitem{Canopoli2014}
-Canopoli A, Herbst JA, Hahnloser RHR.
-\newblock A Higher Sensory Brain Region Is Involved in Reversing
-  Reinforcement-Induced Vocal Changes in a Songbird.
-\newblock The Journal of Neuroscience. 2014;34(20):7018--7026.
-\newblock Available from:
-  \url{http://www.jneurosci.org/content/34/20/7018.full}.
-
-\bibitem{Leonardo1999}
-Leonardo A, Konishi M.
-\newblock Decrystallization of adult birdsong by perturbation of auditory
-  feedback.
-\newblock Nature. 1999 June;399:466--470.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v399/n6735/abs/399466a0.html}.
-
-\bibitem{Skocik2013}
-Skocik M, Kozhevnikov A.
-\newblock Real-time system for studies of the effects of acoustic feedback on
-  animal vocalizations.
-\newblock Frontiers in Neural Circuits. 2013 January;6(111).
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3539774/}.
-
-\bibitem{Poole2012}
-Poole B, Markowitz JE, Gardner TJ.
-\newblock The Song Must Go On: Resilience of the Songbird Vocal Motor Pathway.
-\newblock PLOS One. 2012 June;7(6).
-\newblock Available from:
-  \url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0038173}.
-
-\bibitem{Tchernichovski2011SAP}
-Tchernichovski O. Sound Analysis Pro User Manual; 2011.
-\newblock Available from: \url{http://soundanalysispro.com/manual-1}.
-
-\bibitem{Jacobsen2003SlidingDFT}
-Jacobsen E, Lyons R.
-\newblock The sliding DFT.
-\newblock IEEE Signal Processing Magazine. 2003 Mar;20(2):74--80.
-\newblock Available from:
-  \url{http://www.cmlab.csie.ntu.edu.tw/DSPCourse/reference/Sliding\%20DFT.pdf}.
-
-\bibitem{hkp}
-Hertz JA, Krogh AS, Palmer RG.
-\newblock Introduction to the Theory of Neural Computation.
-\newblock Perseus Books; 1991.
-
-\bibitem{Noll1970pitchdetection}
-Noll AM.
-\newblock Pitch determination of human speech by the harmonic product spectrum,
-  the harmonic sum spectrum, and a maximum likelihood estimate.
-\newblock In: Proceedings of the symposium on computer processing in
-  communications. vol.~19; 1970. p. 779--797.
-
-\end{thebibliography}
+\bibliography{birds}
 

### 2017-03-13 main.tex
New data. Not yet formatted properly.

@@
 % D: LNY4RB 0.25s
-
-Data consisted of extracted, aligned songs, and additional song-length samples of non-song (cage noise, calls, other birds' song).
-These were divided into training and test sets.
-For Bird A, we used 1000 songs and 1000 non-song samples for training, and the remaining 1818 songs and 3636 nonsongs for testing.
-For birds B, C, and D, we used 1000 songs and 1000 non-songs for training, and another 1000 of each for testing.
-
-%\begin{table}
-% \begin{tabular}{c|cccc}
-%    Bird & Training songs & Test songs & Training non-song & Test non-song \\ %\hline
- %   A & 1000 & 1818 & 1000 & 3636 \\
-  %  B & 1000 & 1000 & 1000 & 1000 \\
- %   C & 1000 & 1000 & 1000 & 1000 \\
-%    D & 1000 & 1000 &  1000 & 1000
-%  \end{tabular}
-%  \caption{Sizes of our training and test sets.}
-%  \label{table:datasizes}
-%\end{table}
+% E: LR28 0.14s
+
+\begin{table}
+ \begin{tabular}{c|cccccc}
+    Bird & Training songs & Test songs & Training non-song & Test non-song & Target time  & Time window length\\ %\hline
+   A & 1000 & 1000 & 1000 & 3636 & 150:50:400 & 50 \\
+   B & 1000 & 1000 & 1000 & 1000 & 195 & 50 \\
+    C & 1000 & 1000 & 1000 & 1000 & 380 & 50 \\
+    D & 1000 & 1000 &  1000 & 1000 & 250 & 50 \\
+    E & 1000 & 1000 & 1000 & 1000 & 140 & 80
+  \end{tabular}
+  \caption{Sizes of our training and test sets, times of interest, and detection time window length (milliseconds).}
+  \label{table:datasizes}
+\end{table}
+
+Data consisted of extracted, aligned songs, and additional song-length
+samples of non-song (cage noise, calls, other birds' song). These were
+divided into training and test sets. For all birds, we used 1000 songs
+and 1000 non-song samples for training. For Bird A we used the
+remaining 1818 songs and 3636 nonsongs for testing. For each of the
+other birds, we tested on another 1000 songs and 1000
+nonsongs. Table~\ref{table:datasizes} lists the dataset sizes, along
+with the target detection points and the length of the time history
+window used for detection, which was the only parameter that we
+adjusted.
 
@@
 
-Accuracies for our six test syllables are shown in
+Accuracies for our six test syllables for bird A are shown in
 \fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
@@
     Bird & Target time & \% True positives & \% False positives \\ \hline
-    A&$ t^*_1$ & 99.72 &  0.0025 \\
+    A&$ t^*_1$ & 99.72 &  0.0025 \\ % lny64
     &$ t^*_2$ & 99.81 &  0.00053 \\
@@
     &$ t^*_6$ & 99.54 &  0.0012 \\
-    B&$ t^*$ & 99.37 & 0.0039 \\
-    C&$ t^*$ & 99.13 & 0.0011 \\
-    D&$ t^*$ & 99.34 &0.0026 \\
+    B&$ t^*$ & 99.370992640000068   0.003887539045000 \\ % lny46
+    C&$ t^*$ & 99.218862389999998   0.001205787775200 \\ % lny42
+    D&$ t^*$ & 99.342842499999975   0.00260723410200  \\ % lny4rb
+    E& $t^*$ & 99.726635259999981   0.003169430259900 \\ % lr28
+    E& $t^*$ & 99.464679732673261   0.004515502156436 \\ % lr12
+    E& $t^*$ & 99.862406879999966   0.001215710723800 \\ % lr13
+    E& $t^*$ & 99.646613329999980   0.003110002775000 % lr77
   \end{tabular}
@@
 \noindent Our implementation of the syllable
-detector is available under an open-source license {\em (Tentatively GPL, but we are still evaluating which open source license is the most suitable for this software):}
+detector is available under the GNU General Public License 3.
 \begin{description}

### 2017-03-14 main.tex

@@
 corrective signal to bias song away from those disruptions
-\cite{Andalman2009}. The
-song change is isolated to within roughly 10 milliseconds (ms) of the
-stimulus, and the shape of the learned response can be predicted by
-a simple mechanism \cite{Charlesworth2011learning}. AFP transfers the error signal to
-the robust nucleus of the arcopallium (RA) using
-NMDA-receptor--mediated glutamatergic transmission \cite{Warren2011}. By looking at song
-recovery after applying the same pitch-shift paradigm,
-the caudal medial nidopallium is
-implicated in memorising or recalling a recent song target, but for
-neither auditory processing nor directed motor learning \cite{Canopoli2014}.
+\cite{Andalman2009}. The song change is isolated to within roughly 10
+milliseconds (ms) of the stimulus, and the shape of the learned
+response can be predicted by a simple mechanism
+\cite{Charlesworth2011learning}. AFP transfers the error signal to the
+robust nucleus of the arcopallium (RA) using NMDA-receptor--mediated
+glutamatergic transmission \cite{Warren2011}. The course of song
+recovery after applying such a pitch-shift paradigm showed that the
+caudal medial nidopallium is implicated in memorising or recalling a
+recent song target, but in neither auditory processing nor directed
+motor learning \cite{Canopoli2014}.
 

### 2017-03-15 main.tex
Textual updates for the plethora of new birds.

--- 2017-03-13 main.tex
@@
 depend on the makeup of the non-song data, we have found that training
-with as low as a 1:1 ratio of non-song to song---or roughly 10-20
+with as low as a 1:1 ratio of non-song to song---or roughly 10
 minutes of non-song audio---yields excellent results on most birds.
@@
 44.1\rceil /44.1\approx 1.4966$ ms.  We set the network's input space
-to 30 ms long, and to span frequencies from 1--8 kHz, which contains
+to 50 ms long, and to span frequencies from 1--8 kHz, which contains
 the fundamentals and several overtones of most zebra finch
@@
 We measure accuracy and characterise timing for 6 randomly chosen
-points in one bird's song (Bird A), and present accuracy results for
-three more birds (B, C, D) (timing does not appreciably vary among
+points in one bird's song (lny64), and present accuracy results for
+seven more birds (the detector's timing does not appreciably vary among
 birds). The results presented here are representative of real-world
@@
 
-% A:
-% B: LNY46 0.28s @ 140ms
-% C: LNY42 0.38s
-% D: LNY4RB 0.25s
-% E: LR28 0.14s
-
 \begin{table}
- \begin{tabular}{c|cccccc}
-    Bird & Training songs & Test songs & Training non-song & Test non-song & Target time  & Time window length\\ %\hline
-   A & 1000 & 1000 & 1000 & 3636 & 150:50:400 & 50 \\
-   B & 1000 & 1000 & 1000 & 1000 & 195 & 50 \\
-    C & 1000 & 1000 & 1000 & 1000 & 380 & 50 \\
-    D & 1000 & 1000 &  1000 & 1000 & 250 & 50 \\
-    E & 1000 & 1000 & 1000 & 1000 & 140 & 80
+ \begin{tabular}{c|ccc}
+    Bird & Test songs & Test non-songs & Target time (ms) \\ \hline
+    lny64 & 1818 & 10545 & 150:50:400 \\
+    lny46 & 1000 & 1000 & 195 \\
+    lny42 & 1245 & 1223 & 380 \\
+    lny4rb & 1000 & 1000 & 250 \\
+    lr12 & 2000 & 2008 & 180 \\
+    lr13 & 2000 & 2001 & 300 \\
+    lr28 & 3000 & 3014 & 140 \\
+    lr77 & 2000 & 2008 & 320
   \end{tabular}
-  \caption{Sizes of our training and test sets, times of interest, and detection time window length (milliseconds).}
+  \caption{For each bird, detectors were trained for the specified
+    timepoints, using 1000 songs and 1000 non-song samples of the same
+    length. The rest of the audio samples---shown here---were used to
+    test the learned detectors.}
   \label{table:datasizes}
@@
 divided into training and test sets. For all birds, we used 1000 songs
-and 1000 non-song samples for training. For Bird A we used the
-remaining 1818 songs and 3636 nonsongs for testing. For each of the
-other birds, we tested on another 1000 songs and 1000
-nonsongs. Table~\ref{table:datasizes} lists the dataset sizes, along
-with the target detection points and the length of the time history
-window used for detection, which was the only parameter that we
-adjusted.
+and 1000 non-song samples for training. Table~\ref{table:datasizes}
+lists the test dataset sizes and target detection points. For lny64,
+we arbitrarily chose six linearly spaced target times. For each of the
+other birds, we eyeballed one promising-looking detection point.
 
@@
 
-\fig{fig:song} shows the song for which we present accuracy and timing
-results.  Six example target trigger points are shown, at 150, 200,
-250, 300, 350, and 400 milliseconds after the beginning of the aligned
+\fig{fig:song} shows a typical song. The image is produced by
+averaging the aligned spectrograms, in this case for the bird lny64.
+Six target trigger points are shown for this song, at 150, 200, 250,
+300, 350, and 400 milliseconds after the beginning of the aligned
 samples, as indicated by the red lines, and referred to henceforth as
-$t^*_1\ldots t^*_6$ respectively.  The rectangles show the recognition
-window (30 ms, 1--8 kHz) for each instant in the song.
+$t^*_1\ldots t^*_6$ respectively.  For clarity, the rectangles show a
+shorter recognition window than our standard: here 30 ms, but still
+spanning 1--8 kHz, for each target time in the song.\marginpar{Is this
+  OK?  Should I update to a 50-ms window? They'll overlap!}
 
@@
   \caption{Each plot shows one network output unit's responses to all
-    2818 presentations of the song shown in \fig{fig:song}.  For
+    2818 presentations of lny64's song shown in \fig{fig:song}.  For
     simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
@@
 
-Accuracies for our six test syllables for bird A are shown in
+Accuracies for our six test syllables for bird lny64 are shown
+in\marginpar{Update if I include all birds in this figure!}
 \fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
 songs were always 100\%).  For each test syllable we trained 100
-different networks, which differ from each other in the random
-initialisation of network weights and on which random subset of our
-songs was used for training.  Each network's accuracy is shown as a
-single point in each chart in \fig{fig:accuracies}, and the means are
-shown in Table~\ref{table:accuracies}.  Evidently some syllables are
-easier to identify reliably than others (for example, $t^*_3$ and
-$t^*_6$ are occasionally confused due to their similar appearance, as
-could be guessed from \fig{fig:song}); thus, if the experimental
-design permits, it can be worthwhile to choose an easily identifiable
+different networks, which differ from each other both in the random
+initialisation of network weights and in which random subset of the
+song-and-nonsong corpus was used for training.  Each network's
+accuracy is shown as a single point in each chart in
+\fig{fig:accuracies}, and the means are shown in
+Table~\ref{table:accuracies}.  Evidently some syllables are easier to
+identify reliably than others (for example, $t^*_3$ and $t^*_6$ are
+occasionally confused due to their similar appearance, as could be
+guessed from \fig{fig:song}); thus, if the experimental design
+permits, it can be worthwhile to choose an easily identifiable
 syllable.  Furthermore, as can be seen from \fig{fig:accuracies}, for
@@
   \caption{Accuracy variability over 100 different training runs for
-    each of the six test syllables.  Each dot shows the test-set
+    each of the six test syllables for lny64.  Each dot shows the test-set
     accuracy for an independently trained detector.  Because the
@@
     occlude same-valued measurements, test syllable is also indicated
-    by colour.  The means are given in Table~\ref{table:accuracies}.}
+    by colour.  The means for all birds are given in Table~\ref{table:accuracies}.}
   \label{fig:accuracies}
@@
 detector, the Swift detector with serial output, and for the LabVIEW
-detector, for both the $\delta$-syllable and for the bird song at
+detector, for both the $\delta$-syllable and for lny64's song at
 $t^*_4$.
@@
     Swift+serial implementations, for the constructed
-    $\delta$-syllable and for trigger $t^*_4$ of the bird song.  The
+    $\delta$-syllable and for trigger $t^*_4$ of lny64's song.  The
     lines show latency; error bars are standard deviation (jitter).  Points have
@@
 Syllable choice impacts detector performance, but despite the variety
-of syllables chosen here, performance was fairly stable across
-syllables.  \fig{fig:TimingVsSyllable} shows measured timing data for
-the Swift+serial detector compared to the ideal, with
+of syllables shown here, performance was stable across syllables.
+\fig{fig:TimingVsSyllable} shows measured timing data for lny64's test
+syllables for the Swift+serial detector compared to the ideal, with
 $t_{\textrm{fft}}=1.5$ ms.  Given the variety of the test syllables,
@@
   \end{center}
-  \caption{Timing data for our 6 test syllables, for the ideal and the
+  \caption{Timing data for lny64's 6 test syllables, for the ideal and the
     Swift+serial detectors, with an FFT frame rate of 1.5 ms.  Point
@@
 We compared latency and jitter across our different detector
-implementations for the $\delta$-syllable and $t^*_4$, again with
+implementations for the $\delta$-syllable and lny64's $t^*_4$, again with
 $t_\mathrm{fft}=1.5$ ms.  Results are as follows:
@@
 \begin{tabular}{l|c|c|c| c}
-  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird: $t^*_4$} \\
+  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird lny64: $t^*_4$} \\
   Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
@@
   \caption{The different detectors for the constructed
-    $\delta$-syllable and for the bird song at $t^*_4$.  Point centres
+    $\delta$-syllable and for lny64's song at $t^*_4$.  Point centres
     show latency; error bars show jitter.}
@@
   \caption{Raw timing curves for all detectors measured during
-    detection of $t^*_4$ using 1.5-ms frames.  We extract the trigger events from each curve, from which we obtain the mean---latency---and
+    detection of lny64's $t^*_4$ using 1.5-ms frames.  We extract the trigger events from each curve, from which we obtain the mean---latency---and
     standard deviation---jitter.}
@@
 
-The song presented here was recorded with a fixed microphone mounted
-inside the cage.  We found that higher accuracy is achieved when a
+The songs presented here were recorded with fixed microphones mounted
+inside the cages.  We found that higher accuracy is achieved when a
 microphone is instead mounted on the bird's head, which maintains

### 2017-03-16 main.tex
Table 2 is still being generated. This is just an interim so I can work from home...

@@
 \fig{fig:song} shows a typical song. The image is produced by
-averaging the aligned spectrograms, in this case for the bird lny64.
-Six target trigger points are shown for this song, at 150, 200, 250,
-300, 350, and 400 milliseconds after the beginning of the aligned
-samples, as indicated by the red lines, and referred to henceforth as
-$t^*_1\ldots t^*_6$ respectively.  For clarity, the rectangles show a
-shorter recognition window than our standard: here 30 ms, but still
-spanning 1--8 kHz, for each target time in the song.\marginpar{Is this
-  OK?  Should I update to a 50-ms window? They'll overlap!}
+averaging the aligned spectrograms for the bird lny64.  Six target
+trigger points are shown for this song, at 150, 200, 250, 300, 350,
+and 400 milliseconds after the beginning of the aligned samples, as
+indicated by the red lines, and for convenience will be referred to
+henceforth as $t^*_1\ldots t^*_6$ respectively.  For clarity, the
+rectangles show a shorter recognition window than our standard: here
+30 ms, but still spanning 1--8 kHz, for each target time in the
+song.\marginpar{Is this OK?  Should I update to a 50-ms window?
+  They'll overlap.}
 
@@
   \caption{Each plot shows one network output unit's responses to all
-    2818 presentations of lny64's song shown in \fig{fig:song}.  For
-    simplicity, we show only the syllables $t^*_1$, $t^*_4$, and
-    $t^*_6$, and we do not show the non-response to presentation of non-song. The horizontal axis is time relative to the beginning of
-    the aligned song, and the vertical axis is an index for the 2818
-    individual song presentations. The grey shading shows the audio
-    amplitude of song Y at time T. Detection events on training songs
-    are shown in cyan, with detections of unseen test songs in red.  To provide an intuition of intra-song variability, songs have been
-    stably sorted by the time of detection events; thus, each of the
-    three detection graphs shows the songs in a different order.}
+    2818 presentations of lny64's song shown in \fig{fig:song}.  We
+    show only the syllables $t^*_1$, $t^*_4$, and $t^*_6$, and we do
+    not show the non-response to presentation of non-song. The
+    horizontal axis is time relative to the beginning of the aligned
+    song, and the vertical axis is an index for the 2818 individual
+    song presentations. The grey shading shows the audio amplitude of
+    song Y at time T. Detection events on training songs are shown in
+    cyan, with detections of unseen test songs in red.  To provide an
+    intuition of intra-song variability, songs have been stably sorted
+    by the time of detection events; thus, each of the three detection
+    graphs shows the songs in a different order.}
   \label{fig:detection_raster}
@@
 are gathered on the same hardware setup, this is the digital signal
-that the detector will see---only the non-song data may be different.
+that the detector will see.
 %Accuracy \marginpar{Should be!
@@
     &$ t^*_6$ & 99.54 &  0.0012 \\
-    B&$ t^*$ & 99.370992640000068   0.003887539045000 \\ % lny46
-    C&$ t^*$ & 99.218862389999998   0.001205787775200 \\ % lny42
-    D&$ t^*$ & 99.342842499999975   0.00260723410200  \\ % lny4rb
-    E& $t^*$ & 99.726635259999981   0.003169430259900 \\ % lr28
-    E& $t^*$ & 99.464679732673261   0.004515502156436 \\ % lr12
-    E& $t^*$ & 99.862406879999966   0.001215710723800 \\ % lr13
-    E& $t^*$ & 99.646613329999980   0.003110002775000 % lr77
+    B&$ t^*$ & 99.370992640000068   &0.003887539045000 \\ % lny46
+    C&$ t^*$ & 99.218862389999998   &0.001205787775200 \\ % lny42
+    D&$ t^*$ & 99.342842499999975   &0.00260723410200  \\ % lny4rb
+    E& $t^*$ & 99.726635259999981   &0.003169430259900 \\ % lr28
+    E& $t^*$ & 99.464679732673261   &0.004515502156436 \\ % lr12
+    E& $t^*$ & 99.862406879999966   &0.001215710723800 \\ % lr13
+    E& $t^*$ & 99.646613329999980   &0.003110002775000 % lr77
   \end{tabular}

### 2017-03-20 main.tex
Add data and graph for new accuracy measurements.
Makefile: new target to build the diff (makes some unwarranted assumptions!)

@@
 % Use the asterisk to denote corresponding authorship and provide email address in note below.
-\textsuperscript{\textcurrency} Corresponding author: bwpearre@gmail.com (BP)
+\textsuperscript{\textcurrency} Corresponding author: bwpearre@bu.edu (BP)
 
@@
   isolated and detected with false negative and false positive rates
-  well under 1\% and 0.001\% respectively. The detector can run on a
+  well under 1\% and 0.005\% respectively. The detector can run on a
   stock Mac Mini with triggering delay of less than a millisecond and
@@
 \item Fast: sub-millisecond latencies, with jitter around 2 ms.
-\item Accurate: false negative rates under 1\% and false positive rates under 0.001\% for a variety of syllables.
+\item Accurate: false negative rates under 1\% and false positive rates under 0.005\% for a variety of syllables.
 \item State-of-the-art performance with default parameters.
@@
   \includegraphics[width=\textwidth]{Fig1}
-  \caption{The spectrogram of a typical song, used as the example
+  \caption{The spectrogram of the song of lny64, used as an example
     throughout this paper.  This image was made by superposing the
     spectra of our 2818 aligned songs.  Our example detection points,
-    $t^*_1\ldots t^*_6$, are shown as red lines, with the recognition
-    regions (30 ms $\times$ 1--8 kHz) marked as rectangles.}
+    $t^*_1\ldots t^*_6$, are shown as red lines, with example recognition
+    regions of 30 ms $\times$ 1--8 kHz marked as rectangles.}
   \label{fig:song}
@@
 
-Note that this results in time being discretised into chunks of length
+This results in time being discretised into chunks of length
 $t_\mathrm{fft}$.  Because each Fourier transform computation contains
@@
 values $F$ (for example, 1--8 kHz) at some number of the most recent
-frames $n_\textrm{fft}$.  Any time $t$ falls within frame
-$\tau(t)$, and $t-t_\textrm{fft}$ falls within the previous frame, so the recognition region that the neural network receives as input
-consists of the spectrogram values over $F$ at $\tau(t)$ and those
-from the contiguous set of recent frames: $T = \{$ $\tau(t)$,
+frames $n_\textrm{fft}$.  Any time $t$ falls within frame $\tau(t)$,
+and $t-t_\textrm{fft}$ falls within the previous frame, so the
+recognition region that the neural network receives as input consists
+of the spectrogram values over $F$ at $\tau(t)$ and those from the
+contiguous set of recent frames: $T = \{$ $\tau(t)$,
 $\tau(t-t_\textrm{fft})$, $\tau(t-{2t_\textrm{fft}})\ldots$
-$\tau(t-n_\textrm{fft}t_\textrm{fft})\}$.  We have found that a 30-ms
-time window---which will yield $n_\textrm{fft} = |T|=\lfloor 30\textrm{
+$\tau(t-n_\textrm{fft}t_\textrm{fft})\}$.  Time windows of 30--50
+ms---the latter will yield $n_\textrm{fft} = |T|=\lfloor 50\textrm{
   ms}/t_\textrm{fft}\rceil$ frames---of frequencies spanning 1--8 kHz
-works well for most syllables.
-
-Six examples of chosen target moments in the song, and their
-corresponding recognition regions $F$ and $T$, are shown in
-\fig{fig:song}.
+generally work well.
+
+Six examples of chosen target moments in the song, with recognition
+regions $F = 1--8$ kHz and $T=30$ ms, are shown in \fig{fig:song}.
 
@@
 \end{equation*}
-During the network's training phase, the elements of the matrices and bias vectors are learned by back-propagation of errors
-over a training set.  A more detailed explanation of neural networks
-may be found in \cite{hkp}.
+During the network's training phase, the elements of the matrices and
+bias vectors are learned by back-propagation of errors over a training
+set.  A more detailed explanation of neural networks may be found in
+\cite{hkp}.
 
@@
 
-The network is trained using MATLAB's neural network toolbox, with Scaled Conjugate Gradient ({\tt trainscg}). We tried
-a variety of feedforward neural network geometries, from simple
-1-layer perceptrons to geometries with many hidden nodes, as well as
-autoencoders. Perhaps surprisingly, even the former yields excellent
-results on many syllables, but a 2-layer perceptron with a very small
-hidden layer---with a unit count 2-4 times the number of target
+The network is trained using MATLAB's neural network toolbox, with
+Scaled Conjugate Gradient ({\tt trainscg}). We tried a variety of
+feedforward neural network geometries, from simple 1-layer perceptrons
+to geometries with many hidden nodes, as well as autoencoders. Perhaps
+surprisingly, even the former yields excellent results on many
+syllables, but a 2-layer perceptron with a very small hidden
+layer---with a unit count 2-4 times the number of target
 syllables---was a good compromise between accuracy and training
@@
 decreasing the frame interval generally decreases both latency and
-jitter, but also increases training time.
+jitter, but also increases training time. Sometimes the effects are
+syllable-specific: for example, using a 30-ms time window rather than
+50 ms speeds training while usually having a minimal effect on
+detector performance, but occasionally a syllable will be seen for
+which extending the window to 80 ms is very helpful.
 
@@
 and 400 milliseconds after the beginning of the aligned samples, as
-indicated by the red lines, and for convenience will be referred to
-henceforth as $t^*_1\ldots t^*_6$ respectively.  For clarity, the
-rectangles show a shorter recognition window than our standard: here
-30 ms, but still spanning 1--8 kHz, for each target time in the
-song.\marginpar{Is this OK?  Should I update to a 50-ms window?
-  They'll overlap.}
+indicated by the red lines. In discussion, we will refer to these six
+as $t^*_1\ldots t^*_6$ respectively.  For clarity, the rectangles show
+a shorter recognition window than our standard: here 30 ms, but still
+spanning 1--8 kHz, for each target time in the song.\marginpar{Is this
+  OK?  Should I update to a 50-ms window?  They'll overlap.}
 
@@
 
-Accuracies for our six test syllables for bird lny64 are shown
-in\marginpar{Update if I include all birds in this figure!}
-\fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
+Accuracies are shown in \fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
 songs were always 100\%).  For each test syllable we trained 100
 different networks, which differ from each other both in the random
-initialisation of network weights and in which random subset of the
-song-and-nonsong corpus was used for training.  Each network's
+initialisation of network weights and in the random subset of the
+song-and-nonsong corpus that was used for training.  Each network's
 accuracy is shown as a single point in each chart in
@@
 Table~\ref{table:accuracies}.  Evidently some syllables are easier to
-identify reliably than others (for example, $t^*_3$ and $t^*_6$ are
+identify reliably than others (for example, for lny64, $t^*_3$ (250 ms) and $t^*_6$ (400 ms) are
 occasionally confused due to their similar appearance, as could be
@@
   \caption{Accuracy variability over 100 different training runs for
-    each of the six test syllables for lny64.  Each dot shows the test-set
+    each of the test detection points.  Each dot shows the test-set
     accuracy for an independently trained detector.  Because the
@@
     occlude same-valued measurements, test syllable is also indicated
-    by colour.  The means for all birds are given in Table~\ref{table:accuracies}.}
+    by colour.  The means are given in Table~\ref{table:accuracies}.}
   \label{fig:accuracies}
@@
 \begin{table}
-  \begin{tabular}{rc|rr}
-    Bird & Target time & \% True positives & \% False positives \\ \hline
-    A&$ t^*_1$ & 99.72 &  0.0025 \\ % lny64
-    &$ t^*_2$ & 99.81 &  0.00053 \\
-    &$ t^*_3$ & 99.78 &  0.00062 \\
-    &$ t^*_4$ & 99.69 &  0.00082 \\
-    &$ t^*_5$ & 99.75 &  0.00028 \\
-    &$ t^*_6$ & 99.54 &  0.0012 \\
-    B&$ t^*$ & 99.370992640000068   &0.003887539045000 \\ % lny46
-    C&$ t^*$ & 99.218862389999998   &0.001205787775200 \\ % lny42
-    D&$ t^*$ & 99.342842499999975   &0.00260723410200  \\ % lny4rb
-    E& $t^*$ & 99.726635259999981   &0.003169430259900 \\ % lr28
-    E& $t^*$ & 99.464679732673261   &0.004515502156436 \\ % lr12
-    E& $t^*$ & 99.862406879999966   &0.001215710723800 \\ % lr13
-    E& $t^*$ & 99.646613329999980   &0.003110002775000 % lr77
+  % replot_accuracies_concatanated('tex', true) for confusion_log_perf.txt (1.5 ms).
+  \begin{tabular}{lc|cc}
+    Bird & Target time (ms) & \% True positives & \% False positives \\ \hline
+lny64 & 	150 $[t^*_1]$ & 	99.66 & 	0.0026 \\ % /Volumes/Data/song/lny64/confusion_log_perf.txt
+      & 	200 $[t^*_2]$ & 	99.75 & 	0.00052 \\ %                                               
+      & 	250 $[t^*_3]$ & 	99.79 & 	0.00051 \\ %                                               
+      & 	300 $[t^*_4]$ & 	99.70 & 	0.00077 \\ %                                               
+      & 	350 $[t^*_5]$ & 	99.73 & 	0.00022 \\ %                                               
+      & 	400 $[t^*_6]$ & 	99.52 & 	0.0010 \\ %                                                
+lny42 & 	380 & 	99.09 & 	0.0011 \\ % /Volumes/Data/song/lny42/confusion_log_perf.txt          
+lny46 & 	140 & 	99.53 & 	0.0041 \\ % /Volumes/Data/song/lny46/confusion_log_perf.txt          
+lny4rb & 	250 & 	99.36 & 	0.0025 \\ % /Volumes/Data/song/lny4rb/confusion_log_perf.txt        
+lr12 & 	180 & 	99.52 & 	0.0033 \\ % /Volumes/Data/song/lr12/all/confusion_log_perf.txt        
+lr13 & 	300 & 	99.83 & 	0.0013 \\ % /Volumes/Data/song/lr13/all/confusion_log_perf.txt        
+lr28 & 	140 & 	99.73 & 	0.0028 \\ % /Volumes/Data/song/lr28/all/confusion_log_perf.txt        
+lr77 & 	320 & 	99.58 & 	0.0028  % /Volumes/Data/song/lr77/all/confusion_log_perf.txt          
   \end{tabular}
@@
 
-We repeated the above experiment using only 2 hidden units per syllable.
-Training under this condition is much faster, but the mean true positive rate per syllable decreases by an average of 18\% across our test syllables, and the mean false positive rate increases by roughly 23\%.
+We repeated the above experiment using only 2 hidden units per
+syllable.  Training under this condition is much faster, but the mean
+true positive rate per syllable decreases by an average of 18\% across
+our test syllables, and the mean false positive rate increases by
+roughly 23\%.
+
+Changing the FFT interval to 2 ms significantly reduces the
+computational requirement. We found that some syllables were slightly
+easier to detect and some more difficult, with no significant
+difference on average.
 
@@
 \vspace{8pt}
-\begin{tabular}{l|c|c|c| c}
-  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird lny64: $t^*_4$} \\
+\begin{tabular}{l|c|c|c|c}
+  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird: $t^*_4$} \\
   Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
@@
 \item The detector is accurate. False negative and false positive
-  rates can be well under 1\% and 0.001\% respectively, and trading
+  rates can be well under 1\% and 0.005\% respectively, and trading
   these two numbers off against each other is through a single

### 2017-03-21 main.tex
Fixed training/test set sizes.

--- 2017-03-16 main.tex
@@
  \begin{tabular}{c|ccc}
-    Bird & Test songs & Test non-songs & Target time (ms) \\ \hline
-    lny64 & 1818 & 10545 & 150:50:400 \\
-    lny46 & 1000 & 1000 & 195 \\
-    lny42 & 1245 & 1223 & 380 \\
-    lny4rb & 1000 & 1000 & 250 \\
-    lr12 & 2000 & 2008 & 180 \\
-    lr13 & 2000 & 2001 & 300 \\
-    lr28 & 3000 & 3014 & 140 \\
-    lr77 & 2000 & 2008 & 320
+    Bird & Test songs = Test non-songs & Target time (ms) \\ \hline
+    lny64 & 1818 & 150:50:400 \\
+    lny46 & 1000 & 195 \\
+    lny42 & 1245 & 380 \\
+    lny4rb & 1000 & 250 \\
+    lr12 & 2000 & 180 \\
+    lr13 & 2000 & 300 \\
+    lr28 & 3000 & 140 \\
+    lr77 & 2000 & 320
   \end{tabular}
@@
     timepoints, using 1000 songs and 1000 non-song samples of the same
-    length. The rest of the audio samples---shown here---were used to
-    test the learned detectors.}
+    length. The test set consisted of the remaining songs and an
+    equal number of non-songs.}
   \label{table:datasizes}
@@
 divided into training and test sets. For all birds, we used 1000 songs
-and 1000 non-song samples for training. Table~\ref{table:datasizes}
+and 1000 non-song samples for training. The test set consisted of the remaining songs and an equal number of non-songs. Table~\ref{table:datasizes}
 lists the test dataset sizes and target detection points. For lny64,

### 2017-03-21 main.tex
Ben's final(ish) round of edits. Still need to provide DOIs for code and data repositories, and update diffs.

@@
 mesopallium may hold a representation of song against which auditory
-signals are compared \cite{Keller2009}.  Renditions of a syllable that
-were disrupted by white noise when they were slightly above (or below)
-the syllable's average pitch showed that the apparently random natural
+signals are compared \cite{Keller2009}.  The disruption by white noise
+of renditions of a syllable that were slightly above (or below) the
+syllable's average pitch showed that the apparently random natural
 variability in songbird motor output is used to drive change in the
-song \cite{Tumer2007pitchshift}. Furthermore, AFP produces a
+song \cite{Tumer2007pitchshift}, and AFP produces a
 corrective signal to bias song away from those disruptions
@@
 \item[Latency:] The average delay between the target syllable being sung and the detection.
-\item[Jitter:] The amount that latency changes from instance to instance of song.  Our measure of jitter is the standard deviation $\sigma$ of latency.
+\item[Jitter:] The amount that latency changes from instance to instance of song.  Our measure of jitter is the standard deviation of latency.
 \item[Versatility:] Is detection possible at ``difficult'' syllables?
@@
 operators \cite{Leonardo1999}.  The system had a latency of 50 or 100 ms,
-and jitter or accuracy were not reported.  As access to
+and accuracy and jitter were not reported.  As access to
 computational resources has improved, approaches have changed: in
@@
 based on our experience with a similar technique, accurate.
-2009 also saw the application of a neural network to a spectral image of
+2009 saw the application of a neural network to a spectral image of
 song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
@@
 \begin{enumerate}
-\item Record and align a corpus of training songs.
-The technique has been published in \cite{Poole2012}.
-200 songs generally yield acceptable results, but here we standardise on 1000-song training sets.
+\item Record and align a corpus of training songs.  The technique has
+  been published in \cite{Poole2012}. As few as 200 songs can yield
+  acceptable results, but here we standardise on 1000-song training
+  sets.
 \item Choose one or more instants in the song that should create a trigger event, and train a
@@
 song'') and slightly more variable otherwise (``undirected''); we
-train and test on undirected song since this is more commonly studied,
-as well as more challenging.  Recordings were made inside a
+train and test on undirected song since this is both more commonly
+studied and more challenging.  Recordings were made inside a
 sound-isolating chamber in which was mounted a recording microphone
-(Audio-Technica AT803), using methods similar to those described in \cite{Tchernichovski2011SAP}, Chapter 2.  The songs were
-time-aligned as described in \cite{Poole2012}.
+(Audio-Technica AT803), using methods similar to those described in
+\cite{Tchernichovski2011SAP}, Chapter 2.  The songs were time-aligned
+as described in \cite{Poole2012}.
 
@@
 Six examples of chosen target moments in the song, with recognition
-regions $F = 1--8$ kHz and $T=30$ ms, are shown in \fig{fig:song}.
+regions $F =$ 1--8 kHz and $T=30$ ms, are shown in \fig{fig:song}.
 
@@
 minutes of non-song audio---yields excellent results on most birds.
-These data are also included in our accuracy tests.
 
@@
 
-During runtime, the network may produce above-threshold responses to nearby
-frames.  Thus, after the first response, subsequent responses are suppressed
-for 100 ms.  However, for the accuracy figures presented here, we used
-the un-de-bounced network response.
+During runtime, the network may produce above-threshold responses to
+nearby frames.  Thus, after the first response, subsequent responses
+are suppressed for 100 ms.  However, for the accuracy measurements
+presented here, we used the un-de-bounced network response.
 
@@
 National Instruments PCI-6251 card on a PC with an Intel Core i5-4590
-processor at 3.7GHz (a relatively low-end machine) with 32 gigabytes
+processor at 3.7GHz (a relatively low-end machine) with 24 gigabytes
 of RAM, running Microsoft Windows 8.1 Pro and Windows 10.
@@
 performance per frame. In order to avoid the apparent problem of every
-non-detected non-syllable counting as a true negative, we also tried
-defining accuracy on a per-song basis, such that a song-length sample without the
-target syllable counted as a single true negative.  Computing the
-optimal output thresholds on a per-frame rather than a per-song basis resulted in higher
-thresholds and thus a slight reduction in both the false-positive and true-positive rates (reversible by increasing $C_n$), while also providing a valid
+non-detected non-matching frame counting as a true negative, we also
+tried defining accuracy on a per-song basis, such that a song-length
+sample without the target syllable counted as a single true negative.
+Computing the optimal output thresholds on a per-frame rather than a
+per-song basis resulted in higher thresholds and thus a slight
+reduction in both the false-positive and true-positive rates
+(reversible by increasing $C_n$), while also providing a valid
 definition of false-positive rate for data streams that had not been
@@
 indicated by the red lines. In discussion, we will refer to these six
-as $t^*_1\ldots t^*_6$ respectively.  For clarity, the rectangles show
+as $t^*_1\ldots t^*_6$ respectively.  To avoid overlap in the graphic, the rectangles show
 a shorter recognition window than our standard: here 30 ms, but still
-spanning 1--8 kHz, for each target time in the song.\marginpar{Is this
-  OK?  Should I update to a 50-ms window?  They'll overlap.}
+spanning 1--8 kHz, for each target time in the song.
 
@@
 lny42 &         380 &   99.09 &         0.0011 \\ % /Volumes/Data/song/lny42/confusion_log_perf.txt          
-lny46 & 	140 & 	99.53 & 	0.0041 \\ % /Volumes/Data/song/lny46/confusion_log_perf.txt          
-lny4rb & 	250 & 	99.36 & 	0.0025 \\ % /Volumes/Data/song/lny4rb/confusion_log_perf.txt        
+lny46 &         140 &   99.17 &         0.0037 \\ % /Volumes/Data/song/lny46/confusion_log_perf.txt          
+lny4rb &        250 &   99.37 &         0.0026 \\ % /Volumes/Data/song/lny4rb/confusion_log_perf.txt        
 lr12 &  180 &   99.52 &         0.0033 \\ % /Volumes/Data/song/lr12/all/confusion_log_perf.txt        
@@
 Changing the FFT interval to 2 ms significantly reduces the
-computational requirement. We found that some syllables were slightly
-easier to detect and some more difficult, with no significant
-difference on average.
+computational requirement during learning. For most syllables, results
+were identical, but some syllables are slightly easier to detect and
+some slightly more difficult, with no significant difference on
+average.
 
@@
     \item The main programme is {\tt learn\_detector.m}.
-    \item The top section of {\tt learn\_detector.m} contains global
-      configuration variables.  The important ones are the data
-      directory and the bird name: as supplied, the software looks for
-      {\tt \textless data directory\textgreater/\textless bird
-        name\textgreater/song.mat}.  Modify as necessary.
-    \item The format for the data file {\tt \textless data
-      directory\textgreater/\textless bird name\textgreater/song.mat}
-      is specified in {\tt syllable-detector-learn/README.md}: a
-      MATLAB file containing aligned song, nonsong (calls, cage noise,
-      etc), and the audio sampling frequency.
-    \item The configuration file {\tt \textless data
-      directory\textgreater/\textless bird name\textgreater/params.m}
-      is also described in {\tt README.md}.  If the file is empty, the
-      programme will display an image showing the average of the song
-      spectrograms (frequency vs.~time), tell the user how to define
-      syllables of interest, and exit.  If {\tt times\_of\_interest}
-      is defined, the programme will train the neural network.
-      Optionally, {\tt params.m} can also override any of the
-      configuration parameters at the top of {\tt learn\_detector.m}.
+    \item The top section of {\tt learn\_detector.m}
+      describes the configuration variables.
+    \item The current directory should contain the data in {\tt
+      song.mat}. The format for this file is specified in {\tt
+      README.md}: a MATLAB file containing aligned song, nonsong
+      (calls, cage noise, etc), and the audio sampling frequency.
+    \item {\tt learn\_detector} also looks in the current directory for
+      a configuration file {\tt params.m}, described in {\tt
+        README.md}.  If the file is empty, the programme will display
+      an image showing the average of the song spectrograms (frequency
+      vs.~time), tell the user how to define syllables of interest,
+      and exit.  If {\tt times\_of\_interest\_ms} is defined, the
+      programme will train the neural network.  Optionally, {\tt
+        params.m} can also override any of the configuration
+      parameters at the top of {\tt learn\_detector.m}.
+    \item Configuration may also be controlled by specifying
+      parameter-value pairs to the function call in the common MATLAB
+      fashion; for example, {\tt
+        learn\_detector('times\_of\_interest\_ms', [150 200])}.
     \item The software will produce a neural network file, {\tt

### 2017-03-21 main.tex
Wording.

--- 2017-03-21 main.tex
@@
 50 ms speeds training while usually having a minimal effect on
-detector performance, but occasionally a syllable will be seen for
-which extending the window to 80 ms is very helpful.
+detector performance (as is the case, for example, for all of the
+detection points for lny64), but occasionally a syllable will be seen
+for which extending the window to 80 ms is very helpful.
 
@@
  \begin{tabular}{c|ccc}
-    Bird & Test songs = Test non-songs & Target time (ms) \\ \hline
-    lny64 & 1818 & 150:50:400 \\
-    lny46 & 1000 & 195 \\
-    lny42 & 1245 & 380 \\
-    lny4rb & 1000 & 250 \\
-    lr12 & 2000 & 180 \\
-    lr13 & 2000 & 300 \\
-    lr28 & 3000 & 140 \\
-    lr77 & 2000 & 320
+    Bird & Test songs & Test non-songs & Target time (ms) \\ \hline
+    lny64 & 1818 & 10545 & 150:50:400 \\
+    lny46 & 1000 & 1000 & 195 \\
+    lny42 & 1245 & 1223 & 380 \\
+    lny4rb & 1000 & 1000 & 250 \\
+    lr12 & 2000 & 2008 & 180 \\
+    lr13 & 2000 & 2001 & 300 \\
+    lr28 & 3000 & 3014 & 140 \\
+    lr77 & 2000 & 2008 & 320
   \end{tabular}
@@
     timepoints, using 1000 songs and 1000 non-song samples of the same
-    length. The test set consisted of the remaining songs and an
-    equal number of non-songs.}
+    length. The rest of the audio samples---shown here---were used to
+    test the learned detectors.}
   \label{table:datasizes}
@@
 divided into training and test sets. For all birds, we used 1000 songs
-and 1000 non-song samples for training. The test set consisted of the remaining songs and an equal number of non-songs. Table~\ref{table:datasizes}
+and 1000 non-song samples for training. Table~\ref{table:datasizes}
 lists the test dataset sizes and target detection points. For lny64,

### 2017-03-21 main.tex
Final edits, merges, PDFs (main and diff).

--- 2017-03-21 main.tex
@@
 50 ms speeds training while usually having a minimal effect on
-detector performance, but occasionally a syllable will be seen for
-which extending the window to 80 ms is very helpful.
+detector performance (as is the case, for example, for all of the
+detection points for lny64), but occasionally a syllable will be seen
+for which extending the window to 80 ms is helpful.
 

### 2017-03-22 main.tex
Really final data for Table~\ref{table:accuracies}! Still need to update the figure.
Update version link to Nathan's Swift code.

@@
 lny42 &         380 &   99.09 &         0.0011 \\ % /Volumes/Data/song/lny42/confusion_log_perf.txt          
-lny46 &         140 &   99.17 &         0.0037 \\ % /Volumes/Data/song/lny46/confusion_log_perf.txt          
+lny46 &         140 &   99.11 &         0.0036 \\ % /Volumes/Data/song/lny46/confusion_log_perf.txt          
 lny4rb &        250 &   99.37 &         0.0026 \\ % /Volumes/Data/song/lny4rb/confusion_log_perf.txt        
@@
     \end{description}
-  \item[Swift:] {\tt https://github.com/gardner-lab/\dots\ syllable-detector-swift/releases/tag/0.3.1}
+  \item[Swift:] {\tt https://github.com/gardner-lab/\dots\ syllable-detector-swift/releases/tag/0.3.2}
     \begin{description}

### 2017-03-22 main.tex
Update Fig3. Add DOIs for data and software.

@@
 
-{\em If accepted, before final submission, DOIs for the latest code will be created as described in https://guides.github.com/activities/citable-code/ so that our results may be reproduced and the software deployed.
-For now, in lieu of final release DOIs, we provide these links to our development repositories.
-If you would like final versions and test datasets for review, please contact the corresponding author (bwpearre@bu.edu).}
-%\marginpar{PLOS doesn't really say where such things should go.}
-%\marginpar{Cleanup: mostly done.  But it would be nice to move all this to the lab's GitHub group.}
 \begin{description}
+\item[Data:]
+  Data and training configuration files for the 8 birds discussed here are available from the Open Science Framework at DOI: 10.17605/OSF.IO/BX76R .
+DOIs for the versions of the programmes used here are provided for reproducibility; however, we generally recommend downloading the latest release.
+%\item[Licenses:]
+%  All software is licensed under Open-Source-Initiative--approved licenses. Please see the software packages for individual choices.
 \item[Song alignment:] Jeff Markowitz's song alignment software \cite{Poole2012} can be found at
     {\tt https://github.com/jmarkow/zftftb}
-\end{description}
-
-\noindent Our implementation of the syllable
-detector is available under the GNU General Public License 3.
-\begin{description}
 \item[Training the neural network:]\hfill
@@
   \item {\tt https://github.com/gardner-lab/syllable-detector-learn}
+    \item DOI: 10.5281/zenodo.437555
   \end{trivlist}
@@
   \begin{description}
-  \item[Installation:]\hfill
+  \item[Installation:]\hfill\\
+    See {\tt README.md} in the repository for complete, up-to-date instructions.
     \begin{itemize}
-    \item Clone the Github repository onto your computer.
-    \item You also need the following packages from MATLAB File Exchange somewhere in your MATLAB path:
+    \item Clone the Github repository onto your computer, and add it to your MATLAB path.
+    \item You also need the following packages from MATLAB File Exchange in your MATLAB path:
       \begin{description}
@@
   \begin{description}
-  \item[MATLAB:] {\tt https://github.com/gardner-lab/syllable-detector-matlab}
+
+  \item[MATLAB:]\hfill
+    \begin{trivlist}
+    \item {\tt https://github.com/gardner-lab/syllable-detector-matlab}
+    \item DOI: 10.5281/zenodo.437557
+    \end{trivlist}
     \begin{description}
@@
     \end{description}
-  \item[Swift:] {\tt https://github.com/gardner-lab/\dots\ syllable-detector-swift/releases/tag/0.3.2}
+
+  \item[Swift:]\hfill
+    \begin{trivlist}
+    \item {\tt https://github.com/gardner-lab/syllable-detector-swift}
+    \item DOI: 10.5281/zenodo.437559
+    \end{trivlist}
     \begin{description}
@@
     \end{description}
-  \item[LabVIEW:] {\tt https://github.com/gardner-lab/\dots\ syllable-detector-runtime-labview}
+
+  \item[LabVIEW:]\hfill
+    \begin{trivlist}
+    \item {\tt https://github.com/gardner-lab/syllable-detector-runtime-labview}
+    \item 10.5281/zenodo.437558 
+    \end{trivlist}
     \begin{description}

### 2017-03-22 main.tex
First mention of a bird name, lny64, is now called out as a bird name.

@@
   \includegraphics[width=\textwidth]{Fig1}
-  \caption{The spectrogram of the song of lny64, used as an example
+  \caption{The spectrogram of the song of the bird ``lny64'', used as an example
     throughout this paper.  This image was made by superposing the
@@
 detector performance (as is the case, for example, for all of the
-detection points for lny64), but occasionally a syllable will be seen
+detection points for the bird ``lny64''), but occasionally a syllable will be seen
 for which extending the window to 80 ms is helpful.

### 2017-03-24 main.tex
Final tweaks
Some minor adjustments based on a final readthrough. It looks great, I think it is ready to go. These changes just reflect suggestions.

* Refer to “AFP” as the “the AFP” (debatable)
* Consistent formatting of times (space before ms)
* Switched one instance of target times to target syllable, as it seemed a bit confusing on my read through
* Rearranged sentence on normalization
* Named Apple’s operating system
* Adjusted some performance information for the MATLAB detector
* Change “timeshifted” to “time-shifted”

Feel free to any of the suggestions or discard them all.

@@
 variability in songbird motor output is used to drive change in the
-song \cite{Tumer2007pitchshift}, and AFP produces a
+song \cite{Tumer2007pitchshift}, and the AFP produces a
 corrective signal to bias song away from those disruptions
@@
 response can be predicted by a simple mechanism
-\cite{Charlesworth2011learning}. AFP transfers the error signal to the
+\cite{Charlesworth2011learning}. The AFP transfers the error signal to the
 robust nucleus of the arcopallium (RA) using NMDA-receptor--mediated
@@
 song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
-performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8-ms
+performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8 ms
 segments \cite{Warren2011}.  This detector achieved a jitter of 4.5 ms, and false-negative and
@@
 detection syllable.  That element is, roughly, 1 if the input vector
-comes from the target time ($t=t^*$), 0 otherwise, for each target
+matches the target syllable ($t=t^*$), 0 otherwise, for each target
 syllable (of which there may be any number, although they increase
@@
 In order to present consistent and meaningful inputs to the neural
-network, we normalise the incoming data stream so that changes in the
-song structure of the sound are emphasised over changes in volume, and
-to maximise the effectiveness of the neural network's training
-algorithm.
+network and to maximise the effectiveness of the neural network's training
+algorithm, we normalise the incoming data stream so that changes in the
+song structure of the sound are emphasised over changes in volume.
 
@@
 This detector uses the Swift programming language and Core Audio
-interface included in Apple's operating systems.
+interface included in Apple's Mac OS X operating systems.
 
@@
 hardware and operating systems, and is perhaps the easiest to modify.
-While it did not perform as well on our test system as the other
+While it did not perform as well as the other
 implementations, the convenience may outweigh the timing
@@
 the double buffer required when sending triggers via the audio
-interface, reducing latency by half.
+interface, reducing latency by close to half.
 
@@
 points in one bird's song (lny64), and present accuracy results for
-seven more birds (the detector's timing does not appreciably vary among
+seven more birds (the detector's timing does not appreciably vary across
 birds). The results presented here are representative of real-world
@@
 (reversible by increasing $C_n$), while also providing a valid
-definition of false-positive rate for data streams that had not been
+definition of the false-positive rate for data streams that had not been
 segmented into song-sized chunks and thus allowing easy interpretation
@@
 timing variability within each song and across songs (which is
-responsible for a small amount of our measured jitter, but note that few of the songs are significantly timeshifted).
+responsible for a small amount of our measured jitter, but note that few of the songs are significantly time-shifted).
 

### 2017-03-29 main.tex
Update response2 document. Re-hyphenate compound adjective ;)
Respond to last comments from the editor: moved ylabel on Fig3, and changed one word.

@@
 song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
-performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8 ms
+performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8-ms
 segments \cite{Warren2011}.  This detector achieved a jitter of 4.5 ms, and false-negative and
@@
 
-\subsection{Learning a detector}
+\subsection{Training a detector}
 
@@
 \begin{figure}
-  % Width 8, height 3, font 120%
+  % Width 10, height 4, font 100%
   \includegraphics[width=\textwidth]{Fig3}
@@
 
-\bibliography{birds}
+%\bibliography{birds}
+\begin{thebibliography}{10}
+
+\bibitem{Kao2005}
+Kao MH, Doupe AJ, Brainard MS.
+\newblock Contributions of an avian basal ganglia--forebrain circuit to
+  real-time modulation of song.
+\newblock Letters to Nature. 2005 February;433:638--643.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v433/n7026/abs/nature03127.html}.
+
+\bibitem{Wang2008interhemispheric}
+Wang CZH, Herbst JA, Keller GB, Hahnloser RHR.
+\newblock Rapid Interhemispheric Switching during Vocal Production in a
+  Songbird.
+\newblock PLOS Biology. 2008 October;6(10).
+\newblock Available from:
+  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.0060250}.
+
+\bibitem{Keller2009}
+Keller GB, Hahnloser RHR.
+\newblock Neural processing of auditory feedback during vocal practice in a
+  songbird.
+\newblock Nature. 2009 January;457:187--190.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v457/n7226/abs/nature07467.html}.
+
+\bibitem{Tumer2007pitchshift}
+Tumer EC, Brainard MS.
+\newblock Performance variability enables adaptive plasticity of 'crystallized'
+  adult birdsong.
+\newblock Nature. 2007 December;450:1240--1244.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v450/n7173/abs/nature06390.html}.
+
+\bibitem{Andalman2009}
+Andalman AS, Fee MS.
+\newblock A basal ganglia-forebrain circuit in the songbird biases motor output
+  to avoid vocal errors.
+\newblock Proceedings of the National Academy of Sciences of the United States
+  of America. 2009 July;106(30):12518--12523.
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2709669/}.
+
+\bibitem{Charlesworth2011learning}
+Charlesworth JD, Tumer EC, Warren TL, Brainard MS.
+\newblock Learning the microstructure of successful behavior.
+\newblock Nature Neuroscience. 2011;14(3):373--380.
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3045469/}.
+
+\bibitem{Warren2011}
+Warren TL, Tumer EC, Charlesworth JD, Brainard MS.
+\newblock Mechanisms and time course of vocal learning and consolidation in the
+  adult songbird.
+\newblock Journal of Neurophysiology. 2011 October;106(4):1806--1821.
+\newblock Available from:
+  \url{http://jn.physiology.org/content/106/4/1806.full}.
+
+\bibitem{Canopoli2014}
+Canopoli A, Herbst JA, Hahnloser RHR.
+\newblock A Higher Sensory Brain Region Is Involved in Reversing
+  Reinforcement-Induced Vocal Changes in a Songbird.
+\newblock The Journal of Neuroscience. 2014;34(20):7018--7026.
+\newblock Available from:
+  \url{http://www.jneurosci.org/content/34/20/7018.full}.
+
+\bibitem{Leonardo1999}
+Leonardo A, Konishi M.
+\newblock Decrystallization of adult birdsong by perturbation of auditory
+  feedback.
+\newblock Nature. 1999 June;399:466--470.
+\newblock Available from:
+  \url{http://www.nature.com/nature/journal/v399/n6735/abs/399466a0.html}.
+
+\bibitem{Skocik2013}
+Skocik M, Kozhevnikov A.
+\newblock Real-time system for studies of the effects of acoustic feedback on
+  animal vocalizations.
+\newblock Frontiers in Neural Circuits. 2013 January;6(111).
+\newblock Available from:
+  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3539774/}.
+
+\bibitem{Poole2012}
+Poole B, Markowitz JE, Gardner TJ.
+\newblock The Song Must Go On: Resilience of the Songbird Vocal Motor Pathway.
+\newblock PLOS One. 2012 June;7(6).
+\newblock Available from:
+  \url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0038173}.
+
+\bibitem{Tchernichovski2011SAP}
+Tchernichovski O. Sound Analysis Pro User Manual; 2011.
+\newblock Available from: \url{http://soundanalysispro.com/manual-1}.
+
+\bibitem{Jacobsen2003SlidingDFT}
+Jacobsen E, Lyons R.
+\newblock The sliding DFT.
+\newblock IEEE Signal Processing Magazine. 2003 Mar;20(2):74--80.
+\newblock Available from:
+  \url{http://www.cmlab.csie.ntu.edu.tw/DSPCourse/reference/Sliding\%20DFT.pdf}.
+
+\bibitem{hkp}
+Hertz JA, Krogh AS, Palmer RG.
+\newblock Introduction to the Theory of Neural Computation.
+\newblock Perseus Books; 1991.
+
+\bibitem{Noll1970pitchdetection}
+Noll AM.
+\newblock Pitch determination of human speech by the harmonic product spectrum,
+  the harmonic sum spectrum, and a maximum likelihood estimate.
+\newblock In: Proceedings of the symposium on computer processing in
+  communications. vol.~19; 1970. p. 779--797.
+
+\end{thebibliography}
 

### 2017-07-17 main.tex
Remove author contributions. Revert subsubsection.

@@
 % PLOS wants \paragraph for some reason...
-\renewcommand{\subsubsection}[1]{\paragraph{#1}}
+%\renewcommand{\subsubsection}[1]{\paragraph{#1}}
 
@@
 %\Yinyang These authors contributed equally to this work.
-Author \textsuperscript{a}~contributed the original MATLAB and LabVIEW implementations and most of the manuscript.
-\textsuperscript{b}~contributed the Swift implementation and its description herein, helped debug the training code, and measured the timing for most of the detectors.
-\textsuperscript{c}~wrote several types of filter-bank software against which we compared, made suggestions for the neural network detector, and reformatted the MATLAB code for distribution.
-\textsuperscript{d}~provided a stream of suggested techniques, metrics, and goals.
+
+%Author \textsuperscript{a}~contributed the original MATLAB and LabVIEW %implementations and most of the manuscript.
+%\textsuperscript{b}~contributed the Swift implementation and its description herein, %helped debug the training code, and measured the timing for most of the detectors.
+%textsuperscript{c}~wrote several types of filter-bank software against which we %compared, made suggestions for the neural network detector, and reformatted the %MATLAB code for distribution.
+%\textsuperscript{d}~provided a stream of suggested techniques, metrics, and goals.
 
@@
 
-\subsection{Training a detector}
+\subsection{Learning a detector}
 
@@
 \begin{figure}
-  % Width 10, height 4, font 100%
+  % Width 8, height 3, font 120%
   \includegraphics[width=\textwidth]{Fig3}

### 2017-07-17 main.tex
Add raw tiff figures. They will be rescaled shortly.

@@
 \begin{figure}
-  % Width 8, height 3, font 120%
+  % Width 10, height 4, font 100%
   \includegraphics[width=\textwidth]{Fig3}
-  %\includegraphics[width=\textwidth]{accuracies-2hidden}
 
@@
 $t_{\textrm{fft}}=1.5$ ms.  Given the variety of the test syllables,
-the variability is surprisingly low, and is summarised here with 95\%
-confidence intervals (with n=6):
-\vspace{8pt}\par\noindent
+the variability is surprisingly low, and is summarised in
+Table~\ref{table:TimingVsSyllable}.
+
+\begin{table}
   \begin{tabular}{r|cc}
@@
   \end{tabular}
-\vspace{8pt}\par\noindent
+  \caption{Latency and jitter variability (95\% confidence) for lny64's six test syllables.}
+  \label{table:TimingVsSyllable}
+\end{table}
+
 The negative latency is due to the way in which the network responds
@@
 implementations for the $\delta$-syllable and lny64's $t^*_4$, again with
-$t_\mathrm{fft}=1.5$ ms.  Results are as follows:
-\vspace{8pt}
+$t_\mathrm{fft}=1.5$ ms.  Results are shown in Table~\ref{table:TimingVsDetector}.
+These data are also plotted in \fig{fig:TimingVsDetector}, and
+\fig{fig:timing} gives a more detailed view of what the timing curves
+look like for the five implementations of our detector on lny64's $t^*_4$.
+
+\begin{table}
   \begin{tabular}{l|c|c|c|c}
-  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird: $t^*_4$} \\
+    & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{lny64: $t^*_4$} \\
     Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
@@
   \end{tabular}
-\vspace{8pt}\par\noindent
-These data are also plotted in \fig{fig:TimingVsDetector}, and
-\fig{fig:timing} gives a more detailed view of what the timing curves
-look like for the five implementations of our detector on $t^*_4$.
+  \caption{Latency and jitter for each of our detector implementations on the synthetic $\delta$-syllable and on syllable $t^*_4$ from lny64.}
+  \label{table:TimingVsDetector}
+\end{table}
 
@@
 We would like to thank Winthrop F.~Gillis, William A.~Liberti, and Sanne Moorman for
-their improvements to the draft.  This work was funded by NIH grants
-5R01NS089679-02 and 5U01NS090454-02.  The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+their improvements to the draft.
+%  This work was funded by NIH grants
+%5R01NS089679-02 and 5U01NS090454-02.  The funders had no role in study %design, data collection and analysis, decision to publish, or preparation of the %manuscript.
 

### 2017-07-17 main.tex
Remove includegraphics from the paper (PLOS doesn't want it), and tag all figures as 300dpi.
Whoops, remove author credit assignment fully.
Minor wording change.

@@
 \\
-Ben Pearre\textsuperscript{1,a,\textcurrency},
-L.~Nathan Perkins\textsuperscript{1,b},
-Jeffrey E.~Markowitz\textsuperscript{2,c},
-Timothy J.~Gardner\textsuperscript{1,d}
+Ben Pearre\textsuperscript{1,\textcurrency},
+L.~Nathan Perkins\textsuperscript{1},
+Jeffrey E.~Markowitz\textsuperscript{2},
+Timothy J.~Gardner\textsuperscript{1}
 \\
@@
   % Width: 8, Height: 3, Fonts scale: 120%
-  \includegraphics[width=\textwidth]{Fig1}
+  %\includegraphics[width=\textwidth]{Fig1}
   \caption{The spectrogram of the song of the bird ``lny64'', used as an example
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{Fig2}
+  %\includegraphics[width=\textwidth]{Fig2}
   % 9x10 inches
@@
   % Width 10, height 4, font 100%
-  \includegraphics[width=\textwidth]{Fig3}
+  %\includegraphics[width=\textwidth]{Fig3}
 
@@
 \begin{figure}
-  \includegraphics[width=\textwidth]{Fig4}
+  %\includegraphics[width=\textwidth]{Fig4}
   \caption{Timing varies as the FFT frame interval changes.  Here we
@@
   \begin{center}
-    \includegraphics[width=8cm]{Fig5}
+    %\includegraphics[width=8cm]{Fig5}
   \end{center}
@@
 implementations for the $\delta$-syllable and lny64's $t^*_4$, again with
-$t_\mathrm{fft}=1.5$ ms.  Results are shown in Table~\ref{table:TimingVsDetector}.
-These data are also plotted in \fig{fig:TimingVsDetector}, and
+$t_\mathrm{fft}=1.5$ ms.  Results are shown in Table~\ref{table:TimingVsDetector} and \fig{fig:TimingVsDetector}.
 \fig{fig:timing} gives a more detailed view of what the timing curves
@@
   \begin{center}
-    \includegraphics[width=8cm]{Fig6}
+    %\includegraphics[width=8cm]{Fig6}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=8cm]{Fig7}
+    %\includegraphics[width=8cm]{Fig7}
   \end{center}

### 2017-07-20 S1_Appendix_A.tex
Moved the Resources into a separate file, reformatted with sections, added links. Whew, they've actually accepted this one!

--- 2017-07-17 main.tex
@@
 % Use adjustwidth environment to exceed column width (see example table in text)
-\usepackage{changepage}
+%\usepackage{changepage}
 
@@
 % textcomp package and marvosym package for additional characters
-\usepackage{textcomp,marvosym}
+%\usepackage{textcomp,marvosym}
 
@@
 % cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
+%\usepackage{cite}
 
 % Use nameref to cite supporting information files (see Supporting Information section for more info)
-\usepackage{nameref,hyperref}
+\usepackage{nameref}
+\usepackage[colorlinks=true,urlcolor=blue]{hyperref}
 
 % line numbers
-\usepackage[right]{lineno}
+%\usepackage[right]{lineno}
 
@@
 % rotating package for sideways tables
-\usepackage{rotating}
+%\usepackage{rotating}
 
 % Remove comment for double spacing
-\usepackage{setspace} 
-\doublespacing
+%\usepackage{setspace} 
+%\doublespacing
 
@@
 \raggedright
-\setlength{\parindent}{0.5cm}
-\textwidth 5.25in 
-\textheight 8.75in
+%\setlength{\parindent}{0.5cm}
+%\textwidth 5.25in 
+%\textheight 8.75in
 
@@
 % Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
+%\makeatletter
+%\renewcommand{\@biblabel}[1]{\quad#1.}
+%\makeatother
 
@@
 % Header and Footer with logo
-\usepackage{lastpage,fancyhdr,graphicx}
-\usepackage{epstopdf}
-\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
-\rfoot{\thepage/\pageref{LastPage}}
-\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
-\fancyheadoffset[L]{2.25in}
-\fancyfootoffset[L]{2.25in}
-\lfoot{\sf PLOS}
+%\usepackage{lastpage,fancyhdr,graphicx}
+%\usepackage{epstopdf}
+%\pagestyle{myheadings}
+%\pagestyle{fancy}
+%\fancyhf{}
+%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
+%\rfoot{\thepage/\pageref{LastPage}}
+%\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
+%\fancyheadoffset[L]{2.25in}
+%\fancyfootoffset[L]{2.25in}
+%\lfoot{\sf PLOS}
 
@@
 
-\newcommand{\lorem}{{\bf LOREM}}
-\newcommand{\ipsum}{{\bf IPSUM}}
+%\newcommand{\lorem}{{\bf LOREM}}
+%\newcommand{\ipsum}{{\bf IPSUM}}
 
@@
 \begin{document}
-\vspace*{0.35in}
-
-% Title must be 250 characters or less.
-% Please capitalize all terms in the title except conjunctions, prepositions, and articles.
-\begin{flushleft}
-{\Large
-\textbf\newline{A Fast and Accurate Zebra Finch Syllable Detector}
-}
-\newline
-% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
-\\
-Ben Pearre\textsuperscript{1,a,\textcurrency},
-L.~Nathan Perkins\textsuperscript{1,b},
-Jeffrey E.~Markowitz\textsuperscript{2,c},
-Timothy J.~Gardner\textsuperscript{1,d}
-\\
-\bigskip
-\textsuperscript{1} Department of Biology, Boston University, Boston, Massachusetts, United States of America\\
-\textsuperscript{2} Department of Neurobiology, Harvard Medical School, Boston, Massachusetts, United States of America
-\\
-\bigskip
-
-% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
-% 
-% Remove or comment out the author notes below if they aren't used.
-%
-% Primary Equal Contribution Note
-%\Yinyang These authors contributed equally to this work.
-
-%Author \textsuperscript{a}~contributed the original MATLAB and LabVIEW %implementations and most of the manuscript.
-%\textsuperscript{b}~contributed the Swift implementation and its description herein, %helped debug the training code, and measured the timing for most of the detectors.
-%textsuperscript{c}~wrote several types of filter-bank software against which we %compared, made suggestions for the neural network detector, and reformatted the %MATLAB code for distribution.
-%\textsuperscript{d}~provided a stream of suggested techniques, metrics, and goals.
-
-% Additional Equal Contribution Note
-% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
-%\ddag These authors also contributed equally to this work.
-
-% Current address notes
-%\textcurrency a Insert current address of first author with an address update
-% \textcurrency b Insert current address of second author with an address update
-% \textcurrency c Insert current address of third author with an address update
-
-% Deceased author note
-%\dag Deceased
-
-% Group/Consortium Author Note
-%\textpilcrow Membership list can be found in the Acknowledgments section.
-
-% Use the asterisk to denote corresponding authorship and provide email address in note below.
-\textsuperscript{\textcurrency} Corresponding author: bwpearre@bu.edu (BP)
-
-\end{flushleft}
-
-\reversemarginpar
-%%% End PLOS header template
-
-\begin{abstract}
-  The song of the adult male zebra finch is strikingly stereotyped.
-  Efforts to understand motor output, pattern generation, and learning
-  have taken advantage of this consistency by investigating the bird's
-  ability to modify specific parts of song under external cues, and by
-  examining timing relationships between neural activity and vocal
-  output.  Such experiments require that precise moments during song
-  be identified in real time as the bird sings.  Various
-  syllable-detection methods exist, but many require special hardware,
-  software, and know-how, and details on their implementation and
-  performance are scarce.  We present an accurate, versatile, and fast
-  syllable detector that can control hardware at precisely timed
-  moments during zebra finch song.  Many moments during song can be
-  isolated and detected with false negative and false positive rates
-  well under 1\% and 0.005\% respectively. The detector can run on a
-  stock Mac Mini with triggering delay of less than a millisecond and
-  a jitter of $\sigma\approx 2$ milliseconds.
-\end{abstract}
-
-\linenumbers
-
-\section{Introduction}
-
-The adult zebra finch ({\em Taeniopygia guttata}) sings a song made up
-of 2--6 syllables, with longer songs taking on the order of a
-second. The song may be repeated hundreds of times per day, and is almost
-identical each time.  Several brain areas reflect this consistency
-in highly stereotyped neural firing patterns, which makes
-the zebra finch one of the most popular models for the study of the
-neural basis of learning, audition, and control.
-
-If precise moments in song can reliably be detected quickly enough to
-trigger other apparatus during singing, then this consistency of
-behaviour allows a variety of experiments.  A common area of study
-with song-triggered experiments is the anterior forebrain pathway
-(AFP), a homologue of mammalian basal ganglia consisting of a few
-distinct brain areas concerned with the learning and production of
-song.  For example, stimulation of the lateral magnocellular nucleus
-of the anterior nidopallium (LMAN)---the output nucleus of the
-AFP---at precisely timed moments during song showed that this area
-controls specific variables in song output \cite{Kao2005}.
-Song-synchronised stimulation of LMAN and the high vocal centre (HVC)
-in one hemisphere or the other showed that control of song rapidly
-switches between hemispheres \cite{Wang2008interhemispheric}.
-Feedback experiments have shown that Field L and the caudolateral
-mesopallium may hold a representation of song against which auditory
-signals are compared \cite{Keller2009}.  The disruption by white noise
-of renditions of a syllable that were slightly above (or below) the
-syllable's average pitch showed that the apparently random natural
-variability in songbird motor output is used to drive change in the
-song \cite{Tumer2007pitchshift}, and the AFP produces a
-corrective signal to bias song away from those disruptions
-\cite{Andalman2009}. The song change is isolated to within roughly 10
-milliseconds (ms) of the stimulus, and the shape of the learned
-response can be predicted by a simple mechanism
-\cite{Charlesworth2011learning}. The AFP transfers the error signal to the
-robust nucleus of the arcopallium (RA) using NMDA-receptor--mediated
-glutamatergic transmission \cite{Warren2011}. The course of song
-recovery after applying such a pitch-shift paradigm showed that the
-caudal medial nidopallium is implicated in memorising or recalling a
-recent song target, but in neither auditory processing nor directed
-motor learning \cite{Canopoli2014}.
-
-% THEIR WORDS: (NCM, a high auditory area) impaired recovery of the original pitch even several weeks after withdrawal of the reinforcing stimuli.
-Because NCM lesions spared both successful noise-avoidance behavior and birds' auditory discrimination ability, our results show that NCM is not needed for directed motor changes or for auditory discriminative processing, but is implied in memorizing or recalling the memory of the recent song target.
-
-Despite the power and versatility of vocal feedback experiments, there
-is no standard syllable detector.  Desiderata for such a detector include:
-\begin{description}
-\item[Accuracy:] How often does the system produce false positives or false negatives?
-\item[Latency:] The average delay between the target syllable being sung and the detection.
-\item[Jitter:] The amount that latency changes from instance to instance of song.  Our measure of jitter is the standard deviation of latency.
-\item[Versatility:] Is detection possible at ``difficult'' syllables?
-\item[Ease of use:] How automated is the process of programming a detector?
-\item[Cost:] What are the hardware and software requirements?
-\end{description}
-
-A variety of syllable-triggering systems have been used, but few have
-been documented or characterised in detail.  In 1999,
-detection was achieved by a group of IIR filters with hand-tuned logical
-operators \cite{Leonardo1999}.  The system had a latency of 50 or 100 ms,
-and accuracy and jitter were not reported.  As access to
-computational resources has improved, approaches have changed: in
-2009, hand-tuned filters were implemented on a Tucker-Davis Technologies digital signal processor, bringing latency down to around 4 ms \cite{Andalman2009}.
-But as with other filter-bank techniques, it
-is not strictly a syllable detector but rather a pitch and timbre
-detector---it cannot identify a frequency sweep, or distinguish a
-short chirp from a long one---and thus requires careful selection of
-target syllables.  Furthermore, the method is neither inexpensive nor,
-based on our experience with a similar technique, accurate.
-2009 saw the application of a neural network to a spectral image of
-song \cite{Keller2009}.  They reported a jitter of 4.3 ms, but further implementation and
-performance details are not available.  In 2011, stable portions of syllables were matched to spectral templates in 8-ms
-segments \cite{Warren2011}.  This detector achieved a jitter of 4.5 ms, and false-negative and
-false-positive rates of up to 2\% and 4\% respectively.  Hardware
-requirements and ease of use were not reported.  In 2013, spectral
-images of template syllables were compared to song using a correlation coefficient \cite{Skocik2013}.  With a
-fast desktop (Intel i7 six-core) running Linux and equipped with a
-National Instruments data acquisition card, it boasts a hardware-only
-(not accounting for the time taken to compute a match with a syllable)
-latency and jitter of just a few microseconds, and the detection
-computation they use should not much increase that.  They reported
-false-negative rates around 4\% and 7\% for zebra finches and
-Bengalese finches respectively, measured on a small dataset.  In much
-other work, a syllable detector is alluded to, but not described.
-
-We developed a standalone detector that learns to match moments in the
-song using a neural network applied to the song spectrogram, and
-outputs a TTL pulse (a brief 5-volt pulse) at the chosen moment. The
-approach consists of three steps:
-
-\begin{enumerate}
-\item Record and align a corpus of training songs.  The technique has
-  been published in \cite{Poole2012}. As few as 200 songs can yield
-  acceptable results, but here we standardise on 1000-song training
-  sets.
-\item Choose one or more instants in the song that should create a trigger event, and train a
-  neural network to recognise them. This step is carried out offline.  While any neural network software would produce similar results, we used MATLAB 2015b's neural network toolbox.
-\item Once trained and saved, the neural network is used by a realtime
-  detection programme that listens to an audio signal and indicates detection of the target
-  syllables via a TTL pulse.  We present three detection implementations, in MATLAB, LabVIEW, and Swift, that trade off hardware requirements, ease of maintenance, and performance.
-\end{enumerate}
-This method makes the following contributions:
-\begin{itemize}
-\item Fast: sub-millisecond latencies, with jitter around 2 ms.
-\item Accurate: false negative rates under 1\% and false positive rates under 0.005\% for a variety of syllables.
-\item State-of-the-art performance with default parameters.
-\item Requires almost no programming experience.
-\item Runs on inexpensive hardware.
-\item Described in detail here, with reference implementations provided and benchmarked.
-\end{itemize}
-
-%We present the method in Section~\ref{sec:method}.  Section~\ref{sec:results} presents our measurements.
-%We
-%conclude in Section~\ref{sec:conclusion}, and point to software
-%resources in Appendix~\ref{sec:resources}.
-
-\section{Materials and Methods}
-\label{sec:method}
-
-\subsection{Learning a detector}
-
-We begin with a few hundred recordings of a given bird's song, as well
-as calls, cage noise, and other non-song audio data.  Male zebra finch
-song is most highly stereotyped when a female is present (``directed
-song'') and slightly more variable otherwise (``undirected''); we
-train and test on undirected song since this is both more commonly
-studied and more challenging.  Recordings were made inside a
-sound-isolating chamber in which was mounted a recording microphone
-(Audio-Technica AT803), using methods similar to those described in
-\cite{Tchernichovski2011SAP}, Chapter 2.  The songs were time-aligned
-as described in \cite{Poole2012}.
-
-We rely on two circular buffers:
-\begin{description}
-\item[Audio buffer:] This contains the most recent audio samples, and
-  is of the length required to compute the Fast Fourier Transform
-  (FFT)---usually 256 samples.
-\item[FFT buffer:] The results of each new FFT are placed here.  It
-  contains the $n_\textrm{fft}$ most recent FFTs, which will serve as
-  inputs to the neural network (described below).
-\end{description}
-
-Audio is recorded at some sample rate $1/t_\mathrm{sample}$ (for example, 44.1 kHz), and new data are appended to the circular audio buffer.
-
-The spectrogram is computed at regular intervals---the useful range
-seems to be roughly 1--5 milliseconds, which we refer to as the frame
-interval $t_\textrm{fft}$.  At each frame, a spectrum is computed from
-the most recent 256 audio samples in the audio buffer, and the result
-is appended to the FFT buffer.  For example, if $t_\textrm{fft}=1$ ms
-and the recording sample rate $1/t_\textrm{sample}=40$ kHz, then in
-order to compute a new FFT, $40000\cdot 0.001=40$ new audio samples
-must be appended to the audio buffer, the FFT is computed using the
-most recent 256 samples in that buffer, and the result is appended to
-the FFT buffer.  Over time, the successive spectra will look something
-like \fig{fig:song}.
-
-\begin{figure}
-  % Width: 8, Height: 3, Fonts scale: 120%
-  \includegraphics[width=\textwidth]{Fig1}
-  \caption{The spectrogram of the song of the bird ``lny64'', used as an example
-    throughout this paper.  This image was made by superposing the
-    spectra of our 2818 aligned songs.  Our example detection points,
-    $t^*_1\ldots t^*_6$, are shown as red lines, with example recognition
-    regions of 30 ms $\times$ 1--8 kHz marked as rectangles.}
-  \label{fig:song}
-\end{figure}
-
-This results in time being discretised into chunks of length
-$t_\mathrm{fft}$.  Because each Fourier transform computation contains
-a small number $n_\mathrm{fft}$ of new audio samples (in the example
-above, 40 new samples and $256-40$ that have already been examined),
-we tried implementing the Sliding Discrete Fourier transform (SDFT)
-\cite{Jacobsen2003SlidingDFT}.  This allows
-$t_\textrm{fft}=t_\textrm{sample}$.  Practically, the operating system
-retrieves new audio samples from the hardware several at a time, so
-the full benefit of the SDFT is difficult to see in practice.
-Furthermore, we found that FFT implementations are sufficiently highly
-optimised that discretising time into chunks of $t_\mathrm{fft}$ as we
-have done produced similar results with simpler software.
-
-When using consumer audio hardware that operates best at 44.1 kHz, the
-requested frame interval may not line up with the sample rate, so the
-actual frame interval may be different from the intended.  For
-example, at 44.1 kHz a 1-ms frame interval requires a new FFT every
-44.1 samples.  This must be rounded to 44 samples, resulting in
-$t_\textrm{fft}=\lfloor 44.1 \rceil / 44.1 \approx 0.9977$ ms.
-
-One or more target moments during the song must be chosen.  Our
-interface presents the time-aligned spectrogram averaged over training
-songs, and requires manual input of the target times, $t^*$.  Then we
-assemble the training set from the song data, train the network,
-compute optimal output unit thresholds, and save the network object
-and an audio test file.
-
-\subsubsection{Recognition region}
-\label{sec:recognitionregion}
-
-The neural network's inputs are the FFT values from a rectangular
-region of the spectrogram covering a predefined range of frequency
-values $F$ (for example, 1--8 kHz) at some number of the most recent
-frames $n_\textrm{fft}$.  Any time $t$ falls within frame $\tau(t)$,
-and $t-t_\textrm{fft}$ falls within the previous frame, so the
-recognition region that the neural network receives as input consists
-of the spectrogram values over $F$ at $\tau(t)$ and those from the
-contiguous set of recent frames: $T = \{$ $\tau(t)$,
-$\tau(t-t_\textrm{fft})$, $\tau(t-{2t_\textrm{fft}})\ldots$
-$\tau(t-n_\textrm{fft}t_\textrm{fft})\}$.  Time windows of 30--50
-ms---the latter will yield $n_\textrm{fft} = |T|=\lfloor 50\textrm{
-  ms}/t_\textrm{fft}\rceil$ frames---of frequencies spanning 1--8 kHz
-generally work well.
-
-Six examples of chosen target moments in the song, with recognition
-regions $F =$ 1--8 kHz and $T=30$ ms, are shown in \fig{fig:song}.
-
-%% \subsubsection{Song micro-realignment}
-
-%% As conditions change, and especially during undirected song, syllable
-%% length and relative timing may vary slightly, which introduces
-%% variations in the precise timing of each syllable. The alignment
-%% software we use ensures that songs are aligned at the point midway
-%% through the song, but if the target syllable is not at that point, it
-%% is helpful to re-align the songs at the point of interest.  This may
-%% be accomplished by looking for peaks in the correlation of the
-%% time-domain signal with the song whose spectrogram is closest to
-%% average over the training set.\marginpar{I hope that's the right thing
-%%   to do. Seems to work, anyway\dots}
-
-\subsubsection{Building the training set}
-
-The training set is created in a fashion typical for neural networks:
-at each time frame $t$ the rectangular $|F|\times |T|$ recognition
-region in the spectrogram as of time $t$ is reshaped into a vector
-$\xi_t$, which will have length $|F||T|$ and contain the spectra in
-$F$ taken at all of the times in the set $T$: from
-$\tau(t-n_\textrm{fft}t_\textrm{fft})$ to $\tau(t)$.  These vectors
-are placed into a training matrix, $\Xi$, such that each column
-$\xi_t$ holds the contents of the recognition region---containing
-multiple frames from the spectrogram---as of one value of $t$.
-
-Training targets $y_t$ are vectors with one element for each desired
-detection syllable.  That element is, roughly, 1 if the input vector
-matches the target syllable ($t=t^*$), 0 otherwise, for each target
-syllable (of which there may be any number, although they increase
-training time, and in our implementations the number of distinct
-output pulses is constrained by hardware).  Since the song alignment
-may not be perfect, and due to sample aliasing, a
-strict binary target may ask the network to learn that, of two
-practically identical frames, one should be a match and the other
-not. Thus it is preferable to spread the target in time, such that at
-the target moment it is 1, and at neighbouring moments it is
-nonzero. We found that a Gaussian smoothing kernel around the target
-time with a standard deviation of 2 ms serves well.
-
-With inputs well outside the space on which a neural network has been
-trained, its outputs will be essentially random. In order to reduce
-the false positive rate it is necessary to provide negative training
-examples that include silence, cage noise, wing flapping, non-song
-vocalisations, and perhaps songs from other birds.  Although it will
-depend on the makeup of the non-song data, we have found that training
-with as low as a 1:1 ratio of non-song to song---or roughly 10
-minutes of non-song audio---yields excellent results on most birds.
-
-\subsubsection{Normalisation}
-
-In order to present consistent and meaningful inputs to the neural
-network and to maximise the effectiveness of the neural network's training
-algorithm, we normalise the incoming data stream so that changes in the
-song structure of the sound are emphasised over changes in volume.
-
-The first normalisation step is designed to eliminate differences in
-amplitude due to changes in the bird's location and other variations
-in recording.  Each recognition region vector $\xi_t$---during
-training, each {\em column} of the training matrix $\Xi$---is
-normalised using MATLAB's {\tt zscore} function $\hat{\xi}_t = (\xi_t
-- \mu_{\xi_t}) / \sigma_{\xi_t}$, so that the content of each window
-has mean $\mu_{\hat{\xi}_t}=0$ and standard deviation
-$\sigma_{\hat{\xi}_t}=1$.
-
-The second step is designed to ensure that the neural network's inputs
-have a range and distribution for which the training function can
-easily converge.  Each element $i$ of $\hat{\xi}$ is normalised across
-the entire training set---each {\em row} of $\Xi$---in the same way:
-$\check{\xi}^i = (\hat{\xi}^i -
-\mu_{\hat{\xi}^i})/\sigma_{\hat{\xi}^i}$, so that the values of that
-point across the whole training set have mean $\mu_{\check{\xi}^i}=0$
-and standard deviation $\sigma_{\check{\xi}^i}=1$.  This is
-accomplished during training by setting MATLAB's neural network
-toolbox normalisation scheme to {\tt mapstd}, and the scaling
-transform is saved as part of the network object used by the realtime
-detector so that it may be applied to unseen data at runtime.
-
-These two normalisation steps provide a set of inputs that are more
-robust to outliers and less likely to produce false positives during
-silence than other normalisation schemes, such as linear or
-L\textsubscript{2} normalisation.
-
-\subsubsection{Neural Networks}
-
-While any learned classifier might suffice, we chose a two-layer
-feedforward neural network.  In brief, our network takes an input
-vector $\xi_t$---as described above---and produces an output vector
-$y_t$, and when any element of $y_t$ is above a threshold (described
-below), the detector reports a detection event.  The network uses two
-matrices of weights, $W_0$ and $W_1$, and two vectors of biases, $b_0$
-and $b_1$.  The first takes the input $\xi_t$ to an intermediate
-stage---the ``hidden layer'' vector.  To each element of this vector
-is applied a nonlinear squashing function such as {\tt tanh}, and then
-the second weight matrix $W_1$ is applied.  This produces output
-$y_t$:
-\begin{equation*}
-  y_t = W_1 \tanh (W_0 \xi_t + b_0) + b_1
-\end{equation*}
-During the network's training phase, the elements of the matrices and
-bias vectors are learned by back-propagation of errors over a training
-set.  A more detailed explanation of neural networks may be found in
-\cite{hkp}.
-
-Essentially, after training, the network is available in the form of
-the two matrices $W_0$ and $W_1$ and the two vectors $b_0$ and $b_1$,
-and running the network consists of two matrix multiplications, two
-vector additions and the application of the squashing function.
-
-\subsubsection{Training the network}
-
-The network is trained using MATLAB's neural network toolbox, with
-Scaled Conjugate Gradient ({\tt trainscg}). We tried a variety of
-feedforward neural network geometries, from simple 1-layer perceptrons
-to geometries with many hidden nodes, as well as autoencoders. Perhaps
-surprisingly, even the former yields excellent results on many
-syllables, but a 2-layer perceptron with a very small hidden
-layer---with a unit count 2-4 times the number of target
-syllables---was a good compromise between accuracy and training
-speed. For more variable songs, deep structure-preserving networks may
-be more appropriate, but they are slow to train and unnecessary for
-zebra finch song.
-
-\subsubsection{Computing optimal output thresholds}
-\label{sec:optimalthresholds}
-After the network is trained, outputs of the network for any input are
-now available, and will be in (or, due to noisy inputs and imperfect
-training, close to) the interval (0, 1). We must choose a threshold
-above which the output is considered a positive detection. Finding the
-optimal threshold requires two choices. The first is the relative cost
-of false negatives to false positives, $C_n$. The second is the
-acceptable time interval: if the true event occurs at time $t$, and
-the detector triggers at any time $t\pm\Delta t$, then it is
-considered a correct detection. Then the optimal detection threshold
-is the number that minimises $\textrm{[false positives]}
-+C_n\cdot\textrm{[false negatives]}$ over the training set, using the
-definitions of false positives and negatives given in
-Section~\ref{sec:accuracy}. Since large portions of the cost function
-are flat, we use a brute-force linear search, which requires fractions of a
-second. For the results presented here, we have used
-$\Delta t=10$ ms, and arbitrarily set $C_n=1$.
-
-\subsubsection{De-bouncing}
-
-During runtime, the network may produce above-threshold responses to
-nearby frames.  Thus, after the first response, subsequent responses
-are suppressed for 100 ms.  However, for the accuracy measurements
-presented here, we used the un-de-bounced network response.
-
-\subsubsection{Our parameter choices}
-
-We used an FFT of size 256; a Hamming window; and chose a target
-spectrogram frame interval of $t_\textrm{fft}=1.5$ milliseconds,
-resulting in a true frame interval of $t_\textrm{fft}=\lfloor 1.5\cdot
-44.1\rceil /44.1\approx 1.4966$ ms.  We set the network's input space
-to 50 ms long, and to span frequencies from 1--8 kHz, which contains
-the fundamentals and several overtones of most zebra finch
-vocalisations.
-
-We found these parameters to work well across a variety of target
-syllables, but various other parameter sets yield results similar to
-those presented here.  Some of the parameters trade off detection
-accuracy or temporal precision vs.~training time. For example,
-decreasing the frame interval generally decreases both latency and
-jitter, but also increases training time. Sometimes the effects are
-syllable-specific: for example, using a 30-ms time window rather than
-50 ms speeds training while usually having a minimal effect on
-detector performance (as is the case, for example, for all of the
-detection points for the bird ``lny64''), but occasionally a syllable will be seen
-for which extending the window to 80 ms is helpful.
-
-\subsection{Realtime detection}
-
-The architecture of the realtime detector requires that the most
-recent $n_\textrm{fft}$ spectrograms be fed to the neural network
-every frame interval.  Audio samples from the microphone are appended
-to the circular audio buffer.  Every $t_\textrm{fft}$ seconds a new
-spectrogram is calculated by applying the Hamming window to the
-contents of the buffer, performing an FFT, and extracting the
-power. Outputs of the spectrogram from the target frequency band are
-appended to the circular FFT buffer.  The spectrograms are sent to a
-static implementation of the previously trained neural network.
-
-We tested three implementations of the realtime detector.  For all of
-these tests, we ran the detector processes under the operating
-systems' default schedulers and process priorities, running typical
-operating system daemon processes but no loads from user processes.
-The computers had ample memory resources.
-
-\subsubsection{Swift}
-
-This detector uses the Swift programming language and Core Audio
-interface included in Apple's Mac OS X operating systems.
-
-The Core Audio frameworks provide an adjustable hardware buffer size
-for reading from and writing to audio hardware (different from our two
-circular buffers). Tuning this buffer size provides a tradeoff between
-the jitter in the detection and the processor usage needed to run the
-detector. We used buffer sizes ranging from 8 samples (0.18 ms at 44.1
-kHz) to 32 samples (0.7 ms at 44.1 kHz) depending on the frame size 
-used by the detector.
-
-Vector operations---applying the Hamming window, the FFT, input
-normalisation, matrix multiplication, and the neural network's
-transfer functions---are performed using the Accelerate framework
-(vDSP and vecLib), which use modern vector-oriented processor
-instructions to perform calculations.
-
-When the neural network detects a match, it instructs the computer to
-generate a TTL pulse that can be used to trigger downstream hardware.
-This pulse can be either written to the computer's audio output buffer
-(again, in 8- to 32-sample chunks) or sent to a microcontroller
-(Arduino) via a USB serial interface. Sending the trigger pulse via
-the serial interface and microcontroller is noticeably faster (2.2 ms
-lower latency), likely due to the fact that the audio buffer goes
-through hardware mixing and filtering prior to output.
-
-The above code can be run on multiple channels of audio on consumer
-hardware (such as a 2014 Mac Mini) with little impact on CPU usage
-($<15\%$). Depending on the experimental needs, latency can potentially
-be further decreased (at the expense of processor usage) by adjusting the audio
-buffer sizes.
-
-We ran the Swift detector on a Late 2014 Mac Mini with a Intel Core i5
-processor at 2.6GHz  with 16 gigabytes of RAM, running Mac OS X 10.11.
-
-\subsubsection{LabVIEW}
-
-This implementation requires special software and hardware: LabVIEW
-from National Instruments---we used 2014 service pack 1---and a data acquisition card---we use the
-National Instruments PCI-6251 card on a PC with an Intel Core i5-4590
-processor at 3.7GHz (a relatively low-end machine) with 24 gigabytes
-of RAM, running Microsoft Windows 8.1 Pro and Windows 10.
-
-This implementation has several drawbacks: it requires expensive hardware and software from National Instruments (a data acquisition card and LabVIEW), Windows (we use hardware features of LabVIEW that are unavailable on MacOS or Linux), and due to the programming language it
-is difficult to modify and debug---indeed, a persistent bug in our
-implementation currently renders it substantially less accurate than
-the other detector implementations on some syllables. However, our
-test configuration proved itself capable of excellent performance, and further
-gains should be possible if the implementation were retargeted onto
-field-programmable gate array (FPGA) hardware---which would have the additional benefit of providing
-deterministic ``hard realtime'' guarantees---or just run on a faster
-desktop system.
-
-\subsubsection{MATLAB}
-
-This detector uses the built-in audio input and output hardware on a
-compatible computer.  We tested on a 2014 Mac Mini (the same machine used
-for the Swift detector described above) and 2015 Mac Pro.
-The Mac Pro does not have an audio input jack, so input was through an
-M-Audio MobilePre external USB audio interface.  Despite the faster
-processor, the latter system did not achieve higher performance than
-the former, due to USB data transfer overhead.
-
-Because of how MATLAB's DSP toolbox interfaces with audio hardware,
-there is a double buffer both reading from and writing to audio
-hardware. As a result, much of the code focuses on a lightweight audio
-hardware interface, in order to have the smallest achievable audio
-buffer. To help achieve this, the MATLAB implementation spreads data
-acquisition and processing across two threads, due to the higher
-computational overhead of the interpreted programming language.
-
-The most versatile implementation, MATLAB runs on a variety of
-hardware and operating systems, and is perhaps the easiest to modify.
-While it did not perform as well as the other
-implementations, the convenience may outweigh the timing
-performance penalty for some experiments.  Key to minimising jitter is
-the size of the audio buffer: on a 2014 Mac Mini running MATLAB 2015b
-the smallest buffer size that did not result in read overruns was
-about 4 ms.
-
-As with the Swift detector, it is also possible to modify the
-MATLAB implementation to generate the TTL pulse from a microcontroller
-(Arduino) controlled via a serial over USB interface. This eliminates
-the double buffer required when sending triggers via the audio
-interface, reducing latency by close to half.
-
-\subsection{Quantification}
-\label{sec:quantify}
-
-We measure accuracy and characterise timing for 6 randomly chosen
-points in one bird's song (lny64), and present accuracy results for
-seven more birds (the detector's timing does not appreciably vary across
-birds). The results presented here are representative of real-world
-performance in feedback experiments on many birds in our lab.
-
-\begin{table}
- \begin{tabular}{c|ccc}
-    Bird & Test songs = Test non-songs & Target time (ms) \\ \hline
-    lny64 & 1818 & 150:50:400 \\
-    lny46 & 1000 & 195 \\
-    lny42 & 1245 & 380 \\
-    lny4rb & 1000 & 250 \\
-    lr12 & 2000 & 180 \\
-    lr13 & 2000 & 300 \\
-    lr28 & 3000 & 140 \\
-    lr77 & 2000 & 320
-  \end{tabular}
-  \caption{For each bird, detectors were trained for the specified
-    timepoints, using 1000 songs and 1000 non-song samples of the same
-    length. The test set consisted of the remaining songs and an
-    equal number of non-songs.}
-  \label{table:datasizes}
-\end{table}
-
-Data consisted of extracted, aligned songs, and additional song-length
-samples of non-song (cage noise, calls, other birds' song). These were
-divided into training and test sets. For all birds, we used 1000 songs
-and 1000 non-song samples for training. The test set consisted of the remaining songs and an equal number of non-songs. Table~\ref{table:datasizes}
-lists the test dataset sizes and target detection points. For lny64,
-we arbitrarily chose six linearly spaced target times. For each of the
-other birds, we eyeballed one promising-looking detection point.
-
-The MATLAB neural network toolbox further divides our ``training'' set
-into internal training, validation, and test sets.  Because we do our
-own testing on a distinct data set, we instructed the
-toolbox to use 80\% of its input set for training, and the
-remaining 20\% for validation, with no hold-out test data.
-
-Because the data set consists of temporally aligned songs, the moment
-at which detection should occur during each song sample is available
-(albeit with minor variations due to the alignment process used
-\cite{Poole2012}).  The detector can be checked by comparing its
-response to the recorded training songs against these ``canonical''
-detection events.  To this end, besides the trained network object,
-our learning code produces an audio file consisting of all of the
-training data on the left audio channel and a delta function at each
-aligned moment of target syllable presentation on the right
-channel. Thus, when played on any audio player, the left channel may
-be provided as input to the detector, and the the detector's output
-pulses may be compared against the ``ground truth'' given by the
-canonical detection event provided by the right channel.
-
-\subsubsection{Accuracy}
-\label{sec:accuracy}
-
-We define the accuracy of the network based on its classification
-performance per frame. In order to avoid the apparent problem of every
-non-detected non-matching frame counting as a true negative, we also
-tried defining accuracy on a per-song basis, such that a song-length
-sample without the target syllable counted as a single true negative.
-Computing the optimal output thresholds on a per-frame rather than a
-per-song basis resulted in higher thresholds and thus a slight
-reduction in both the false-positive and true-positive rates
-(reversible by increasing $C_n$), while also providing a valid
-definition of the false-positive rate for data streams that had not been
-segmented into song-sized chunks and thus allowing easy interpretation
-of false-positive rates from calls, cage noise, etc.
-
-%% We define accuracy on a per-song basis, as follows:
-
-%% \begin{description}
-%%   \item[True positive:] A song for which the detector fires within
-%%     20ms of the intended moment.
-%%   \item[True negative:] So that every non-detected non-syllable does
-%%     not count, a true negative is a complete song during which the
-%%     detector did not fire at any point outside the true positive
-%%     region.
-%%   \item[False positive:] A detection event more than 20ms from the
-%%     target syllable.\marginpar{Or a song with $>0$ false pos?}
-%%   \item[False negative:] A song's target interval during which a
-%%     detection event should have occurred, but didn't.
-%% \end{description}
-
-The accuracy as defined above is used for computing the optimal
-thresholds above which the network's output should be interpreted as a
-match on the training data as described in
-\sref{sec:optimalthresholds}, for evaluation of the detectors on the
-training songs, and while live.
-
-\noprint{ Since the network will output values $o_t$ between 0 and 1
-  at each moment $t$ in an attempt to match the training output, the
-  optimal threshold $\tau\in[0,1]$ for the output neuron should be
-  computed.  Given the relative cost of false positives vs.~false
-  negatives $C$, and the acceptable time difference between target
-  syllable and correct output $\Delta t_d$, we compute the optimal
-  threshold for an output element according to the definitions above:
-\begin{eqnarray*}
-  \textrm{true positives}_\tau &=& \textrm{size of set}_{s\in \textrm{target songs}} o_t > \tau, \left| t \leq \Delta t_d \right| \\
-  \textrm{false negatives}_\tau &=& \textrm{size of set} {s\in\textrm{target songs}} - \textrm{size of set} \textrm{true positives} \\
-  \textrm{false positives}_\tau &=& \textrm{size of set}_{s\in \textrm{target songs}} o_t > \tau, \left| t > \Delta t_d \right| \\
-  \widehat{\tau} &=& \argmin_\tau C\textrm{false positive} + \textrm{false negatives}
-\end{eqnarray*}
-}
-
-\subsubsection{Timing}
-
-We evaluate the time taken from the complete presentation of each
-instance of the target syllable to the firing of the detector's TTL
-pulse.  For example, if the target trigger point is at 200 ms with
-respect to the corpus of aligned songs, and if the detection region is
-30 ms long, then when the detector has seen the region from 170--200
-ms, it should recognise the syllable.  If for a given presentation of
-the song the detector fires at 203 ms, we take the latency on that
-trial to be 3 ms.
-
-While playing the audio test file from any audio playback device, we
-used the TTL output from the ground-truth channel of the audio output
-as the trigger pulse for an oscilloscope, and compared it to the TTL
-pulse produced by the detector, which sees only the birdsong channel
-of the audio file. For this purpose we used a pulse generator (Philips
-PM 5715, with a listed latency of 50 ns, jitter of $\leq 0.1\%$ or 50
-ps, whichever is greater) to widen the detector's output spike to a
-number much larger than the jitter ($\sim 100$ ms).  This obviates
-pulse length variability in the output device by essentially
-discarding the falling edge of the output pulse.  The oscilloscope is
-then set to averaging mode (128-trigger average) in order to collect
-timing data. The canonical signal is the trigger at $t=0$, and the
-average of the detector's detection events will be seen as a
-low-to-high transition with form approximating the cumulative
-probability distribution function (CDF) of the detector's output in
-response to the chosen song event, with individual detection events
-visible as steps in this curve.
-
-We define latency as the time between the training target (during our tests, this is indicated by the canonical
-signal in the test file) and the corresponding detection event.  It is a helpful number,
-but not a critical one; due to the stereotyped song of the zebra finch, a detector with high but
-constant latency can often be trained to trigger at a point somewhat
-before the true moment of interest (for example, if triggering should
-occur at the very beginning of a syllable, the detector may be trained 
-to recognise the previous syllable and the gap thereafter).  Usually the 
-variability in latency is the more important number.  We define jitter as 
-the standard deviation of detection latency, measured over the test songs.
-
-When measuring timing, it is useful to compare against a theoretical
-optimal, in order to control for any imprecision in the song
-extraction and alignment of our real data inherent in the method we
-use (described in \cite{Poole2012}).  We propose two measures thereof:
-
-First we test the ``ideal'' latency and jitter in the time shift used
-in calculating new columns of the spectrogram. By passing a recorded
-audio sequence into the detector offline, and assuming no input or
-output latency, we compare how many additional audio samples beyond
-the syllable are needed before the spectrogram can match the inputs
-needed to trigger the neural network.  This latency reflects the FFT
-size used for calculating the spectrogram, the FFT time shift between
-columns in the spectrogram, and the width of the Gaussian smoothing
-kernel applied to the ground truth data when training the neural
-network, but ignores computation times, audio buffers, and other
-operating system overhead.
-
-Next we use a ``$\delta$-syllable'' consisting of a $\delta$ function
-in the time domain, and train the network to trigger 5 ms after this
-pulse.  This song is fed into the live detector.  The results for this
-measurement show the latency and jitter inherent in the complete
-detector including audio read and write buffers, classifier
-computation, and FFT time shift aliasing, but excluding imprecision in
-the song alignment as well as detection timing effects due to
-differences between instances of the bird's song.
-
-Finally, we measure latency on real bird song aligned as in
-\cite{Poole2012}.  This extracts and aligns the centres of the sampled
-songs, and the canonical signal is given with respect to that
-alignment.  These timing measurements reflect not only detector
-performance, but also the variability of the bird's song.
-
-\subsection{Ethics}
-
-All procedures were approved by the Institutional Animal Care and Use
-Committee of Boston University (protocol number 14-029).
-
-\section{Results}
-\label{sec:results}
-
-\fig{fig:song} shows a typical song. The image is produced by
-averaging the aligned spectrograms for the bird lny64.  Six target
-trigger points are shown for this song, at 150, 200, 250, 300, 350,
-and 400 milliseconds after the beginning of the aligned samples, as
-indicated by the red lines. In discussion, we will refer to these six
-as $t^*_1\ldots t^*_6$ respectively.  To avoid overlap in the graphic, the rectangles show
-a shorter recognition window than our standard: here 30 ms, but still
-spanning 1--8 kHz, for each target time in the song.
-
-The first two triggers, $t^*_1$ at 150 ms and $t^*_2$ at 200 ms, are
-far from pure tones, but are closer to bandpass-filtered white noise.  They would be
-difficult for a harmonic-stack filter bank to detect, and especially
-to pinpoint in time.  $t^*_3$ (250 ms) and $t^*_6$ (400 ms) are rich
-in overtones and contain downward slides with changing timbres.
-$t^*_4$ (300 ms) occurs near the beginning of a more typical harmonic
-stack amenable to a variety of recognition techniques, although
-consistently detecting a point partway through the syllable demands a
-detector that can use multiple time steps.  $t^*_5$ (350 ms) shows a
-steady pitch followed by a complex changing tone structure.
-
-An example of the detector's output for three of these syllables is
-shown in \fig{fig:detection_raster}.  The total audio energy of each
-song is shown as a single grayscale row.  By sorting according to time
-of each syllable's detection, the rest of the song is shown in the
-context of the timing for that moment.  This gives an intuition of the
-timing variability within each song and across songs (which is
-responsible for a small amount of our measured jitter, but note that few of the songs are significantly time-shifted).
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{Fig2}
-  % 9x10 inches
-  % Fonts: Helvetica, scaled 130%
-  \caption{Each plot shows one network output unit's responses to all
-    2818 presentations of lny64's song shown in \fig{fig:song}.  We
-    show only the syllables $t^*_1$, $t^*_4$, and $t^*_6$, and we do
-    not show the non-response to presentation of non-song. The
-    horizontal axis is time relative to the beginning of the aligned
-    song, and the vertical axis is an index for the 2818 individual
-    song presentations. The grey shading shows the audio amplitude of
-    song Y at time T. Detection events on training songs are shown in
-    cyan, with detections of unseen test songs in red.  To provide an
-    intuition of intra-song variability, songs have been stably sorted
-    by the time of detection events; thus, each of the three detection
-    graphs shows the songs in a different order.}
-  \label{fig:detection_raster}
-\end{figure}
-
-\subsection{Accuracy}
-
-We allowed our training software to allocate our default of 4 hidden
-units per syllable, and computed a new FFT every 1.5 ms.  Because our
-timing test files are designed for our stereo playback software,
-allowing only one channel for the ground-truth pulse, we trained one
-detector for each syllable for the following results.  In order to
-eliminate the large number of variables involved in microphone, cage,
-playback and re-digitising, we evaluated the neural network's accuracy
-directly on the digitised recording.  When training and runtime data
-are gathered on the same hardware setup, this is the digital signal
-that the detector will see.
-%Accuracy \marginpar{Should be!
-Verify!
-And/or explain that there is a bug in the LabVIEW implementation, if there in fact is.
-ALSO need to present a picture of the test syllable or etc \dots} is identical across the three detectors.
-\noprint{
-  Typical confusion matrices for these syllables are:
-
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_1$
-    & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.96451\% & 0.000298\%\\ 
-    neg & 0.0355\% & 99.99970\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_2$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.92903\% & 0.000596\%\\ 
-    neg & 0.0710\% & 99.99940\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_3$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.68062\% & 0.00194\%\\ 
-    neg & 0.319\% & 99.99806\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_4$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.82257\% & 0.00239\%\\ 
-    neg & 0.177\% & 99.99761\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_5$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.78708\% & 0\\ 
-    neg & 0.213\% & 100\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-  \begin{tabular}{r|cc}
-    $t^*_6$ & \multicolumn{2}{c}{True} \\ 
-    & pos & neg \\ 
-    \hline  Detected pos & 99.71611\% & 0.00552\%\\ 
-    neg & 0.284\% & 99.99448\%\\ 
-  \end{tabular}
-  \vspace{8pt}\par\noindent
-}
-
-Accuracies are shown in \fig{fig:accuracies} (accuracies for the synthetic $\delta$-function
-songs were always 100\%).  For each test syllable we trained 100
-different networks, which differ from each other both in the random
-initialisation of network weights and in the random subset of the
-song-and-nonsong corpus that was used for training.  Each network's
-accuracy is shown as a single point in each chart in
-\fig{fig:accuracies}, and the means are shown in
-Table~\ref{table:accuracies}.  Evidently some syllables are easier to
-identify reliably than others (for example, for lny64, $t^*_3$ (250 ms) and $t^*_6$ (400 ms) are
-occasionally confused due to their similar appearance, as could be
-guessed from \fig{fig:song}); thus, if the experimental design
-permits, it can be worthwhile to choose an easily identifiable
-syllable.  Furthermore, as can be seen from \fig{fig:accuracies}, for
-difficult syllables the training process occasionally yields a network
-that performs unusually poorly.  It is easy to spot these networks by
-their performance either as measured as part of the training process
-or on the test audio file, perhaps taking the best of a few training
-runs.
-
-\begin{figure}
-  % Width 8, height 3, font 120%
-  \includegraphics[width=\textwidth]{Fig3}
-  %\includegraphics[width=\textwidth]{accuracies-2hidden}
-
-  \caption{Accuracy variability over 100 different training runs for
-    each of the test detection points.  Each dot shows the test-set
-    accuracy for an independently trained detector.  Because the
-    horizontal positions have been randomised slightly so as not to
-    occlude same-valued measurements, test syllable is also indicated
-    by colour.  The means are given in Table~\ref{table:accuracies}.}
-  \label{fig:accuracies}
-\end{figure}
-
-\begin{table}
-  % replot_accuracies_concatanated('tex', true) for confusion_log_perf.txt (1.5 ms).
-  \begin{tabular}{lc|cc}
-    Bird & Target time (ms) & \% True positives & \% False positives \\ \hline
-lny64 & 	150 $[t^*_1]$ & 	99.66 & 	0.0026 \\ % /Volumes/Data/song/lny64/confusion_log_perf.txt
-      & 	200 $[t^*_2]$ & 	99.75 & 	0.00052 \\ %                                               
-      & 	250 $[t^*_3]$ & 	99.79 & 	0.00051 \\ %                                               
-      & 	300 $[t^*_4]$ & 	99.70 & 	0.00077 \\ %                                               
-      & 	350 $[t^*_5]$ & 	99.73 & 	0.00022 \\ %                                               
-      & 	400 $[t^*_6]$ & 	99.52 & 	0.0010 \\ %                                                
-lny42 & 	380 & 	99.09 & 	0.0011 \\ % /Volumes/Data/song/lny42/confusion_log_perf.txt          
-lny46 & 	140 & 	99.11 & 	0.0036 \\ % /Volumes/Data/song/lny46/confusion_log_perf.txt          
-lny4rb & 	250 & 	99.37 & 	0.0026 \\ % /Volumes/Data/song/lny4rb/confusion_log_perf.txt        
-lr12 & 	180 & 	99.52 & 	0.0033 \\ % /Volumes/Data/song/lr12/all/confusion_log_perf.txt        
-lr13 & 	300 & 	99.83 & 	0.0013 \\ % /Volumes/Data/song/lr13/all/confusion_log_perf.txt        
-lr28 & 	140 & 	99.73 & 	0.0028 \\ % /Volumes/Data/song/lr28/all/confusion_log_perf.txt        
-lr77 & 	320 & 	99.58 & 	0.0028  % /Volumes/Data/song/lr77/all/confusion_log_perf.txt          
-  \end{tabular}
-  \caption{Mean values for the detection accuracies shown in \fig{fig:accuracies}.}
-  \label{table:accuracies}
-\end{table}
-
-We repeated the above experiment using only 2 hidden units per
-syllable.  Training under this condition is much faster, but the mean
-true positive rate per syllable decreases by an average of 18\% across
-our test syllables, and the mean false positive rate increases by
-roughly 23\%.
-
-Changing the FFT interval to 2 ms significantly reduces the
-computational requirement during learning. For most syllables, results
-were identical, but some syllables are slightly easier to detect and
-some slightly more difficult, with no significant difference on
-average.
-
-\subsection{Timing}
-
-We evaluate variability in timing performance across three variables:
-FFT frame interval; syllable choice; and detector implementation.
-
-\subsubsection{FFT frame interval}
-
-Detector latency and jitter depend on the FFT frame rate.  Our
-1.5-ms--frame default is a compromise: shorter frames increase the
-precision of timing, but also increase computational requirements both
-during training and at runtime.  \fig{fig:TimingVsFrame} shows how
-these numbers vary over a useful range of frame rates on our ideal
-detector, the Swift detector with serial output, and for the LabVIEW
-detector, for both the $\delta$-syllable and for lny64's song at
-$t^*_4$.
-
-\begin{figure}
-  \includegraphics[width=\textwidth]{Fig4}
-  \caption{Timing varies as the FFT frame interval changes.  Here we
-    show results for the ideal detector and the LabVIEW and
-    Swift+serial implementations, for the constructed
-    $\delta$-syllable and for trigger $t^*_4$ of lny64's song.  The
-    lines show latency; error bars are standard deviation (jitter).  Points have
-    been shifted horizontally slightly for clarity; original positions are [0.5 1
-      1.5 2 4] ms.}
-  \label{fig:TimingVsFrame}
-\end{figure}
-
-\subsubsection{Syllable choice}
-
-Syllable choice impacts detector performance, but despite the variety
-of syllables shown here, performance was stable across syllables.
-\fig{fig:TimingVsSyllable} shows measured timing data for lny64's test
-syllables for the Swift+serial detector compared to the ideal, with
-$t_{\textrm{fft}}=1.5$ ms.  Given the variety of the test syllables,
-the variability is surprisingly low, and is summarised here with 95\%
-confidence intervals (with n=6):
-\vspace{8pt}\par\noindent
-\begin{tabular}{r|cc}
-  Detector & Latency & Jitter \\ 
-  \hline   Ideal & $-0.8\pm 0.3$ ms & $2.1\pm 0.1$ ms \\
-  Swift+serial & $0.0\pm 0.6$ ms & $2.0\pm 0.1$ ms
-\end{tabular}
-\vspace{8pt}\par\noindent
-The negative latency is due to the way in which the network responds
-to the song through time: as the recognition region looks increasingly
-similar to the trained match, the network's evaluation of similarity
-rises, and will generally cross the triggering threshold before it
-reaches its maximum value.  A heuristic as simple as triggering at an
-apparent turning point after crossing the threshold might improve
-timing consistency at the expense of latency, but we did not test this.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=8cm]{Fig5}
-  \end{center}
-  \caption{Timing data for lny64's 6 test syllables, for the ideal and the
-    Swift+serial detectors, with an FFT frame rate of 1.5 ms.  Point
-    centres show latency; error bars show jitter.}
-  \label{fig:TimingVsSyllable}
-\end{figure}
-
-\subsubsection{Detector Implementations}
-
-We compared latency and jitter across our different detector
-implementations for the $\delta$-syllable and lny64's $t^*_4$, again with
-$t_\mathrm{fft}=1.5$ ms.  Results are as follows:
-\vspace{8pt}
-\begin{tabular}{l|c|c|c|c}
-  & \multicolumn{2}{c}{$\delta$-syllable} & \multicolumn{2}{c}{Bird: $t^*_4$} \\
-  Detector & Latency (ms) & Jitter (ms) & Latency (ms) & Jitter (ms) \\
-  \hline
-  Ideal & 0.66 & 0.38 & -1.0 & 2.0 \\
-  LabVIEW & 1.5 & 0.69 & -0.18 & 2.4 \\
-  Swift+serial & 0.88 & 1.6 & -0.075 & 2.0 \\
-  Swift+audio & 5.9 & 2.6 & 1.2 & 4.9 \\
-  MATLAB+serial & 8.1 & 2.3 & 3.8 & 2.0 \\
-  MATLAB+audio & 13 & 5.9 & 8.4 & 5.9
-\end{tabular}
-\vspace{8pt}\par\noindent
-These data are also plotted in \fig{fig:TimingVsDetector}, and
-\fig{fig:timing} gives a more detailed view of what the timing curves
-look like for the five implementations of our detector on $t^*_4$.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=8cm]{Fig6}
-  \end{center}
-  \caption{The different detectors for the constructed
-    $\delta$-syllable and for lny64's song at $t^*_4$.  Point centres
-    show latency; error bars show jitter.}
-  \label{fig:TimingVsDetector}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=8cm]{Fig7}
-  \end{center}
-  \caption{Raw timing curves for all detectors measured during
-    detection of lny64's $t^*_4$ using 1.5-ms frames.  We extract the trigger events from each curve, from which we obtain the mean---latency---and
-    standard deviation---jitter.}
-  \label{fig:timing}
-\end{figure}
-
-\section{Discussion}
-\label{sec:conclusion}
-
-This syllable detector is appropriate for zebra finch song, and
-although our tests were carried out on songs from that species, it is
-also likely to work well for Bengalese finches due to their similarly stereotyped syllables.  It offers the
-following benefits:
-\begin{itemize}
-\item The detector is accurate. False negative and false positive
-  rates can be well under 1\% and 0.005\% respectively, and trading
-  these two numbers off against each other is through a single
-  relative-cost parameter.
-\item Latency is generally under a millisecond, with jitter around 2
-  ms.
-\item Works on a wide range of target syllables using the default
-  values described here, generally eliminating the need for
-  hand-tuning.
-\item Runs fast enough for syllable-modification experiments on
-  inexpensive consumer-grade hardware, although we recommend that the
-  training phase be run on a fast desktop system with 32 GB of RAM.
-\item A single detector can generate different target pulses for
-  multiple syllables at almost no additional computational cost during
-  runtime, although training time will increase.
-\end{itemize}
-
-Although there are differences, the Swift+serial detector and the
-LabVIEW implementation are roughly comparable in performance.  We
-prefer the Swift implementation due to its lower hardware and software
-requirements and the difficulty of debugging LabVIEW programmes.  With
-serial output, MATLAB's performance is good, although its buffer
-handling is sensitive to system load.
-
-The songs presented here were recorded with fixed microphones mounted
-inside the cages.  We found that higher accuracy is achieved when a
-microphone is instead mounted on the bird's head, which maintains
-volume and reduces changes in timbre as the bird moves.
-% We have head mounted microphone data?
-
-A common experimental paradigm requires detecting the frequency of
-syllables.  Many pitch detection techniques rely on the spectrum,
-which incurs no additional computational cost here since it is already
-available.  For example, \cite{Canopoli2014} achieved good results
-with the Harmonic Product Spectrum algorithm
-\cite{Noll1970pitchdetection}.
-
-In order to monitor syllable duration, the beginning and end of a
-syllable may be detected by looking at the ratio of total energy in
-the singing frequency band to the total energy, over some small time
-window.  Any syllable thus identified that also contains a trigger
-event may be monitored for duration.  Alternatively, the network can
-be trained to recognise both the beginning and the end of the syllable
-of interest.
-
-In feedback experiments such as frequency- or duration-shifting, vocal
-output changes over the course of the experiment.  The neural network
-excels at identifying syllables close to its training set, so as vocal
-output changes the detector may not recognise a match.  If the
-detector must be robust to this shift, it may be retrained as often as
-necessary as the bird learns, or data consisting of synthetically
-pitch-shifted or duration-shifted target syllables over the
-recognition region may be added to the training set.  We will test
-these approaches in future work.
-
-\section{Acknowledgments}
-We would like to thank Winthrop F.~Gillis, William A.~Liberti, and Sanne Moorman for
-their improvements to the draft.  This work was funded by NIH grants
-5R01NS089679-02 and 5U01NS090454-02.  The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
-
 \appendix
 
-\section{Resources}
-\label{sec:resources}
+\section{Online Resources}
 
-\begin{description}
-\item[Data:]
+\subsection{Data}
 Data and training configuration files for the 8 birds discussed here are available from the Open Science Framework at DOI: 10.17605/OSF.IO/BX76R .
 DOIs for the versions of the programmes used here are provided for reproducibility; however, we generally recommend downloading the latest release.
-%\item[Licenses:]
-%  All software is licensed under Open-Source-Initiative--approved licenses. Please see the software packages for individual choices.
-\item[Song alignment:] Jeff Markowitz's song alignment software \cite{Poole2012} can be found at
-    {\tt https://github.com/jmarkow/zftftb}
-\item[Training the neural network:]\hfill
+
+\subsection{Song alignment}
+Jeff Markowitz's song alignment software [11] can be found at
+\url{https://github.com/jmarkow/zftftb}
+
+\subsection{Training the neural network}
 \begin{trivlist}
-  \item {\tt https://github.com/gardner-lab/syllable-detector-learn}
+\item \url{https://github.com/gardner-lab/syllable-detector-learn}
 \item DOI: 10.5281/zenodo.437555
@@
 \begin{description}
-  \item[Installation:]\hfill\\
+\item[Installation:]\hfill
 See {\tt README.md} in the repository for complete, up-to-date instructions.
@@
     \end{itemize}
+
 \item[Usage:]\hfill
@@
 
-\item[Runtime:]\hfill
-  \begin{description}
+\subsection{Runtime}
 
-  \item[MATLAB:]\hfill
+\subsubsection{MATLAB}
     \begin{trivlist}
-    \item {\tt https://github.com/gardner-lab/syllable-detector-matlab}
+    \item \url{https://github.com/gardner-lab/syllable-detector-matlab}
     \item DOI: 10.5281/zenodo.437557
@@
 
-  \item[Swift:]\hfill
+\subsubsection{Swift}
     \begin{trivlist}
-    \item {\tt https://github.com/gardner-lab/syllable-detector-swift}
+    \item \url{https://github.com/gardner-lab/syllable-detector-swift}
     \item DOI: 10.5281/zenodo.437559
@@
 
-  \item[LabVIEW:]\hfill
+\subsubsection{LabVIEW}
     \begin{trivlist}
-    \item {\tt https://github.com/gardner-lab/syllable-detector-runtime-labview}
-    \item 10.5281/zenodo.437558 
+    \item \url{https://github.com/gardner-lab/syllable-detector-runtime-labview}
+    \item DOI: 10.5281/zenodo.437558 
     \end{trivlist}
@@
     \end{description}
-  \end{description}
-
-\end{description}
-
-%\bibliography{birds}
-\begin{thebibliography}{10}
-
-\bibitem{Kao2005}
-Kao MH, Doupe AJ, Brainard MS.
-\newblock Contributions of an avian basal ganglia--forebrain circuit to
-  real-time modulation of song.
-\newblock Letters to Nature. 2005 February;433:638--643.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v433/n7026/abs/nature03127.html}.
-
-\bibitem{Wang2008interhemispheric}
-Wang CZH, Herbst JA, Keller GB, Hahnloser RHR.
-\newblock Rapid Interhemispheric Switching during Vocal Production in a
-  Songbird.
-\newblock PLOS Biology. 2008 October;6(10).
-\newblock Available from:
-  \url{http://journals.plos.org/plosbiology/article?id=10.1371/journal.pbio.0060250}.
-
-\bibitem{Keller2009}
-Keller GB, Hahnloser RHR.
-\newblock Neural processing of auditory feedback during vocal practice in a
-  songbird.
-\newblock Nature. 2009 January;457:187--190.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v457/n7226/abs/nature07467.html}.
-
-\bibitem{Tumer2007pitchshift}
-Tumer EC, Brainard MS.
-\newblock Performance variability enables adaptive plasticity of 'crystallized'
-  adult birdsong.
-\newblock Nature. 2007 December;450:1240--1244.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v450/n7173/abs/nature06390.html}.
-
-\bibitem{Andalman2009}
-Andalman AS, Fee MS.
-\newblock A basal ganglia-forebrain circuit in the songbird biases motor output
-  to avoid vocal errors.
-\newblock Proceedings of the National Academy of Sciences of the United States
-  of America. 2009 July;106(30):12518--12523.
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC2709669/}.
-
-\bibitem{Charlesworth2011learning}
-Charlesworth JD, Tumer EC, Warren TL, Brainard MS.
-\newblock Learning the microstructure of successful behavior.
-\newblock Nature Neuroscience. 2011;14(3):373--380.
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3045469/}.
-
-\bibitem{Warren2011}
-Warren TL, Tumer EC, Charlesworth JD, Brainard MS.
-\newblock Mechanisms and time course of vocal learning and consolidation in the
-  adult songbird.
-\newblock Journal of Neurophysiology. 2011 October;106(4):1806--1821.
-\newblock Available from:
-  \url{http://jn.physiology.org/content/106/4/1806.full}.
-
-\bibitem{Canopoli2014}
-Canopoli A, Herbst JA, Hahnloser RHR.
-\newblock A Higher Sensory Brain Region Is Involved in Reversing
-  Reinforcement-Induced Vocal Changes in a Songbird.
-\newblock The Journal of Neuroscience. 2014;34(20):7018--7026.
-\newblock Available from:
-  \url{http://www.jneurosci.org/content/34/20/7018.full}.
-
-\bibitem{Leonardo1999}
-Leonardo A, Konishi M.
-\newblock Decrystallization of adult birdsong by perturbation of auditory
-  feedback.
-\newblock Nature. 1999 June;399:466--470.
-\newblock Available from:
-  \url{http://www.nature.com/nature/journal/v399/n6735/abs/399466a0.html}.
-
-\bibitem{Skocik2013}
-Skocik M, Kozhevnikov A.
-\newblock Real-time system for studies of the effects of acoustic feedback on
-  animal vocalizations.
-\newblock Frontiers in Neural Circuits. 2013 January;6(111).
-\newblock Available from:
-  \url{http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3539774/}.
-
-\bibitem{Poole2012}
-Poole B, Markowitz JE, Gardner TJ.
-\newblock The Song Must Go On: Resilience of the Songbird Vocal Motor Pathway.
-\newblock PLOS One. 2012 June;7(6).
-\newblock Available from:
-  \url{http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0038173}.
-
-\bibitem{Tchernichovski2011SAP}
-Tchernichovski O. Sound Analysis Pro User Manual; 2011.
-\newblock Available from: \url{http://soundanalysispro.com/manual-1}.
-
-\bibitem{Jacobsen2003SlidingDFT}
-Jacobsen E, Lyons R.
-\newblock The sliding DFT.
-\newblock IEEE Signal Processing Magazine. 2003 Mar;20(2):74--80.
-\newblock Available from:
-  \url{http://www.cmlab.csie.ntu.edu.tw/DSPCourse/reference/Sliding\%20DFT.pdf}.
-
-\bibitem{hkp}
-Hertz JA, Krogh AS, Palmer RG.
-\newblock Introduction to the Theory of Neural Computation.
-\newblock Perseus Books; 1991.
-
-\bibitem{Noll1970pitchdetection}
-Noll AM.
-\newblock Pitch determination of human speech by the harmonic product spectrum,
-  the harmonic sum spectrum, and a maximum likelihood estimate.
-\newblock In: Proceedings of the symposium on computer processing in
-  communications. vol.~19; 1970. p. 779--797.
-
-\end{thebibliography}
 

### 2017-07-20 main.tex
They want the appendix after the bibliography. Done.
Moved the Resources into a separate file, reformatted with sections, added links. Whew, they've actually accepted this one!

@@
 
-\appendix
-
-\section{Resources}
-\label{sec:resources}
-
-\begin{description}
-\item[Data:]
-  Data and training configuration files for the 8 birds discussed here are available from the Open Science Framework at DOI: 10.17605/OSF.IO/BX76R .
-DOIs for the versions of the programmes used here are provided for reproducibility; however, we generally recommend downloading the latest release.
-%\item[Licenses:]
-%  All software is licensed under Open-Source-Initiative--approved licenses. Please see the software packages for individual choices.
-\item[Song alignment:] Jeff Markowitz's song alignment software \cite{Poole2012} can be found at
-    {\tt https://github.com/jmarkow/zftftb}
-\item[Training the neural network:]\hfill
-  \begin{trivlist}
-  \item {\tt https://github.com/gardner-lab/syllable-detector-learn}
-    \item DOI: 10.5281/zenodo.437555
-  \end{trivlist}
-
-  \begin{description}
-  \item[Installation:]\hfill\\
-    See {\tt README.md} in the repository for complete, up-to-date instructions.
-    \begin{itemize}
-    \item Clone the Github repository onto your computer, and add it to your MATLAB path.
-    \item You also need the following packages from MATLAB File Exchange in your MATLAB path:
-      \begin{description}
-      \item[Significant Figures] by Teck Por Lam.  File ID \#10669.
-      \item[Generate maximally perceptually-distinct colors] by Tim Holy.  File ID \#29702.
-      \end{description}
-    \end{itemize}
-  \item[Usage:]\hfill
-    \begin{itemize}
-    \item The main programme is {\tt learn\_detector.m}.
-    \item The top section of {\tt learn\_detector.m}
-      describes the configuration variables.
-    \item The current directory should contain the data in {\tt
-      song.mat}. The format for this file is specified in {\tt
-      README.md}: a MATLAB file containing aligned song, nonsong
-      (calls, cage noise, etc), and the audio sampling frequency.
-    \item {\tt learn\_detector} also looks in the current directory for
-      a configuration file {\tt params.m}, described in {\tt
-        README.md}.  If the file is empty, the programme will display
-      an image showing the average of the song spectrograms (frequency
-      vs.~time), tell the user how to define syllables of interest,
-      and exit.  If {\tt times\_of\_interest\_ms} is defined, the
-      programme will train the neural network.  Optionally, {\tt
-        params.m} can also override any of the configuration
-      parameters at the top of {\tt learn\_detector.m}.
-    \item Configuration may also be controlled by specifying
-      parameter-value pairs to the function call in the common MATLAB
-      fashion; for example, {\tt
-        learn\_detector('times\_of\_interest\_ms', [150 200])}.
-    \item The software will produce a neural network file, {\tt
-      detector\_bird\_times\_\dots.mat} (the ellipses signify other
-      information encoded in the filename).  It will also produce the
-      audio test file described in Methods.
-    \end{itemize}
-  \end{description}
-
-\item[Runtime:]\hfill
-  \begin{description}
-
-  \item[MATLAB:]\hfill
-    \begin{trivlist}
-    \item {\tt https://github.com/gardner-lab/syllable-detector-matlab}
-    \item DOI: 10.5281/zenodo.437557
-    \end{trivlist}
-    \begin{description}
-    \item[Installation:]\hfill
-      \begin{itemize}
-      \item Clone the Github repository onto your computer.
-      \item Usage requires MATLAB with the Data Acquisition Toolbox and the 
-      Parallel Computing Toolbox.
-      \end{itemize}
-    \item[Connection:]\hfill
-      \begin{itemize}
-      \item Either connect the microphone directly into the microphone jack or through a 
-      preamp into the audio input jack. The current implementation only supports a single 
-      channel of audio.
-      \item To generate TTL signals via the audio output, use the headphone or audio output
-      jack.
-      \item To generate TTL signals via an Arduino, switch to the {\tt arduino} branch, 
-      load the MATLAB Arduino IO sketch onto the Arduino (the sketch is available in the 
-      Github repository, and provides a simple serial protocol for controlling pins). 
-      Connect the Arduino to the computer via USB. The TTL signal will be available on
-      pin 7.
-      \end{itemize}
-    \item[Usage:]\hfill
-      \begin{itemize}
-      \item Switch to the directory of the repository and run {\tt nndetector\_live}. You
-      can optionally specify parameters as string/value pairs. When you run the command,
-      you will be prompted to select an input device, output device and select a ``.mat''
-      file containing the trained detector.
-      \item Note that you may need to tune the input and output buffer based on computer 
-      performance. Try decreasing the size of the buffers until you see warnings about buffer
-      under/overruns, then increase the value slightly. We usually found 2-3ms worked reliably.
-      To specify the buffer sizes, use the {\tt buffer\_size\_input} and 
-      {\tt buffer\_size\_output} parameters.
-      \item Information about the detector is logged to a file called ``detector\_status.log''.
-      \end{itemize}
-    \end{description}
-
-  \item[Swift:]\hfill
-    \begin{trivlist}
-    \item {\tt https://github.com/gardner-lab/syllable-detector-swift}
-    \item DOI: 10.5281/zenodo.437559
-    \end{trivlist}
-    \begin{description}
-    \item[Installation:]\hfill
-      \begin{trivlist}
-      \item Download the most recent release from the Github repository onto your computer.
-      \end{trivlist}
-    \item[Connection:]\hfill
-      \begin{itemize}
-      \item Either connect the microphone directly into the microphone jack or through a 
-      preamp into the audio input jack. If the input jack supports multiple channels (e.g., 
-      audio input jacks are usually stereo), you can connect multiple audio sources and run 
-      different detectors on each channel simultaneously.
-      \item To generate TTL signals via the audio output, use the headphone jack. The 
-      audio output must have at least the same number of channels as the input.
-      \item To generate TTL signals via an Arduino, load the MATLAB Arduino IO sketch 
-      onto the Arduino (the sketch is available in the Github repository, and provides a 
-      simple serial protocol for controlling pins). Connect the Arduino 
-      to the computer via USB. The TTL signal for the first input will be available on 
-      pin 7, the second input on pin 8, etc.
-      \end{itemize}
-    \item[Usage:]\hfill
-      \begin{itemize}
-      \item Use {\tt convert\_to\_text.m} to convert the Matlab output file to a format 
-      easily readable by Swift. This will generate a text version of the neural network.
-      \item Launch the application downloaded above. It will provide an 
-      menu to select an audio input and either an audio output or an Arduino output (if 
-      detected). Once you select an input and an output, a new window will show all audio 
-      channels associated with the input. For each channel, select the text-encoded version of 
-      the detector created in the previous step. Once configured, press run to begin 
-      monitoring the inputs.
-      \end{itemize}
-    \item[Customisation:]\hfill
-      \begin{itemize}
-      \item You can clone the Github repository and use Xcode to edit the project, if you 
-      are interested in extending or modifying the functionality. To hook into either 
-      the raw input or output, look at the ``Processor.swift'' file. The 
-      {\tt Processor} class acts as a coordinator, receiving incoming audio, passing that
-      to the detector and generating appropriate output TTL signals.
-      \end{itemize}
-    \end{description}
-
-  \item[LabVIEW:]\hfill
-    \begin{trivlist}
-    \item {\tt https://github.com/gardner-lab/syllable-detector-runtime-labview}
-    \item 10.5281/zenodo.437558 
-    \end{trivlist}
-    \begin{description}
-    \item[Installation:]\hfill
-      \begin{trivlist}
-      \item Clone the Github repository onto your computer.
-      \end{trivlist}
-    \item[Connection:]\hfill
-      \begin{itemize}
-      \item The microphone should be connected, optionally through a preamplifier, to the data acquisition card, by default on Analog Input 0 (AI0).
-      \item Outputs are on the Digital I/O (DIO) channels.  See the block diagram of {\tt nnfft.vi} for documentation, or just connect to DIO0.
-      \end{itemize}
-    \item[Usage:]\hfill
-      \begin{itemize}
-      \item Use LabVIEW to open {\tt nnfft.vi}.
-        \item A file chooser is provided in the front panel of {\tt nnfft.vi}.  Use it to point to the network file
-      produced by {\tt learn\_detector.m}, and use LabView's ``Run''
-      button to begin detection.
-      \end{itemize}
-    \end{description}
-  \end{description}
-
-\end{description}
-
 %\bibliography{birds}
@@
 
+\section{Supporting information captions}
+
+\begin{description}
+  \item[S1 Appendix A: Online resources.] Where to find the datasets and software introduced in this paper, with an overview of how to install and use the packages.
+\end{description}
+
 \end{document}
